### 2018-09-14 paper/paper.tex
Add paper skeleton

\documentclass[conference]{IEEEtran}
%\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Towards Accelerating checkpointed adjoint computations in full waveform
  inversion using lossy compression
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
}

\maketitle

\begin{abstract}
Full-waveform inversion is an adjoint-based optimization problem in
seismic imaging that processes terabytes of data, often beyond the
memory capacity of most modern computer systems. A common technique to
carry out the adjoint computation in an environment with
less-than-optimal memory is checkpoint-recomputation. In this
technique, a subset of the forward-computation states are stored in
memory checkpoints and the others are recomputed from the stored
states. The compression of checkpoints can increase the number of
checkpoints available, in a given amount of memory, which, in turn,
can reduce the recompute factor, reducing the time-to-solution of the
adjoint computation. This compression, often comes at an additional
computational cost, and an associated loss of precision that must be
accounted for when choosing between compression schemes. In this
paper, we study the effect of compression on checkpoint-recomputation
methods in memory-constrained scenarios for a realistic FWI workload.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}
Introduce fulll waveform inversion as an adjoint based optimization
problem

introduce revolve based checkpointing - cite revolve, pyrevolve

introduction on compression techniques

relate compression to checkpoint compression

\section{Related Work}

Towards efficient backward-in-time adjoint computations using data
compression techniques E.C. Cyr, J.N. Shadid, T. Wildey

Lossy Data Compression with DCT Transforms F. Rubio Dalmau,
M. Hanzich, J. de la Puente and N. Gutierrez

Full seismic waveform tomography for upper-mantle structure in the
Australasian region using adjoint methods
Andreas Fichtner,1 Brian L. N. Kennett,2 Heiner Igel1 and Hans-Peter
Bunge1

\section{Devito, pyRevolve, test case}
short introduction to devito and how to setup a forward-backward
propagator using it - some justification on why this is a realistic
setup even though it might seem like a toy problem

how does pyrevolve implement checkpointing

test case setup.
requirements: should be realistic, numerically stable (to get
realistic compression), big domain,
large number of timesteps
Can we use a SEAM model here?
\section{Compression algorithms}
\subsection{Lossless}
ZLib

variations in compression ratios
\subsection{Lossy}
ZFP
@inproceedings{lossless,
    Acmid = {1126035},
    Address = {Washington, DC, USA},
    Author = {Ratanaworabhan, Paruj and Ke, Jian and Burtscher, Martin},
    Booktitle = {Proceedings of the Data Compression Conference},
    Date-Added = {2015-05-03 20:40:52 +0000},
    Date-Modified = {2018-01-31 10:20:09 +0000},
    Isbn = {0-7695-2545-8},
    Numpages = {10},
    Pages = {133--142},
    Publisher = {IEEE Computer Society},
    Series = {DCC '06},
    Title = {Fast Lossless Compression of Scientific Floating-Point Data},
    Year = {2006}}

@conference{gpu-compression,
    Author = {O'Neil, Molly A. and Burtscher, Martin},
    Booktitle = {Proceedings of the Fourth Workshop on General Purpose Processing on Graphics Processing Units},
    Date-Added = {2015-05-03 22:40:52 +0000},
    Date-Modified = {2015-05-03 22:42:57 +0000},
    Keywords = {GPGPU, floating-point data, lossless data compression, real-time compression},
    Number = {10.1145/1964179.1964189},
    Publisher = {ACM},
    Title = {Floating-point Data Compression at 75 Gb/s on a {GPU}},
    Year = {2011},
}
@inbook{Kaklamanis:2012aa,
    Author = {Kaklamanis, Christos and Papatheodorou, Theodore and Spirakis, PaulG. and Iverson, Jeremy and Kamath, Chandrika and Karypis, George},
    Booktitle = {Euro-Par 2012 Parallel Processing},
    Chapter = {Fast and Effective Lossy Compression Algorithms for Scientific Datasets},
    Da = {2012/01/01},
    Date-Added = {2015-05-03 22:46:47 +0000},
    Date-Modified = {2018-01-31 10:20:09 +0000},
    Isbn = {978-3-642-32819-0},
    La = {English},
    Pages = {843-856},
    Publisher = {Springer Berlin Heidelberg},
    Se = {83},
    Title = {Lecture Notes in Computer Science},
    Title1 = {Fast and Effective Lossy Compression Algorithms for Scientific Datasets},
    Ty = {CHAP},
    Volume = {7484},
    Year = {2012}}

p-adaptive compression from Boehm et al. 
\section{Results}
\subsection{Error analysis}
Discuss the error incurred due to lossy compression. This could be the
angular deviation in the gradient from Boehm et al. as a factor of the
precision parameter, for about 3 different compression schemes  What if a minor
deviation is at a point where you're pointing directly at a
constraint? or a corner point between two constraints? The line search
may mistakenly terminate early. This should be used to derive a bound
for compression ratio based on desired precision.

\subsection{Compression vs recomputation}

- Time to solution for different compression ratios, for multiple
compression algorithms. at least one lossless compression included
here.

at least 3 compression algorithms compared

vary compression ratio within bound established in previous section

\section{Conclusions and Future work}

What are the problems?

Disk checkpointing

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}

### 2018-09-25 difference.py
Add fixed rate testing as well as openmp

@@ -3,30 +3,67 @@ from simple import overthrust_setup
 import h5py
 from devito import TimeFunction
 import numpy as np
+import matplotlib.animation as animation
+import matplotlib.pyplot as plt
+from matplotlib import cm
 
+
+def plot_field(u):
+    return plt.imshow(np.transpose(u[0]),
+                    animated=True, vmin=-1e0, vmax=1e0,
+                    cmap=cm.RdGy)
+
+def init_video(solver, correct_u, lossy_u):
+    model = solver.model
+    fig = plt.figure(figsize=(24, 10))
+    im_correct_u = plot_field(correct_u)
+    plt.xlabel('X position (km)',  fontsize=20)
+    plt.ylabel('Depth (km)',  fontsize=20)
+    plt.tick_params(labelsize=20)
+    im_model = plt.imshow(np.transpose(model.vp[0]), vmin=1.5, vmax=4.5, cmap=cm.jet,
+                 alpha=.4)
+    im_lossy_u = plot_field(lossy_u)
+    return fig, im_correct_u, im_lossy_u, im_model
+
+def finalize_video(fig, ims, filename):
+    ani = animation.ArtistAnimation(fig, ims, interval=50, blit=True, repeat_delay=1000)
+    ani.save(filename)
 
 def run(original, lossy, filename='', space_order=4, kernel='OT2', nbpml=40, **kwargs):
 
     solver = overthrust_setup(filename=filename, nbpml=nbpml, space_order=space_order, kernel=kernel, **kwargs)
-
+    # Load two timesteps of each
+    # Uncompressed should be saved by simple.py
+    # Compressed one should be generated on demand by makefile
+    # (dedicated script to convert uncompressed previous timestep to compressed)
     with h5py.File(original) as original_file:
         correct_u = original_file['data'][()]
     with h5py.File(lossy) as lossy_file:
         lossy_u = lossy_file['data'][()]
+
     u = TimeFunction(name="u", grid=solver.model.grid, time_order=2, space_order=solver.space_order)
+    fig, im_correct_u, im_lossy_u, im_model = init_video(solver, correct_u, lossy_u)
+    im_corrects = [[im_correct_u, im_model]]
+    im_lossys = [[im_lossy_u, im_model]]
     errors = []
     print("steps, error_norm, max_error")
-    for steps in range(1, 1000, 10):
+    for steps in range(1, 50, 10):
[…]

### 2018-09-25 rate.py

@@ -1,0 +1,24 @@
+import h5py
+import sys
+from util import to_hdf5
+import numpy as np
+from contexttimer import Timer
+from zfp import compress, decompress
+
+
+filename = sys.argv[1]
+
+f = h5py.File(filename, 'r')
+uncompressed = f['data'][()].astype(np.dtype('float32'))
+print("\"Size of compressed field\", \"Compression Factor\", \"Compression time\", \"Decompression time\", \"Rate\", \"Error norm\", \"Maximum error\"")
+for p_i in range(1, 10):
+    rate = p_i
+    with Timer(factor=1000) as t:
+        compressed = compress(uncompressed, rate=rate)
+    
+    with Timer(factor=1000) as t2:
+        decompressed = decompress(compressed, uncompressed.shape, uncompressed.dtype, rate=rate)
+
+    to_hdf5(decompressed, "decompressed-r-%d.h5"%p_i)
+    error_matrix = decompressed-uncompressed
+    print("%f, %f, %f, %f, %f, %f, %f" % (len(compressed), len(uncompressed.tostring())/float(len(compressed)), t.elapsed, t2.elapsed, rate, np.linalg.norm(error_matrix), np.max(error_matrix)))

### 2018-09-25 simple.py

@@ -68,8 +68,10 @@ def run(space_order=4, kernel='OT2', nbpml=40, filename='', **kwargs):
     solver = overthrust_setup(filename=filename, nbpml=nbpml, space_order=space_order, kernel=kernel, **kwargs)
 
     rec, u, summary = solver.forward(save=False)
-    uncompressed = u.data[0]
-    to_hdf5(uncompressed, "uncompressed.h5")
+    last_time_step = solver.nt - 1
+    uncompressed = u
+    to_hdf5(uncompressed[last_time_step], "uncompressed.h5")
+    to_hdf5(uncompressed[last_time_step-1], "uncompressed.h5.1")
     
 
 

### 2018-09-25 video.py

@@ -1,0 +1,20 @@
+fig = plt.figure(figsize=(24, 10))
+im = plt.imshow(np.transpose(u.data[0,40:-40,40:-40]), animated=True, vmin=-1e0, vmax=1e0, cmap=cm.RdGy, aspect=1.5,
+                extent=[origin[0], origin[0] + 1e-3 * dimensions[0] * spacing[0],
+                        origin[1] + 1e-3*dimensions[1] * spacing[1], origin[1]])
+plt.xlabel('X position (km)',  fontsize=20)
+plt.ylabel('Depth (km)',  fontsize=20)
+plt.tick_params(labelsize=20)
+im2 = plt.imshow(np.transpose(vp), vmin=1.5, vmax=4.5, cmap=cm.jet, aspect=1.5,
+                 extent=[origin[0], origin[0] + 1e-3 * dimensions[0] * spacing[0],
+                         origin[1] + 1e-3*dimensions[1] * spacing[1], origin[1]], alpha=.4)
+import matplotlib.animation as animation
+
+def updatefig(i):
+    im.set_array(np.transpose(u.data[i,40:-40,40:-40]))
+    im2.set_array(np.transpose(vp))
+    return im, im2
+
+ani = animation.FuncAnimation(fig, updatefig, frames=np.linspace(0, nt), blit=True)
+ani.save('Adjoint.mp4')
+plt.show()

### 2018-09-25 zfp.py

@@ -26,6 +26,8 @@ class ZfpField(ctypes.Structure):
 libzfp = cdll.LoadLibrary("libzfp.so")
 libzfp.zfp_stream_set_accuracy.argtypes = (ctypes.POINTER(ZfpStream), ctypes.c_double)
 libzfp.zfp_stream_set_precision.argtypes = (ctypes.POINTER(ZfpStream), ctypes.c_int)
+libzfp.zfp_stream_set_rate.argtypes = (ctypes.POINTER(ZfpStream), ctypes.c_double,
+                                       ctypes.c_int, ctypes.c_int, ctypes.c_int)
 libzfp.zfp_stream_maximum_size.argtypes = (ctypes.c_void_p, ctypes.c_void_p)
 libzfp.zfp_stream_set_bit_stream.argtypes = (ctypes.c_void_p, ctypes.c_void_p)
 libzfp.zfp_stream_rewind.argtypes = (ctypes.c_void_p,)
@@ -34,7 +36,6 @@ libzfp.zfp_field_1d.restype = ctypes.POINTER(ZfpField)
 libzfp.zfp_field_2d.restype = ctypes.POINTER(ZfpField)
 libzfp.zfp_field_3d.restype = ctypes.POINTER(ZfpField)
 libzfp.stream_open.restype = ctypes.POINTER(BitStream)
-#libzfp.zfp_stream_open.restype = ctypes.c_void_p
 
 
 
@@ -42,14 +43,16 @@ def raw_pointer(numpy_array):
     #return numpy_array.__array_interface__['data'][0]
     return ctypes.c_void_p(numpy_array.ctypes.data)
 
-def compress(indata, tolerance=None, precision=None):
-    assert(tolerance or precision)
+def compress(indata, tolerance=None, precision=None, rate=None, parallel=True):
+    assert(tolerance or precision or rate)
     assert(not(tolerance is not None and precision is not None))
+    assert(not(tolerance is not None and rate is not None))
+    assert(not(rate is not None and precision is not None))
     zfp_types = {np.dtype('float32'): 3, np.dtype('float64'): 4}
     zfp_fields = {1: libzfp.zfp_field_1d, 2: libzfp.zfp_field_2d, 3: libzfp.zfp_field_3d}
     data_type = zfp_types[indata.dtype]
     status = 1
-    shape = indata.shape
+    shape = list(reversed(indata.shape))
     
[…]

### 2018-09-25 paper/paper.tex

@@
 
-\title{Towards Accelerating checkpointed adjoint computations in full waveform
-  inversion using lossy compression
+\title{Combining checkpointing and data compression\\ for high frequency seismic imaging
 }
@@
 \section{Introduction}
-Introduce fulll waveform inversion as an adjoint based optimization
-problem
-
 introduce revolve based checkpointing - cite revolve, pyrevolve
@@
 relate compression to checkpoint compression
+
+Briefly introduce full waveform inversion as an adjoint based optimization
+problem. This can be relatively short, as this is only one of many use cases
+for our technique.
 
@@
 Can we use a SEAM model here?
+
 \section{Compression algorithms}
@@
 
+For this paper, we can probably focus on only zfp using 1-3 different precision settings
+(nearly lossless, very lossy, something reasonable).
+
 - Time to solution for different compression ratios, for multiple
@@
 
-at least 3 compression algorithms compared
+%at least 3 compression algorithms compared
 
-vary compression ratio within bound established in previous section
+%vary compression ratio within bound established in previous section
 
@@
 What are the problems?
+
+Future work:
+\begin{itemize}
+\item Acceptable error bounds
+\item A posteriori and a priori metrics
+\item Relationship between compression error and field error, propagation of errors in forward simulation
+\item Relationship between compression error and gradient error in cross-correlation
+\item What else?
+\end{itemize}
 

### 2018-09-27 precision.py
Parallel compression mode. Update scripts to use parallel mode. New Makefile for reduced results

@@ -4,9 +4,22 @@ from util import to_hdf5
 import numpy as np
 from contexttimer import Timer
 from zfp import compress, decompress
+from argparse import ArgumentParser
 
 
-filename = sys.argv[1]
+
+
+description = ("Test for fixed precision mode of zfp")
+parser = ArgumentParser(description=description)
+parser.add_argument("filename", type=str, help="Filename")
+p_parser = parser.add_mutually_exclusive_group(required=False)
+p_parser.add_argument('--parallel', dest='parallel', action='store_true')
+p_parser.add_argument('--no-parallel', dest='parallel', action='store_false')
+parser.set_defaults(parallel=True)
+args = parser.parse_args()
+
+filename = args.filename
+parallel = args.parallel
 
 f = h5py.File(filename, 'r')
 uncompressed = f['data'][()].astype(np.dtype('float32'))
@@ -14,10 +27,10 @@ print("\"Size of compressed field\", \"Compression Factor\", \"Compression time\
 for p_i in range(6, 20):
     precision = p_i
     with Timer(factor=1000) as t:
-        compressed = compress(uncompressed, precision=precision)
+        compressed = compress(uncompressed, precision=precision, parallel=parallel)
     
     with Timer(factor=1000) as t2:
-        decompressed = decompress(compressed, uncompressed.shape, uncompressed.dtype, precision=precision)
+        decompressed = decompress(compressed, uncompressed.shape, uncompressed.dtype, precision=precision, parallel=parallel)
 
     to_hdf5(decompressed, "decompressed-p-%d.h5"%p_i)
     error_matrix = decompressed-uncompressed

### 2018-09-27 rate.py

@@ -4,9 +4,22 @@ from util import to_hdf5
 import numpy as np
 from contexttimer import Timer
 from zfp import compress, decompress
+from argparse import ArgumentParser
 
 
-filename = sys.argv[1]
+
+
+description = ("Test for fixed rate mode of zfp")
+parser = ArgumentParser(description=description)
+parser.add_argument("filename", type=str, help="Filename")
+p_parser = parser.add_mutually_exclusive_group(required=False)
+p_parser.add_argument('--parallel', dest='parallel', action='store_true')
+p_parser.add_argument('--no-parallel', dest='parallel', action='store_false')
+parser.set_defaults(parallel=True)
+args = parser.parse_args()
+
+filename = args.filename
+parallel = args.parallel
 
 f = h5py.File(filename, 'r')
 uncompressed = f['data'][()].astype(np.dtype('float32'))
@@ -14,10 +27,10 @@ print("\"Size of compressed field\", \"Compression Factor\", \"Compression time\
 for p_i in range(1, 10):
     rate = p_i
     with Timer(factor=1000) as t:
-        compressed = compress(uncompressed, rate=rate)
+        compressed = compress(uncompressed, rate=rate, parallel=parallel)
     
     with Timer(factor=1000) as t2:
-        decompressed = decompress(compressed, uncompressed.shape, uncompressed.dtype, rate=rate)
+        decompressed = decompress(compressed, uncompressed.shape, uncompressed.dtype, rate=rate, parallel=parallel)
 
     to_hdf5(decompressed, "decompressed-r-%d.h5"%p_i)
     error_matrix = decompressed-uncompressed

### 2018-09-27 simple.py

@@ -68,7 +68,7 @@ def run(space_order=4, kernel='OT2', nbpml=40, filename='', **kwargs):
     solver = overthrust_setup(filename=filename, nbpml=nbpml, space_order=space_order, kernel=kernel, **kwargs)
 
     rec, u, summary = solver.forward(save=False)
-    last_time_step = solver.nt - 1
+    last_time_step = len(rec) - 1
     uncompressed = u
     to_hdf5(uncompressed[last_time_step], "uncompressed.h5")
     to_hdf5(uncompressed[last_time_step-1], "uncompressed.h5.1")

### 2018-09-27 batch-submit.sh
Modify batch submit

@@ -3,15 +3,15 @@
 #cd compression
 export PYTHONUNBUFFERED=1
 export KMP_HW_SUBSET=28c,1t
-export PATH=$PATH:~/intel/bin:/home/nkukreja1/compression/zfp-0.5.3/lib
+export PATH=$PATH:~/intel/bin
 export DEVITO_ARCH=intel
 export DEVITO_OPENMP=1
 export DEVITO_LOGGING=DEBUG
-source activate devito
+source activate devito2
 export PYTHONPATH=$PYTHONPATH:~/checkpointer
-export LD_LIBRARY_PATH=$LD_LIBRARY_PATH:/home/nkukreja1/compression/zfp-0.5.3/lib
+export LD_LIBRARY_PATH=$LD_LIBRARY_PATH:/home/nkukreja1/compression/zfp-0.5.3/lib:~/intel/lib
 python ~/devito/scripts/clear_devito_cache.py
-[ -z "$CHUNK_SIZE" ] && echo "Need to set CHUNK SIZE" && exit 1;
-[ -z "$ALGO" ] && echo "Need to set ALGO" && exit 1;
-[ -z "$SHUFFLE" ] && echo "Need to set SHUFFLE" && exit 1;
-python run-experiment.py -c $CHUNK_SIZE -s $SHUFFLE -z $ALGO
+#[ -z "$CHUNK_SIZE" ] && echo "Need to set CHUNK SIZE" && exit 1;
+#[ -z "$ALGO" ] && echo "Need to set ALGO" && exit 1;
+#[ -z "$SHUFFLE" ] && echo "Need to set SHUFFLE" && exit 1;
+#python run-experiment.py -c $CHUNK_SIZE -s $SHUFFLE -z $ALGO

### 2018-09-27 simple.py
Update simple

@@ -68,7 +68,7 @@ def run(space_order=4, kernel='OT2', nbpml=40, filename='', **kwargs):
     solver = overthrust_setup(filename=filename, nbpml=nbpml, space_order=space_order, kernel=kernel, **kwargs)
 
     rec, u, summary = solver.forward(save=False)
-    last_time_step = len(rec) - 1
+    last_time_step = rec.shape[0] - 1
     uncompressed = u
     to_hdf5(uncompressed[last_time_step], "uncompressed.h5")
     to_hdf5(uncompressed[last_time_step-1], "uncompressed.h5.1")

### 2018-09-27 simple.py
Simple.py isn't able to save two timesteps separately so disable it

@@ -70,8 +70,8 @@ def run(space_order=4, kernel='OT2', nbpml=40, filename='', **kwargs):
     rec, u, summary = solver.forward(save=False)
     last_time_step = rec.shape[0] - 1
     uncompressed = u
-    to_hdf5(uncompressed[last_time_step], "uncompressed.h5")
-    to_hdf5(uncompressed[last_time_step-1], "uncompressed.h5.1")
+    to_hdf5(uncompressed, "uncompressed.h5")
+    #to_hdf5(uncompressed[last_time_step-1, :], "uncompressed.h5.1")
     
 
 

### 2018-09-27 simple.py
Revert to old way of saving snapshot in hdf5

@@ -69,7 +69,7 @@ def run(space_order=4, kernel='OT2', nbpml=40, filename='', **kwargs):
 
     rec, u, summary = solver.forward(save=False)
     last_time_step = rec.shape[0] - 1
-    uncompressed = u
+    uncompressed = u.data[0]
     to_hdf5(uncompressed, "uncompressed.h5")
     #to_hdf5(uncompressed[last_time_step-1, :], "uncompressed.h5.1")
     

### 2018-09-28 performance.py
Add TTI

@@ -152,7 +152,30 @@ def plot(x, y, filename, title, xlabel, ylabel, hline=None, more_y=None, more_y_
             plt.legend()
     plt.savefig(filename, bbox_inches='tight')
 
+
+    
 # Plots
+
+# Problem params
+nt = 2526
+cp_size = 287*881*881*4/1000000
+
+#System params
+compute_ts = 0.36
+bandwidth = 6777
+
+# Compressions Params
+c_factor = 29.72
+c_time = 0.94
+d_time = 1.5
+t_d_time = 0.396
+
+# Plot params
+peak_mem = 2000
+more_peak_mem = [8000, 16000, 24000]
+
+compression_filename="tolerance-richter.csv"
+compression_label = "Tolerance"
 
 # (Speedup vs) varying peak memory
 
@@ -166,13 +189,13 @@ def varying_peak_memory(nt, size_ts, compute_ts, bw, c_factor, c_time, d_time,
     plot(mems, speedups, "varying-memory.png", "Speedup for varying peak memory", "Memory (MB)",
              "Speedup (x)", hline=1, more_y=[theoretical_speedup])
 
-#varying_peak_memory(2000, 287*881*881*4/1000000, 2.276333333, #time to compute one timestep (s)
-#                    7234, # Memory bandwidth from stream (MB/s)
-#                    29.72, # Compression factor
-#                    1.45, # Compression time(s)
-#                    1.65, # Decompression time(s)
-#                    0.7 # Theoretical decompression time (s)
-#                    )
+varying_peak_memory(nt, cp_size, compute_ts, #time to compute one timestep (s)
+                    bandwidth, # Memory bandwidth from stream (MB/s)
+                    c_factor, # Compression factor
+                    c_time, # Compression time(s)
+                    d_time, # Decompression time(s)
+                    t_d_time # Theoretical decompression time (s)
+                    )
 
 # (Speedup vs) varying AI. Fixed compression.
 def varying_compute(nt, size_ts, peak_mem, bw, c_factor, c_time, d_time, more_peak_mem=None):
@@ -192,13 +215,13 @@ def varying_compute(nt, size_ts, peak_mem, bw, c_factor, c_time, d_time, more_pe
     plot(computes, speedups, "varying-compute.png", "Speedup for varying compute time/timestep", "Compute time (s)",
[…]

### 2018-09-28 simple.py

@@ -1,8 +1,9 @@
 from argparse import ArgumentParser
-from examples.seismic.acoustic.acoustic_example import acoustic_setup, smooth10
+from examples.seismic.acoustic.acoustic_example import acoustic_setup
 from devito.logger import info, debug
 from examples.seismic import PointSource, Receiver, TimeAxis, RickerSource
 from examples.seismic.acoustic import AcousticWaveSolver
+from examples.seismic.tti import AnisotropicWaveSolver
 from examples.seismic.model import Model
 from devito import TimeFunction, Function
 import numpy as np
@@ -63,12 +64,44 @@ def overthrust_setup(filename, kernel='OT2', space_order=2, nbpml=40, **kwargs):
                                 space_order=space_order, **kwargs)
     return solver
 
+
+def overthrust_setup_tti(filename, kernel='OT2', space_order=2, nbpml=40, **kwargs):
+    model = from_hdf5(filename, space_order=space_order, nbpml=nbpml, datakey='m0', dtype=np.float32)
+    shape = model.vp.shape
+    spacing = model.shape
+    nrec = shape[0]
+    tn = 4000
+
+    # Derive timestepping from model spacing
+    dt = model.critical_dt
+    t0 = 0.0
+    time_range = TimeAxis(start=t0, stop=tn, step=dt)
+
+    # Define source geometry (center of domain, just below surface)
+    src = RickerSource(name='src', grid=model.grid, f0=0.015, time_range=time_range)
+    src.coordinates.data[0, :] = np.array(model.domain_size) * .5
+    if len(shape) > 1:
+        src.coordinates.data[0, -1] = model.origin[-1] + 2 * spacing[-1]
+    # Define receiver geometry (spread across x, just below surface)
+    rec = Receiver(name='rec', grid=model.grid, time_range=time_range, npoint=nrec)
+    rec.coordinates.data[:, 0] = np.linspace(0., model.domain_size[0], num=nrec)
+    if len(shape) > 1:
+        rec.coordinates.data[:, 1:] = src.coordinates.data[0, 1:]
+
+    # Create solver object to provide relevant operators
+    return AnisotropicWaveSolver(model, source=src, receiver=rec,
+                                 space_order=space_order, **kwargs)
+
 def run(space_order=4, kernel='OT2', nbpml=40, filename='', **kwargs):
 
[…]

### 2018-09-29 performance.py
Explain fixed params on plots

@@ -116,16 +116,17 @@ class Problem(object):
 
         return total_time_revolve/total_time_compression
 
-class AIProblem(object):
-    def __init__(self, nt, size_ts, ai, compute_speed, bw):
-        compute_ts = (ai * size_ts) / compute_speed
-        super(AIProblem, self).__init__(nt, size_ts, compute_ts, bw)
-
 
 def linspace(lower, upper, length):
     return [lower + x*(upper-lower)/length for x in range(length)]
 
-def plot(x, y, filename, title, xlabel, ylabel, hline=None, more_y=None, more_y_labels=None):
+
+def platform_name():
+    global platform
+    return platform
+
+def plot(x, y, filename, title, xlabel, ylabel, hline=None, more_y=None, more_y_labels=None,
+         fixed=None):
     plt.gcf().clear()
     if more_y_labels is not None:
         # We need labels for every series in more_y and one for the main series
@@ -137,7 +138,7 @@ def plot(x, y, filename, title, xlabel, ylabel, hline=None, more_y=None, more_y_
     plt.plot(x, y, label=label)
     plt.xlabel(xlabel)
     plt.ylabel(ylabel)
-    plt.title(title)
+    
     plt.xticks(rotation=90)
     if hline is not None:
         plt.axhline(hline, linestyle='dashed', color='red')
@@ -150,6 +151,12 @@ def plot(x, y, filename, title, xlabel, ylabel, hline=None, more_y=None, more_y_
             plt.plot(x, series, linestyle='dotted', label=label)
         if more_y_labels is not None:
             plt.legend()
+    
+    if fixed is not None:
+        fixed.update({"Platform" : platform_name()})
+        fixed_params = ", \n".join(["%s: %s" % (key, value) for key, value in fixed.items()])
+        title += "\n" + fixed_params
+    plt.title(title)
     plt.savefig(filename, bbox_inches='tight')
 
 
@@ -163,6 +170,7 @@ cp_size = 287*881*881*4/1000000
 #System params
 compute_ts = 0.36
 bandwidth = 6777
+platform = "2 x Intel(R) Xeon(R) CPU E5-2640 v3 @ 2.60GHzr (8C)" # Richter
 
 # Compressions Params
 c_factor = 29.72
@@ -186,15 +194,20 @@ def varying_peak_memory(nt, size_ts, compute_ts, bw, c_factor, c_time, d_time,
     mems = linspace(2*size_ts, overall_peak_mem/c_factor, 200)
[…]

### 2018-10-01 paper/paper.tex
Add performance model to paper

@@
 \usepackage{xcolor}
+\usepackage{cuted}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 
+One possible way of using compression to speed up full-waveform inversion or any adjoint-based
+optimization is by simply compressing the field data before storing it and decompressing it when it
+is required again. If the memory available is sufficient to store the complete forward data after
+compression, the use of compression effectively replaces the use of revolve-based checkpointing.
+This scenario has been discussed in different studies before - some of which we cover in the next
+section. This paper focusses on a broader region, of which this scenario is just one part. In particular,
+we discuss the scenarios where the forward field can not fit in memory even after compression, where
+compression and Revolve would be used in conjunction.
+
+Although most of the compression we discuss here introduces some error which may or may not affect
+the final result of the computation, we try to avoid discussing the error in this paper and focus on the 
+performance impact of compression instead.
+
 \section{Related Work}
@@
 p-adaptive compression from Boehm et al.
+
+\section{Compression in adjoint-based optimization}
+We work with the assumption that the computation of a single forward time step takes the same wall time
+as the computation of a single reverse time step - calling this $\mathbf{C}$. If the size of a single 
+timestep in memory is $\mathbf{S}$ and the simulation involves $\mathbf{N}$ timesteps, the minimum
+wall time required to run a single forward-adjoint evaluation is given by:
+\begin{equation}
+T_N = 2 \times \mathbf{C} \times \mathbf{N}
+\end{equation}
+However, this would require $\mathbf{S} \times \mathbf{N}$ memory. On a system where enough memory
+is available, this would obviously be the optimal strategy to compute the adjoint solution.
+
+When the memory available on a system is less than $\mathbf{S} \times \mathbf{N}$, Revolve provides
+an optimal strategy to solve for the adjoint field by storing a subset of the $\mathbf{N}$ total checkpoints
+and recompute the remaining ones. The overhead introduced by this method can be broken down into
+the recomputation overhead $\mathbf{O}_R$ and the storage overhead $\mathbf{O}_S$. The recomputation
+overhead is largely fixed and given by the following relationship (TODO cite revolve):
+\begin{strip}
+\begin{equation}
+\mathbf{O}_R(N, M) = \begin{cases}
+      N(N-1) /2, & \text{if}\ M=1 \\
+      \min\limits_{1<=\widetilde{N}<=N} \{\widetilde{N} + \mathbf{O}_R(\widetilde{N}, M) + \mathbf{O}_R(N-\widetilde{N}, M-1)\}, & \text{if}\ M>1
+    \end{cases}
+    \label{eqn:recompute}
+\end{equation}
+\end{strip}
+where M is the number of checkpoints that can be stored in memory. Note that for $M >=N$, $\mathbf{O}_R$
+would be zero. For $M < N$, $\mathbf{O}_R$ goes up rather quickly as M is reduced in relation to N.
+
+In a perfect implementation, the storage overhead $\mathbf{O}_S$ might be zero, since the computation could
+be done "in-place", but in practice, checkpoints are generally stored in a separate section of memory and they
+need to be transferred to a "computational" section of the memory where the compute is performed, and then
+the results copied back to the checkpointing memory. This copying is a common feature of checkpointing
+implementations, and might pose a non-trivial overhead in scenarios where the computation is heavily bandwidth-bound. 
+This storage overhead is given by:
+\begin{equation}
+\mathbf{O}_SR = \mathbf{N}_W \times \frac{\mathbf{S}}{\mathbf{B}} + \mathbf{N}_R \times \frac{\mathbf{S}}{\mathbf{B}}
+\label{eqn:storage}
+\end{equation}
+where $\mathbf{N}_W$ is the total number of times Revolve writes checkpoints for a single run, $ \mathbf{N}_R$ 
+is the number of times checkpoints are read, and $\mathbf{B}$ is the memory bandwidth of the target system.
+
+The total time to solution in this scenario becomes:
+\begin{equation}
+T_R = 2 \times \mathbf{C} \times \mathbf{N} + \mathbf{O}_R(N, M) + \mathbf{O}_S
+\end{equation}
+
+Adding compression to the mix will increase the number of checkpoints available ($M$ in equation \ref{eqn:recompute}),
+hence reducing $\mathbf{O}_R$ while adding overheads related to compression and decompression in $\mathbf{O}_S$,
+as well as some error introduced by the compression. Assuming that the error remains within the allowed tolerance,
+we hypothesize that, at least for some circumstances, the reduction in $\mathbf{O}_R$ should offset the increase in 
+$\mathbf{O}_S$, i.e. lead to an overall decrease in the total time to solution ($T$). To build a performance model that
+enables us to find the regions where compression pays off, we assume that the compression algorithm behaves uniformly
+across the different time steps of the simulation, i.e. that we get the same compression ratio, compression time and 
+decompression time, no matter which of the $N$ possible checkpoints we try to compress/decompress. The storage overhead
+now becomes:
+\begin{equation}
+\mathbf{O}_SR = \mathbf{N}_W \times (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} + t_c) + \mathbf{N}_R \times (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} + t_d)
+\end{equation}
+where $\mathbf{F}$ is the compression ratio (i.e. the ratio between the uncompressed and compressed checkpoint), and $t_c$
+and $t_d$ are compression and decompression times, respectively. At the same time, the recomputation overhead goes down
+because $\mathbf{F}$ times more checkpoints are now available.
+
+TODO add figure here that shows memory on x-axis and speedup on y axis. Split into 3 regions. Region 1, 
+where Revolve+Compression makes sense, Region 2, where compression would probably not bring in enough benefit, 
+and region 3 where compression would enable the computation to be done without revolve. 
+\section{Performance model}
+We can predict the performance
 \section{Results}

### 2018-10-01 performance.py

@@ -126,7 +126,7 @@ def platform_name():
     return platform
 
 def plot(x, y, filename, title, xlabel, ylabel, hline=None, more_y=None, more_y_labels=None,
-         fixed=None):
+         fixed=None, xscale=None):
     plt.gcf().clear()
     if more_y_labels is not None:
         # We need labels for every series in more_y and one for the main series
@@ -157,6 +157,8 @@ def plot(x, y, filename, title, xlabel, ylabel, hline=None, more_y=None, more_y_
         fixed_params = ", \n".join(["%s: %s" % (key, value) for key, value in fixed.items()])
         title += "\n" + fixed_params
     plt.title(title)
+    if xscale is not None:
+        plt.xscale(xscale, basex=2)
     plt.savefig(filename, bbox_inches='tight')
 
 
@@ -168,9 +170,9 @@ nt = 2526
 cp_size = 287*881*881*4/1000000
 
 #System params
-compute_ts = 0.36
+compute_ts = 1.11
 bandwidth = 6777
-platform = "2 x Intel(R) Xeon(R) CPU E5-2640 v3 @ 2.60GHzr (8C)" # Richter
+platform = "2 x Intel(R) Xeon(R) CPU E5-2640 v3 @ 2.60GHz (8C)" # Richter
 
 # Compressions Params
 c_factor = 29.72
@@ -179,7 +181,7 @@ d_time = 1.5
 t_d_time = 0.396
 
 # Plot params
-peak_mem = 2000
+peak_mem = 4000
 more_peak_mem = [8000, 16000, 24000]
 
 compression_filename="tolerance-richter.csv"
@@ -193,6 +195,9 @@ def varying_peak_memory(nt, size_ts, compute_ts, bw, c_factor, c_time, d_time,
     min_time, overall_peak_mem = p.naive_strategy()
     mems = linspace(2*size_ts, overall_peak_mem/c_factor, 200)
     speedups = [p.compression_speedup(x, c_factor, c_time, d_time) for x in mems]
+    print(speedups)
+    print("***")
+    print(mems)
     theoretical_speedup = [p.compression_speedup(x, c_factor, c_time, theoretical_d_time) for x in mems]
     fixed = {'Timesteps': nt, 'Size of checkpoint (MB)': size_ts,
              'Time for compute step (s)': compute_ts, 'Bandwidth (MB/s)': bw,
[…]

### 2018-10-02 paper/paper.tex
Filling up introduction sections

@@
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\newcommand*\Laplace{\mathop{}\!\mathbin\bigtriangleup}
 \begin{document}
@@
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
+\author{\IEEEauthorblockN{Navjot Kukreja}
+\IEEEauthorblockA{\textit{Department of Earth Science and Engineering} \\
+\textit{Imperial College London}\\
+London, UK \\
+n.kukreja@imperial.ac.uk}
 \and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
+\IEEEauthorblockN{Jan H\"uckelheim}
+\IEEEauthorblockA{\textit{Department of Earth Science and Engineering}\\
+\textit{Imperial College London}\\
+London, UK }
 \and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
+\IEEEauthorblockN{Mathias Louboutin}
+\IEEEauthorblockA{\textit{TODO Mathias} \\
+\textit{Georgia Institute of Technology}\\
+Atlanta, GA, USA\\}
 \and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
+\IEEEauthorblockN{Kaiyuan Hou}
+\IEEEauthorblockA{\textit{Department of EECS} \\
+\textit{Northwestern University}\\
+Evanston, IL, USA \\
+}
 \and
@@
 \and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
+\IEEEauthorblockN{Gerard Gorman}
+\IEEEauthorblockA{\textit{Department of Earth Science and Engineering}\\
+\textit{Imperial College London}\\
+London, UK }
 }
@@
 \section{Introduction}
-introduce revolve based checkpointing - cite revolve, pyrevolve
+The class of adjoint-based optimization problems has a very typical data access pattern where a
+simulation is run ``forward" in simulation time, which produces data that is to be used by a
+subsequent adjoint pass which is run backward in simulation time TODO CITE. Figure TODO
+shows the typical data access pattern. This class consists of many different problems of interest in
+science and engineering. One problem in this class is full-waveform inversion (FWI) where physical
+experiments are carried out by propagating seismic waves through the earth's subsurface and the
+solution to the optimization problem gives the most likely structure of the earth's subsurface. Figure
+TODO shows the physical setup of the experiment that produces the data that this problem operates
+on. The data access pattern from Figure TODO applies here as well, since every step during the
+adjoint propagation requires data from the corresponding step from the forward propagation. In
+particular for FWI, this data can reach tens of terabytes in size, which is beyond the memory
+available on most hardware. While this paper is written to be focussed on FWI as the target problem, 
+the analysis can be applied to any problem with a similar data-access pattern.
 
-introduction on compression techniques
+Various strategies are routinely employed to reduce this memory requirement, for example 
+Domain decomposition (TODO cite) where the domain of the simulation is distributed over
+multiple nodes so as to fit within the memory available on each node. This introduces the overhead
+of network communication and might not efficiently use the computational resources on each node.
+Another common strategy, which is specific to the domain of FWI, is where only the boundaries of
+the domain are stored at each timestep and the rest of the wavefield is reconstructed from the
+boundaries when required (TODO CITE). Another commonly used strategy that is applied across
+a much wider domain of problems is checkpointing. In this strategy, a subset of the data from the
+forward pass is stored (and the rest discarded). The discarded data is recomputed when required
+by rerunning the forward pass. Starting from Revolve (TODO cite), which was a provably optimal
+scheduling algorithm to implement this strategy, various studies have used and extended this
+technique to various scenarios (TODO cite other checkpointing papers).
 
-relate compression to checkpoint compression
-
-Briefly introduce full waveform inversion as an adjoint based optimization
-problem. This can be relatively short, as this is only one of many use cases
-for our technique.
+introduction on compression techniques - TODO Jan
+Compression is an obvious addition to the tools applied to the specific data access pattern
 
@@
 \section{Related Work}
+\cite{cyr2015towards}
 
-Towards efficient backward-in-time adjoint computations using data
-compression techniques E.C. Cyr, J.N. Shadid, T. Wildey
+\cite{dalmau2014lossy}
 
-Lossy Data Compression with DCT Transforms F. Rubio Dalmau,
-M. Hanzich, J. de la Puente and N. Gutierrez
+\cite{fichtner2009full}
 
-Full seismic waveform tomography for upper-mantle structure in the
-Australasian region using adjoint methods
-Andreas Fichtner,1 Brian L. N. Kennett,2 Heiner Igel1 and Hans-Peter
-Bunge1
+\cite{gpu-compression}
 
-\section{Devito, pyRevolve, test case}
-short introduction to devito and how to setup a forward-backward
-propagator using it - some justification on why this is a realistic
-setup even though it might seem like a toy problem
+\cite{boehm2016wavefield}
+\cite{gpu-compression}
+\cite{Kaklamanis:2012aa}
 
-how does pyrevolve implement checkpointing
+TODO cite SZ (Frank Cappello)
+\section{Test case using Devito and pyRevolve}
+We use Devito \cite{devito-compiler} \cite{devito-api} to setup wave equation solvers for forward and
+adjoint modes. Devito is a domain-specific language to enable the rapid development of finite-difference
+solvers from a high-level description of the partial differential equation. The simplest version of the wave
+equation is the acoustic:
+\begin{equation}
+m(x)\frac{\partial^2 u(t, x)}{\partial t^2} - \Laplace u(t, x) = q(t, x)
+\end{equation}
+where $m(x) = \frac{1}{c^2(x)}$ is the squared slowness with $c(x)$ the spatially dependent speed of sound, 
+$u(t, x)$ is the wavefield, $\Laplace u(t, x)$ denotes the laplacian of the wavefield and $q(t, x)$ is the source.
+Some of the kernels used in the performance measurements corresponded to a more physically representative
+but complex version of this equation called Tilted Transverse Isotropy (TTI) \cite{zhang2011stable}. We leave the 
+equation itself out of this paper for brevity.
 
-test case setup.
-requirements: should be realistic, numerically stable (to get
-realistic compression), big domain,
-large number of timesteps
-Can we use a SEAM model here?
+The values of $m(x)$ were derived from the SEAM Overthrust (TODO Cite SEAM) model which was over a grid
+of $287 \times 881 \times 881$ points, including an absorbing layer of 40 points on each side. The grid spacing
+was TODO in space and TODO in time. This was run for TODO timesteps, and the final field stored as
+representative of a typical snapshot. Figure TODO shows this stored field that was used for the compression tests.
+
+To implement Revolve with Devito, we used pyRevolve (TODO CITE 2 papers) which is a python wrapper over the
+original Revolve utility (TODO CITE Revolve). The performance model that will be developed in section TODO of
+this paper assumes that the implementation is similar to pyRevolve, which stores a checkpoint by copying a portion
+of the operator's working memory to the checkpointing memory and similarly loads a checkpoint by copying from the
+checkpointing memory to the operator's working memory. Although a perfect implementation of checkpointing may
+be able to avoid these copies, the overhead attached to these copies can be ignored for an operator that is
+sufficiently computationally intensive. However, we include the overheads in the model to verify this assumption.
+
+The experiments were run on two different systems, TODO description of Richter, henceforth called Richter and a
+TODO description of skylake, henceforth called Skylake.
 
@@
 \subsection{Lossless}
-ZLib
+For our initial experiments, we used the python package \emph{blosc} (TODO CITE). It includes implementations for
+6 different lossless compression algorithms, namely ZLIB, ZSTD, BLOSC, LZ4, LZ4HC and TODO (TODO cite all of them).
+We did a parameter sweep over all available parameters to find the best-performing combinations. 
+\subsection{Lossy}
+Next we tried lossy compression using ZFP \cite{ratanaworabhan2006fast} (TODO verify citation).
+We tried all three modes of ZFP, namely fixed-tolerance, fixed-precision and fixed-rate modes. Figure TODO shows the compression ratios
+we saw for a variety of compression settings for the three modes. As expected, the highest compression efficiency was
+observed in the fixed-tolerance mode albeit with highly unpredictable compression ratios. Figure TODO shows a sample
+wavefield that was recovered after compression and decompression using TODO mode. Since, in our experiments, the time
+taken to compress a checkpoint was in the same order of magnitude as the time taken to compute a timestep, we build a
+performance model in the next section to evaluate the utility of compression in different scenarios of problem setups and test
+platforms.
 
-variations in compression ratios
-\subsection{Lossy}
-ZFP
-@inproceedings{lossless,
-    Acmid = {1126035},
-    Address = {Washington, DC, USA},
-    Author = {Ratanaworabhan, Paruj and Ke, Jian and Burtscher, Martin},
-    Booktitle = {Proceedings of the Data Compression Conference},
-    Date-Added = {2015-05-03 20:40:52 +0000},
-    Date-Modified = {2018-01-31 10:20:09 +0000},
-    Isbn = {0-7695-2545-8},
-    Numpages = {10},
-    Pages = {133--142},
-    Publisher = {IEEE Computer Society},
-    Series = {DCC '06},
-    Title = {Fast Lossless Compression of Scientific Floating-Point Data},
-    Year = {2006}}
-
-@conference{gpu-compression,
-    Author = {O'Neil, Molly A. and Burtscher, Martin},
-    Booktitle = {Proceedings of the Fourth Workshop on General Purpose Processing on Graphics Processing Units},
-    Date-Added = {2015-05-03 22:40:52 +0000},
-    Date-Modified = {2015-05-03 22:42:57 +0000},
-    Keywords = {GPGPU, floating-point data, lossless data compression, real-time compression},
-    Number = {10.1145/1964179.1964189},
-    Publisher = {ACM},
-    Title = {Floating-point Data Compression at 75 Gb/s on a {GPU}},
-    Year = {2011},
-}
-@inbook{Kaklamanis:2012aa,
-    Author = {Kaklamanis, Christos and Papatheodorou, Theodore and Spirakis, PaulG. and Iverson, Jeremy and Kamath, Chandrika and Karypis, George},
-    Booktitle = {Euro-Par 2012 Parallel Processing},
-    Chapter = {Fast and Effective Lossy Compression Algorithms for Scientific Datasets},
-    Da = {2012/01/01},
-    Date-Added = {2015-05-03 22:46:47 +0000},
-    Date-Modified = {2018-01-31 10:20:09 +0000},
-    Isbn = {978-3-642-32819-0},
-    La = {English},
-    Pages = {843-856},
-    Publisher = {Springer Berlin Heidelberg},
-    Se = {83},
-    Title = {Lecture Notes in Computer Science},
-    Title1 = {Fast and Effective Lossy Compression Algorithms for Scientific Datasets},
-    Ty = {CHAP},
-    Volume = {7484},
-    Year = {2012}}
-
-p-adaptive compression from Boehm et al.
-
-\section{Compression in adjoint-based optimization}
+\section{Performance model for a combination of Revolve and compression}
 We work with the assumption that the computation of a single forward time step takes the same wall time
@@
 \end{equation}
-However, this would require $\mathbf{S} \times \mathbf{N}$ memory. On a system where enough memory
-is available, this would obviously be the optimal strategy to compute the adjoint solution.
+However, this would require $\mathbf{S} \times \mathbf{N}$ memory. On a hardware platform where enough memory
+is available, this would be the optimal strategy to compute the adjoint solution.
 
@@
 an optimal strategy to solve for the adjoint field by storing a subset of the $\mathbf{N}$ total checkpoints
-and recompute the remaining ones. The overhead introduced by this method can be broken down into
+and recomputing the remaining ones. The overhead introduced by this method can be broken down into
 the recomputation overhead $\mathbf{O}_R$ and the storage overhead $\mathbf{O}_S$. The recomputation
@@
 and region 3 where compression would enable the computation to be done without revolve. 
-\section{Performance model}
-We can predict the performance
 \section{Results}
-\subsection{Error analysis}
-Discuss the error incurred due to lossy compression. This could be the
-angular deviation in the gradient from Boehm et al. as a factor of the
-precision parameter, for about 3 different compression schemes  What if a minor
-deviation is at a point where you're pointing directly at a
-constraint? or a corner point between two constraints? The line search
-may mistakenly terminate early. This should be used to derive a bound
-for compression ratio based on desired precision.
-
-\subsection{Compression vs recomputation}
-
-For this paper, we can probably focus on only zfp using 1-3 different precision settings
-(nearly lossless, very lossy, something reasonable).
-
-- Time to solution for different compression ratios, for multiple
-compression algorithms. at least one lossless compression included
-here.
-
-%at least 3 compression algorithms compared
-
-%vary compression ratio within bound established in previous section
 
@@
 
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
+TODO (Acknowledgments for Navjot)
+TODO (Acknowledgments for Jan)
+TODO (Acknowledgments for Mathias)
 
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
-
-\begin{thebibliography}{00}
-\bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-\bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-\end{thebibliography}
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
+\bibliographystyle{plain}
+\bibliography{compression}
 

### 2018-10-02 paper/paper.tex
new abstract

@@
 \begin{abstract}
-Full-waveform inversion is an adjoint-based optimization problem in
-seismic imaging that processes terabytes of data, often beyond the
-memory capacity of most modern computer systems. A common technique to
-carry out the adjoint computation in an environment with
-less-than-optimal memory is checkpoint-recomputation. In this
-technique, a subset of the forward-computation states are stored in
-memory checkpoints and the others are recomputed from the stored
-states. The compression of checkpoints can increase the number of
-checkpoints available, in a given amount of memory, which, in turn,
-can reduce the recompute factor, reducing the time-to-solution of the
-adjoint computation. This compression, often comes at an additional
-computational cost, and an associated loss of precision that must be
-accounted for when choosing between compression schemes. In this
-paper, we study the effect of compression on checkpoint-recomputation
-methods in memory-constrained scenarios for a realistic FWI workload.
+Full-waveform inversion is an adjoint-based optimization problem in seismic
+imaging that processes terabytes of data and regularly exceeds the memory
+capacity of available computers. A popular strategy to reduce the memory
+requirement is data compression, which can reduce the memory
+footprint by a certain factor. An alternative is
+checkpointing, where data is stored at selected points in
+time, and values at other times are recomputed as needed from the last stored state.
+This allows arbitrarily large adjoint computations with limited memory, at the cost
+of additional recomputations.
+
+In this paper we combine compression and checkpointing for the first
+time to compute a realistic full-waveform inversion. The combination of
+checkpointing and compression allows
+larger adjoint computations compared to using only compression, and
+reduces the recomputation overhead significantly compared to using only checkpointing.
+
 \end{abstract}

### 2018-10-02 paper/paper.tex
new keywords
Add images

--- 2018-10-02 paper/paper.tex
@@
 \usepackage{cuted}
+\usepackage{tikz}
+\usepackage{pgfplots}
+
+\usetikzlibrary{calc,trees,positioning,arrows,chains,shapes.geometric,%
+    decorations.pathreplacing,decorations.pathmorphing,shapes,%
+    matrix,shapes.symbols}
+
+\tikzstyle{line} = [draw, -, thick]
+\tikzstyle{nodraw} = [draw, fill, circle, minimum width=0pt, inner sep=0pt]
+\tikzstyle{box} = [line, rectangle, rounded corners, text centered]
+
+\tikzset{
+>=stealth',
+  punktchain/.style={
+    rectangle, 
+    rounded corners, 
+    draw=black, very thick,
+    text width=2em, 
+    minimum height=3em, 
+    text centered, 
+    on chain},
+  line/.style={draw, thick, <-},
+  element/.style={
+    tape,
+    top color=white,
+    bottom color=blue!50!black!60!,
+    minimum width=1em,
+    draw=blue!40!black!90, very thick,
+    text width=2em, 
+    minimum height=3.5em, 
+    text centered, 
+    on chain},
+  every join/.style={->, thick,shorten >=1pt},
+  decoration={brace},
+  tuborg/.style={decorate},
+  tubnode/.style={midway, right=2pt},
+}
+
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 simulation is run ``forward" in simulation time, which produces data that is to be used by a
-subsequent adjoint pass which is run backward in simulation time TODO CITE. Figure TODO
+subsequent adjoint pass which is run backward in simulation time TODO CITE. Figure \ref{fig:dataflow}
 shows the typical data access pattern. This class consists of many different problems of interest in
@@
 solution to the optimization problem gives the most likely structure of the earth's subsurface. Figure
-TODO shows the physical setup of the experiment that produces the data that this problem operates
-on. The data access pattern from Figure TODO applies here as well, since every step during the
+\ref{fig:offshore_survey} shows the physical setup of the experiment that produces the data that this problem operates
+on. The data access pattern from Figure \ref{fig:dataflow} applies here as well, since every step during the
 adjoint propagation requires data from the corresponding step from the forward propagation. In
@@
 the analysis can be applied to any problem with a similar data-access pattern.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.8\linewidth]{images/survey-ship-diagram.png}
+\end{center}
+\caption{Illustration of an offshore seismic survey that collects the input data for FWI}
+\label{fig:offshore_survey}
+\end{figure}
 
@@
 technique to various scenarios (TODO cite other checkpointing papers).
+
+\begin{figure*}
+\begin{center}
+\input{dataflow}
+\end{center}
+\caption{The dataflow pattern that is typical of adjoint-based optimization problems}
+\label{fig:dataflow}
+\end{figure*}
 
@@
 was TODO in space and TODO in time. This was run for TODO timesteps, and the final field stored as
-representative of a typical snapshot. Figure TODO shows this stored field that was used for the compression tests.
+representative of a typical snapshot. Figure \ref{fig:uncompressed} shows this stored field that was used for the compression tests.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.8\linewidth]{images/uncompressed.png}
+\end{center}
+\caption{Cross-section of the wavefield used as a reference sample for compression and decompression}
+\label{fig:uncompressed}
+\end{figure}
 
@@
 \subsection{Lossy}
-Next we tried lossy compression using ZFP \cite{ratanaworabhan2006fast} (TODO verify citation).
+Next we tried lossy compression using ZFP \cite{ratanaworabhan2006fast} (TODO verify citation). To use ZFP from python, 
+we first wrote a python wrapper over the reference implementation of ZFP, which is in C. We plan to release this wrapper
+in the open source community shortly.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.8\linewidth]{images/decompressed-t-8.png}
+\end{center}
+\caption{Cross-section of the wavefield recovered after compression and decompression using the fixed-tolerance mode}
+\label{fig:decompressed}
+\end{figure}
+
 We tried all three modes of ZFP, namely fixed-tolerance, fixed-precision and fixed-rate modes. Figure TODO shows the compression ratios
 we saw for a variety of compression settings for the three modes. As expected, the highest compression efficiency was
-observed in the fixed-tolerance mode albeit with highly unpredictable compression ratios. Figure TODO shows a sample
-wavefield that was recovered after compression and decompression using TODO mode. Since, in our experiments, the time
+observed in the fixed-tolerance mode albeit with highly unpredictable compression ratios. Figure \ref{fig:decompressed} shows a sample
+wavefield that was recovered after compression and decompression using fixed-tolerance mode. Since, in our experiments, the time
 taken to compress a checkpoint was in the same order of magnitude as the time taken to compute a timestep, we build a

### 2018-10-02 performance.py

@@ -1,5 +1,6 @@
 from pyrevolve import Operator, Checkpoint, Revolver
 
+from itertools import product
 import math
 import numpy as np
 import matplotlib.pyplot as plt
@@ -195,9 +196,6 @@ def varying_peak_memory(nt, size_ts, compute_ts, bw, c_factor, c_time, d_time,
     min_time, overall_peak_mem = p.naive_strategy()
     mems = linspace(2*size_ts, overall_peak_mem/c_factor, 200)
     speedups = [p.compression_speedup(x, c_factor, c_time, d_time) for x in mems]
-    print(speedups)
-    print("***")
-    print(mems)
     theoretical_speedup = [p.compression_speedup(x, c_factor, c_time, theoretical_d_time) for x in mems]
     fixed = {'Timesteps': nt, 'Size of checkpoint (MB)': size_ts,
              'Time for compute step (s)': compute_ts, 'Bandwidth (MB/s)': bw,
@@ -298,3 +296,24 @@ varying_nt(cp_size, compute_ts, #time to compute one timestep (s)
 #Where is the tipping point?
 
 # Gives a speedup for everything that is costly and for cheap stuff, when memory is very constrained
+
+
+def vary_all(size_ts, bw, f, c, d):
+    nts = linspace(2000, 20000, 200)
+    nts = [math.floor(x) for x in nts]
+    computes = linspace(0, 100, 200)
+
+    results = np.zeros((200, 200, 200))
+
+    for i, nt in enumerate(nts):
+        for j, compute in enumerate(computes):
+            p = Problem(nt, size_ts, compute, bw)
+            min_time, overall_peak_mem = p.naive_strategy()
+            mems = linspace(2*size_ts, overall_peak_mem/c_factor, 200)
+            for k, mem in enumerate(mems):
+                results[i, j, k] = p.compression_speedup(mem, f, c, d)
+
+    np.save('results', results)
+
+vary_all(cp_size, bandwidth, c_factor, c_time, d_time)
+    

### 2018-10-02 plotter.py

@@ -14,7 +14,7 @@ def read_csv_file(filename):
             if i == 0:
                 headings = row
             else:
-                data.append(row)
+                data.append([float(x) for x in row])
     data = list(zip(*data))
     return data, headings
         
@@ -25,6 +25,7 @@ def plot(x, y, filename, title, xlabel, ylabel):
     plt.ylabel(ylabel)
     plt.title(title)
     plt.xticks(rotation=90)
+    plt.xscale('log')
     plt.savefig(filename, bbox_inches='tight')
 
 if __name__ == "__main__":

### 2018-10-02 paper/paper.tex

@@
 \begin{IEEEkeywords}
-component, formatting, style, styling, insert
+Checkpointing, compression, adjoints, inversion
 \end{IEEEkeywords}

### 2018-10-02 paper/paper.tex
revised abstract again

@@
 Full-waveform inversion is an adjoint-based optimization problem in seismic
-imaging that processes terabytes of data and regularly exceeds the memory
-capacity of available computers. A popular strategy to reduce the memory
-requirement is data compression, which can reduce the memory
-footprint by a certain factor. An alternative is
-checkpointing, where data is stored at selected points in
-time, and values at other times are recomputed as needed from the last stored state.
-This allows arbitrarily large adjoint computations with limited memory, at the cost
-of additional recomputations.
+imaging that processes terabytes of data, regularly exceeding the memory
+capacity of available computers. Data compression is an effective strategy to
+reduce this memory requirement by a certain factor, particularly if some loss in
+accuracy is acceptable. A popular alternative is checkpointing, where data is
+stored at selected points in time, and values at other times are recomputed as
+needed from the last stored state.  This allows arbitrarily large adjoint
+computations with limited memory, at the cost of additional recomputations.
 

### 2018-10-02 paper/paper.tex
Change equation to float

--- 2018-10-02 paper/paper.tex
@@
 the recomputation overhead $\mathbf{O}_R$ and the storage overhead $\mathbf{O}_S$. The recomputation
-overhead is largely fixed and given by the following relationship (TODO cite revolve):
-\begin{strip}
+overhead is simply the amount of time spent in recomputation, given by:
+\begin{equation}
+\mathbf{O}_R(N, M) = p(N, M) \cdot \mathbf{C}
+\end{equation}
+where $p(N, M)$ is the minimum number of recomputed steps from TODO cite revolve, reproduced
+here in equation \ref{eqn:recompute}.
+\begin{figure*}
 \begin{equation}
@@
 \end{equation}
-\end{strip}
-where M is the number of checkpoints that can be stored in memory. Note that for $M >=N$, $\mathbf{O}_R$
+\end{figure*}
+In equation \ref{eqn:recompute}, M is the number of checkpoints that can be stored in memory. Note that for $M >=N$, $\mathbf{O}_R$
 would be zero. For $M < N$, $\mathbf{O}_R$ goes up rather quickly as M is reduced in relation to N.
@@
 \begin{equation}
-\mathbf{O}_SR = \mathbf{N}_W \times \frac{\mathbf{S}}{\mathbf{B}} + \mathbf{N}_R \times \frac{\mathbf{S}}{\mathbf{B}}
+\mathbf{O}_SR = \mathbf{N}_W \cdot \frac{\mathbf{S}}{\mathbf{B}} + \mathbf{N}_R \times \frac{\mathbf{S}}{\mathbf{B}}
 \label{eqn:storage}

### 2018-10-02 paper/paper.tex

--- 2018-10-02 paper/paper.tex
@@
 reduces the recomputation overhead significantly compared to using only checkpointing.
-
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Checkpointing, compression, adjoints, inversion
+component, formatting, style, styling, insert
 \end{IEEEkeywords}

### 2018-10-02 plot_numpy_hdf5.py
Change the order of x, y, z

@@ -4,16 +4,23 @@ import matplotlib.pyplot as plt
 import sys
 import h5py
 import numpy as np
-
+from simple import from_hdf5
 
 input_file = sys.argv[1]
 output_file = sys.argv[2]
 
 with h5py.File(input_file) as ifile:
     data = ifile['data'][()]
+    from IPython import embed
+    embed()
+
+velocity_model='overthrust_3D_initial_model.h5'
+
+model = from_hdf5(velocity_model, space_order=2, nbpml=20, datakey='m0', dtype=np.float32)
 
 vmax = np.max(data)
 
 plt.imshow(data[143], vmin=-vmax, vmax=vmax, cmap="seismic")
+plt.imshow(model.vp[143], alpha=0.5)
 
 plt.savefig(output_file)

### 2018-10-02 simple.py

@@ -33,6 +33,7 @@ def from_hdf5(filename, **kwargs):
     dtype = kwargs.pop('dtype', None)
     data_m = f[datakey][()]
     data_vp = np.sqrt(1/data_m).astype(dtype)
+    data_vp = np.flip(data_vp, (1, 2, 0))
     return Model(space_order=space_order, vp=data_vp, origin=origin, shape=shape,
                      dtype=dtype, spacing=spacing, nbpml=nbpml)
 
@@ -82,6 +83,7 @@ def overthrust_setup_tti(filename, kernel='OT2', space_order=2, nbpml=40, **kwar
     src.coordinates.data[0, :] = np.array(model.domain_size) * .5
     if len(shape) > 1:
         src.coordinates.data[0, -1] = model.origin[-1] + 2 * spacing[-1]
+
     # Define receiver geometry (spread across x, just below surface)
     rec = Receiver(name='rec', grid=model.grid, time_range=time_range, npoint=nrec)
     rec.coordinates.data[:, 0] = np.linspace(0., model.domain_size[0], num=nrec)

### 2018-10-02 simple.py
Replace flip with transpose

@@ -33,7 +33,7 @@ def from_hdf5(filename, **kwargs):
     dtype = kwargs.pop('dtype', None)
     data_m = f[datakey][()]
     data_vp = np.sqrt(1/data_m).astype(dtype)
-    data_vp = np.flip(data_vp, (1, 2, 0))
+    data_vp = np.transpose(data_vp, (1, 2, 0))
     return Model(space_order=space_order, vp=data_vp, origin=origin, shape=shape,
                      dtype=dtype, spacing=spacing, nbpml=nbpml)
 

### 2018-10-02 simple.py
Still fixing issues with model loading

@@ -28,12 +28,12 @@ def from_hdf5(filename, **kwargs):
     datakey = kwargs.pop('datakey', None)
     if datakey is None:
         raise ValueError("datakey must be known - what is the name of the data in the file?")
-    shape = f[datakey].shape
     space_order=kwargs.pop('space_order', None)
     dtype = kwargs.pop('dtype', None)
     data_m = f[datakey][()]
     data_vp = np.sqrt(1/data_m).astype(dtype)
     data_vp = np.transpose(data_vp, (1, 2, 0))
+    shape = data_vp.shape
     return Model(space_order=space_order, vp=data_vp, origin=origin, shape=shape,
                      dtype=dtype, spacing=spacing, nbpml=nbpml)
 

### 2018-10-02 plot_numpy_hdf5.py
Change to use OT4 acoustic

@@ -11,8 +11,6 @@ output_file = sys.argv[2]
 
 with h5py.File(input_file) as ifile:
     data = ifile['data'][()]
-    from IPython import embed
-    embed()
 
 velocity_model='overthrust_3D_initial_model.h5'
 
@@ -20,7 +18,8 @@ model = from_hdf5(velocity_model, space_order=2, nbpml=20, datakey='m0', dtype=n
 
 vmax = np.max(data)
 
-plt.imshow(data[143], vmin=-vmax, vmax=vmax, cmap="seismic")
-plt.imshow(model.vp[143], alpha=0.5)
-
+slice_loc = 400
+plt.imshow(np.transpose(data[slice_loc]), vmin=-vmax, vmax=vmax, cmap="seismic")
+plt.imshow(np.transpose(model.vp[slice_loc]), alpha=0.5)
+plt.colorbar()
 plt.savefig(output_file)

### 2018-10-02 simple.py

@@ -94,11 +94,11 @@ def overthrust_setup_tti(filename, kernel='OT2', space_order=2, nbpml=40, **kwar
     return AnisotropicWaveSolver(model, source=src, receiver=rec,
                                  space_order=space_order, **kwargs)
 
-def run(space_order=4, kernel='OT2', nbpml=40, filename='', **kwargs):
+def run(space_order=4, kernel='OT4', nbpml=40, filename='', **kwargs):
 
-    #solver = overthrust_setup(filename=filename, nbpml=nbpml, space_order=space_order, kernel=kernel, **kwargs)
+    solver = overthrust_setup(filename=filename, nbpml=nbpml, space_order=space_order, kernel=kernel, **kwargs)
 
-    solver = overthrust_setup_tti(filename=filename, nbpml=nbpml, space_order=space_order, kernel=kernel, **kwargs)
+    #solver = overthrust_setup_tti(filename=filename, nbpml=nbpml, space_order=space_order, kernel=kernel, **kwargs)
 
     return_values = solver.forward(save=False)
     rec = return_values[0]

### 2018-10-02 simple.py
Numerical stability

@@ -50,7 +50,7 @@ def overthrust_setup(filename, kernel='OT2', space_order=2, nbpml=40, **kwargs):
     time_range = TimeAxis(start=t0, stop=tn, step=dt)
 
     # Define source geometry (center of domain, just below surface)
-    src = RickerSource(name='src', grid=model.grid, f0=0.01, time_range=time_range)
+    src = RickerSource(name='src', grid=model.grid, f0=0.005, time_range=time_range)
     src.coordinates.data[0, :] = np.array(model.domain_size) * .5
     if len(shape) > 1:
         src.coordinates.data[0, -1] = model.origin[-1] + 2 * spacing[-1]

### 2018-10-02 plot_numpy_hdf5.py
Spacing bug

@@ -18,7 +18,7 @@ model = from_hdf5(velocity_model, space_order=2, nbpml=20, datakey='m0', dtype=n
 
 vmax = np.max(data)
 
-slice_loc = 400
+slice_loc = 440
 plt.imshow(np.transpose(data[slice_loc]), vmin=-vmax, vmax=vmax, cmap="seismic")
 plt.imshow(np.transpose(model.vp[slice_loc]), alpha=0.5)
 plt.colorbar()

### 2018-10-02 simple.py

@@ -40,7 +40,7 @@ def from_hdf5(filename, **kwargs):
 def overthrust_setup(filename, kernel='OT2', space_order=2, nbpml=40, **kwargs):
     model = from_hdf5(filename, space_order=space_order, nbpml=nbpml, datakey='m0', dtype=np.float32)
     shape = model.vp.shape
-    spacing = model.shape
+    spacing = model.spacing
     nrec = shape[0]
     tn = 4000
 
@@ -57,6 +57,8 @@ def overthrust_setup(filename, kernel='OT2', space_order=2, nbpml=40, **kwargs):
     # Define receiver geometry (spread across x, just below surface)
     rec = Receiver(name='rec', grid=model.grid, time_range=time_range, npoint=nrec)
     rec.coordinates.data[:, 0] = np.linspace(0., model.domain_size[0], num=nrec)
+    from IPython import embed
+    embed()
     if len(shape) > 1:
         rec.coordinates.data[:, 1:] = src.coordinates.data[0, 1:]
 

### 2018-10-02 paper/paper.tex
Change the order of x, y, z
fix minor detail

--- 2018-10-02 paper/paper.tex
@@
 forward pass is stored (and the rest discarded). The discarded data is recomputed when required
-by rerunning the forward pass. Starting from Revolve (TODO cite), which was a provably optimal
+by rerunning the forward pass. Starting from Revolve (TODO cite), which is a provably optimal
 scheduling algorithm to implement this strategy, various studies have used and extended this

### 2018-10-02 paper/paper.tex

--- 2018-10-02 paper/paper.tex
@@
 \begin{equation}
-\mathbf{O}_R(N, M) = \begin{cases}
+p(N, M) = \begin{cases}
       N(N-1) /2, & \text{if}\ M=1 \\

### 2018-10-03 plot_numpy_hdf5.py
Add plots for wavefields

@@ -15,11 +15,19 @@ with h5py.File(input_file) as ifile:
 velocity_model='overthrust_3D_initial_model.h5'
 
 model = from_hdf5(velocity_model, space_order=2, nbpml=20, datakey='m0', dtype=np.float32)
-
+shape = model.vp.shape
 vmax = np.max(data)
 
 slice_loc = 440
-plt.imshow(np.transpose(data[slice_loc]), vmin=-vmax, vmax=vmax, cmap="seismic")
-plt.imshow(np.transpose(model.vp[slice_loc]), alpha=0.5)
-plt.colorbar()
-plt.savefig(output_file)
+im = plt.imshow(np.transpose(data[slice_loc]), vmax=.01, vmin=-.01, cmap="seismic",
+           extent = [0, 20, 0.001*(shape[-1]-1)*25, 0])
+
+plt.xlabel("X (km)")
+plt.ylabel("Depth (km)")
+cb = plt.colorbar(shrink=.3, pad=.01, aspect=10)
+for l in cb.ax.yaxis.get_ticklabels():
+    l.set_fontsize(12)
+
+cb.set_label('Pressure')
+
+plt.savefig(output_file, bbox_inches='tight')

### 2018-10-03 paper/paper.tex
Add plots for wavefields
added some compression blurp

@@
 
-introduction on compression techniques - TODO Jan
-Compression is an obvious addition to the tools applied to the specific data access pattern
+Another strategy to reduce the memory footprint of scientific applications is
+data compression. Well-known general purpose data compression packages such as
+gzip TODO cite or 7zip TODO cite, and compression for video and image
+data such as JPEG-2000 TODO cite have been used for scientific data. More
+recently, special purpose compression algorithms for scientific data have been
+developed, such as ZFP or SZ TODO cite.
 
-One possible way of using compression to speed up full-waveform inversion or any adjoint-based
+Lossless algorithms guarantee that the exact original data can be
+recovered during decompression, whereas lossy algorithms introduce an error, but
+often guarantee that the error does not exceed certain absolute or relative
+error metrics. Typically, lossy compression is more effective in reducing the
+data size, or in other words, offer a greater compression ratio. Most popular
+compression packages offer various settings that allow a tradeoff between
+compression ratio, accuracy, and compression and decompression time. In this
+work we use the ZFP package to perform lossy compresssion.
+
+It is worth noting that another popular data reduction strategy is to typecast
+values into a lower precision format, for example, from double precision to
+single precision TODO cite. This can be seen as a computationally cheap lossy
+compression algorithm with a compression ratio of $2$.
+
+Perhaps counterintuitively, compression can not only reduce the memory
+footprint, but also speed up an application. Previous work has observed that the
+compression and decompression time can be less than the time saved from the
+reduction in data that needs to be communicated across MPI nodes or between a
+GPU and a host computer TODO cite.
+
+Another possible way of using compression to speed up full-waveform inversion or any adjoint-based
 optimization is by simply compressing the field data before storing it and decompressing it when it

### 2018-10-03 paper/paper.tex
Expanding related work

--- 2018-10-02 paper/paper.tex
@@
 forward pass is stored (and the rest discarded). The discarded data is recomputed when required
-by rerunning the forward pass. Starting from Revolve (TODO cite), which is a provably optimal
+by rerunning the forward pass. Starting from Revolve (TODO cite), which was a provably optimal
 scheduling algorithm to implement this strategy, various studies have used and extended this
@@
 \section{Related Work}
-\cite{cyr2015towards}
+\cite{cyr2015towards} studied compression for adjoint-based optimisation in the context of the Navier-Stokes
+equation, focussing on the scenario where the entire forward field could be stored in memory after
+compression. They looked at Revolve as an alternative to compression.
 
-\cite{dalmau2014lossy}
+\cite{dalmau2014lossy} also looked at compression as an alternative to Revolve, although in the same context
+as this paper - seismic imaging. Their implementation was, however, storing the compressed data to disk so a
+lot of their benefits came from more efficient utilization of the low disk bandwidth. Their focus was on developing
+and fine-tuning a domain-specific compression algorithm for seismic imaging which was based on discrete
+cosine transforms (DCT).
 
@@
 TODO cite SZ (Frank Cappello)
+TODO Jan check whether SZ supports 4D - it does support up to 4D. 
 \section{Test case using Devito and pyRevolve}
@@
 
+TODO expand to include FWI objective function and explain that the gradient is computed using the adjoint-state method
+
 The values of $m(x)$ were derived from the SEAM Overthrust (TODO Cite SEAM) model which was over a grid
 of $287 \times 881 \times 881$ points, including an absorbing layer of 40 points on each side. The grid spacing
-was TODO in space and TODO in time. This was run for TODO timesteps, and the final field stored as
+was 25m in space and TODO in time. This was run for 2500 timesteps, and the final field stored as
 representative of a typical snapshot. Figure \ref{fig:uncompressed} shows this stored field that was used for the compression tests.
@@
 \begin{center}
-\includegraphics[width=0.8\linewidth]{images/uncompressed.png}
+\includegraphics[width=0.8\linewidth]{images/uncompressed.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.8\linewidth]{images/decompressed-t-8.png}
+\includegraphics[width=0.8\linewidth]{images/errors.pdf}
 \end{center}
-\caption{Cross-section of the wavefield recovered after compression and decompression using the fixed-tolerance mode}
-\label{fig:decompressed}
+\caption{Cross-section of the field that shows errors introduced during compression and decompression using the fixed-tolerance mode}
+\label{fig:decompressed_error}
 \end{figure}
@@
 
+More than 3D fields (use time correlation)
+
+Parallel decompression
+
 Disk checkpointing

### 2018-10-03 paper/paper.tex
Addressing tODOs

@@
 simulation is run ``forward" in simulation time, which produces data that is to be used by a
-subsequent adjoint pass which is run backward in simulation time TODO CITE. Figure \ref{fig:dataflow}
+subsequent adjoint pass which is run backward in simulation time. Figure \ref{fig:dataflow}
 shows the typical data access pattern. This class consists of many different problems of interest in
@@
 \ref{fig:offshore_survey} shows the physical setup of the experiment that produces the data that this problem operates
-on. The data access pattern from Figure \ref{fig:dataflow} applies here as well, since every step during the
+on. \cite{plessix2006review} The data access pattern from Figure \ref{fig:dataflow} applies here as well, since every step during the
 adjoint propagation requires data from the corresponding step from the forward propagation. In
@@
 Various strategies are routinely employed to reduce this memory requirement, for example 
-Domain decomposition (TODO cite) where the domain of the simulation is distributed over
+Domain decomposition where the domain of the simulation is distributed over
 multiple nodes so as to fit within the memory available on each node. This introduces the overhead
-of network communication and might not efficiently use the computational resources on each node.
+of network communication and might not efficiently use the computational resources on each node.\cite{virieux2009seismic}
 Another common strategy, which is specific to the domain of FWI, is where only the boundaries of
 the domain are stored at each timestep and the rest of the wavefield is reconstructed from the
-boundaries when required (TODO CITE). Another commonly used strategy that is applied across
+boundaries when required \cite{clapp2009reverse}\cite{yang2014rtm}. Another commonly used strategy that is applied across
 a much wider domain of problems is checkpointing. In this strategy, a subset of the data from the
 forward pass is stored (and the rest discarded). The discarded data is recomputed when required
-by rerunning the forward pass. Starting from Revolve (TODO cite), which is a provably optimal
+by rerunning the forward pass. Starting from Revolve \cite{griewank2000algorithm}, which is a provably optimal
 scheduling algorithm to implement this strategy, various studies have used and extended this
-technique to various scenarios (TODO cite other checkpointing papers).
+technique to various scenarios \cite{wang2009minimal} \cite{aupy2016optimal} \cite{schanen2016asynchronous}.
 
@@
 Another strategy to reduce the memory footprint of scientific applications is
-data compression. Well-known general purpose data compression packages such as
-gzip TODO cite or 7zip TODO cite, and compression for video and image
-data such as JPEG-2000 TODO cite have been used for scientific data. More
-recently, special purpose compression algorithms for scientific data have been
-developed, such as ZFP or SZ TODO cite.
+data compression. Well-known general purpose data compression algorithms like
+Zlib (which is a part of gzip) \cite{deutsch1996zlib}, and compression for video and image
+data such as JPEG-2000 \cite{skodras2001jpeg}have been used for scientific data. More
+recently, special purpose compression algorithms for floating-point scientific data have been
+developed, such as ZFP or SZ \cite{lindstrom2014fixed} \cite{di2018efficient} 
+\cite{Kaklamanis:2012aa}.
 
@@
 reduction in data that needs to be communicated across MPI nodes or between a
-GPU and a host computer TODO cite.
+GPU and a host computer. There has been some work on streaming compression
+of floating-point data for this specific intent. \cite{gpu-compression}
 
-Another possible way of using compression to speed up full-waveform inversion or any adjoint-based
+One possible way of using compression to speed up full-waveform inversion or any adjoint-based
 optimization is by simply compressing the field data before storing it and decompressing it when it
@@
 compression, the use of compression effectively replaces the use of revolve-based checkpointing.
-This scenario has been discussed in different studies before - some of which we cover in the next
-section. This paper focusses on a broader region, of which this scenario is just one part. In particular,
-we discuss the scenarios where the forward field can not fit in memory even after compression, where
-compression and Revolve would be used in conjunction.
+This scenario has been discussed in different studies before - they look at Revolve as an alternative
+to compression. This approach was seen for a study of adjoint-based methods on the Navier-Stokes
+equation \cite{cyr2015towards}. This was also studied in the context of seismic imaging and FWI
+to develop tailored compression methods for the wavefield data that is typically seen in this context - 
+including a compression method based on DCT \cite{dalmau2014lossy} and one based on adaptive
+coarsening \cite{boehm2016wavefield}. This paper focusses on a broader region, of which this scenario
+is just one part. In particular, we discuss the scenarios where the forward field can not fit in memory even
+after compression, where compression and Revolve would be used in conjunction.
 
@@
 
-\section{Related Work}
-\cite{cyr2015towards} studied compression for adjoint-based optimisation in the context of the Navier-Stokes
-equation, focussing on the scenario where the entire forward field could be stored in memory after
-compression. They looked at Revolve as an alternative to compression.
-
-\cite{dalmau2014lossy} also looked at compression as an alternative to Revolve, although in the same context
-as this paper - seismic imaging. Their implementation was, however, storing the compressed data to disk so a
-lot of their benefits came from more efficient utilization of the low disk bandwidth. Their focus was on developing
-and fine-tuning a domain-specific compression algorithm for seismic imaging which was based on discrete
-cosine transforms (DCT).
-
-\cite{fichtner2009full}
-
-\cite{gpu-compression}
-
-\cite{boehm2016wavefield}
-\cite{gpu-compression}
-\cite{Kaklamanis:2012aa}
-
-TODO cite SZ (Frank Cappello)
-TODO Jan check whether SZ supports 4D - it does support up to 4D. 
 \section{Test case using Devito and pyRevolve}
@@
 m(x)\frac{\partial^2 u(t, x)}{\partial t^2} - \Laplace u(t, x) = q(t, x)
+\label{eqn:wave}
 \end{equation}
@@
 but complex version of this equation called Tilted Transverse Isotropy (TTI) \cite{zhang2011stable}. We leave the 
-equation itself out of this paper for brevity.
+TTI equations out of this paper for brevity.
 
-TODO expand to include FWI objective function and explain that the gradient is computed using the adjoint-state method
+The solution to equation \ref{eqn:wave} form the forward problem. The complete optimization problem involves
+the minimization of the misfit between the simulated signal and the observed signal as follows:
 
-The values of $m(x)$ were derived from the SEAM Overthrust (TODO Cite SEAM) model which was over a grid
+\begin{equation}
+\min_{m} \phi_s(m) = \frac{1}{2} \left\lVert d_{sim} - d_{obs} \right\rVert_2^2
+\end{equation}
+
+This optimization problem is usually solved using gradient-descent methods and the gradient calculation is done
+using the adjoint-state method, which involves the data-flow pattern from figure \ref{fig:dataflow}.
+
+The values of $m(x)$ were derived from the SEAM Overthrust model \cite{aminzadeh1996three} which was over a grid
 of $287 \times 881 \times 881$ points, including an absorbing layer of 40 points on each side. The grid spacing
@@
 
-To implement Revolve with Devito, we used pyRevolve (TODO CITE 2 papers) which is a python wrapper over the
-original Revolve utility (TODO CITE Revolve). The performance model that will be developed in section TODO of
+To implement Revolve with Devito, we used pyRevolve \cite{kukreja2018high} which is a python wrapper over the
+original Revolve utility. The performance model that will be developed in section \ref{sec:performance_model} of
 this paper assumes that the implementation is similar to pyRevolve, which stores a checkpoint by copying a portion
@@
 we saw for a variety of compression settings for the three modes. As expected, the highest compression efficiency was
-observed in the fixed-tolerance mode albeit with highly unpredictable compression ratios. Figure \ref{fig:decompressed} shows a sample
-wavefield that was recovered after compression and decompression using fixed-tolerance mode. Since, in our experiments, the time
+observed in the fixed-tolerance mode albeit with highly unpredictable compression ratios. Figure \ref{fig:decompressed_error} shows the
+spatial distribution of the error introduced after compression and decompression using fixed-tolerance mode. Since, in our experiments, the time
 taken to compress a checkpoint was in the same order of magnitude as the time taken to compute a timestep, we build a
@@
 \section{Performance model for a combination of Revolve and compression}
+\label{sec:performance_model}
 We work with the assumption that the computation of a single forward time step takes the same wall time
@@
 \begin{equation}
-T_N = 2 \times \mathbf{C} \times \mathbf{N}
+T_N = 2 \cdot \mathbf{C} \cdot \mathbf{N}
 \end{equation}
-However, this would require $\mathbf{S} \times \mathbf{N}$ memory. On a hardware platform where enough memory
+However, this would require $\mathbf{S} \cdot \mathbf{N}$ memory. On a hardware platform where enough memory
 is available, this would be the optimal strategy to compute the adjoint solution.
 
-When the memory available on a system is less than $\mathbf{S} \times \mathbf{N}$, Revolve provides
+When the memory available on a system is less than $\mathbf{S} \cdot \mathbf{N}$, Revolve provides
 an optimal strategy to solve for the adjoint field by storing a subset of the $\mathbf{N}$ total checkpoints
@@
 \begin{equation}
-\mathbf{O}_SR = \mathbf{N}_W \cdot \frac{\mathbf{S}}{\mathbf{B}} + \mathbf{N}_R \times \frac{\mathbf{S}}{\mathbf{B}}
+\mathbf{O}_{SR} = \mathbf{N}_W \cdot \frac{\mathbf{S}}{\mathbf{B}} + \mathbf{N}_R \cdot \frac{\mathbf{S}}{\mathbf{B}}
 \label{eqn:storage}
@@
 \begin{equation}
-T_R = 2 \times \mathbf{C} \times \mathbf{N} + \mathbf{O}_R(N, M) + \mathbf{O}_S
+T_R = 2 \cdot \mathbf{C} \cdot \mathbf{N} + \mathbf{O}_R(N, M) + \mathbf{O}_S
 \end{equation}
@@
 \begin{equation}
-\mathbf{O}_SR = \mathbf{N}_W \times (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} + t_c) + \mathbf{N}_R \times (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} + t_d)
+\mathbf{O}_{SR} = \mathbf{N}_W \cdot (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} + t_c) + \mathbf{N}_R \cdot (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} + t_d)
 \end{equation}

### 2018-10-03 paper/paper.tex
revised Introduction

@@
 \section{Introduction}
-The class of adjoint-based optimization problems has a very typical data access pattern where a
-simulation is run ``forward" in simulation time, which produces data that is to be used by a
-subsequent adjoint pass which is run backward in simulation time. Figure \ref{fig:dataflow}
-shows the typical data access pattern. This class consists of many different problems of interest in
-science and engineering. One problem in this class is full-waveform inversion (FWI) where physical
-experiments are carried out by propagating seismic waves through the earth's subsurface and the
-solution to the optimization problem gives the most likely structure of the earth's subsurface. Figure
-\ref{fig:offshore_survey} shows the physical setup of the experiment that produces the data that this problem operates
-on. \cite{plessix2006review} The data access pattern from Figure \ref{fig:dataflow} applies here as well, since every step during the
-adjoint propagation requires data from the corresponding step from the forward propagation. In
-particular for FWI, this data can reach tens of terabytes in size, which is beyond the memory
-available on most hardware. While this paper is written to be focussed on FWI as the target problem, 
-the analysis can be applied to any problem with a similar data-access pattern.
+Adjoint-based optimization problems typically consist of a simulation that is
+run forward in simulation time, producing data that is used in reverse order by
+a subsequent adjoint computation that is run backwards in simulation time.
+Figure~\ref{fig:dataflow} shows the resulting data flow. Many important
+applications in science and engineering have this structure, including
+full-waveform inversion (FWI).
+
+In FWI, the propagation of seismic waves through the earth's subsurface is
+simulated and compared with data from real-world measurements. The model of the
+subsurface is iteratively improved by minimizing the misfit between simulation and
+measurement in an adjoint optimization problem.
+Figure~\ref{fig:offshore_survey} shows the setup of the real-world experiment that
+produces the measurement data~\cite{plessix2006review}. Storing all intermediate
+results as required for the data flow shown in Figure~\ref{fig:dataflow} is not
+feasible, since this would require tens of terabytes of memory, which is not
+commonly available on the machines used to solve these problems. To solve this
+problem, a number of strategies are used that are discussed in the following
+paragraphs. While this paper uses FWI as a test case, most of these ideas,
+including the one presented in this work, are also applicable to other adjoint
+optimization problems.
 
@@
 
-Various strategies are routinely employed to reduce this memory requirement, for example 
-Domain decomposition where the domain of the simulation is distributed over
-multiple nodes so as to fit within the memory available on each node. This introduces the overhead
-of network communication and might not efficiently use the computational resources on each node.\cite{virieux2009seismic}
-Another common strategy, which is specific to the domain of FWI, is where only the boundaries of
-the domain are stored at each timestep and the rest of the wavefield is reconstructed from the
-boundaries when required \cite{clapp2009reverse}\cite{yang2014rtm}. Another commonly used strategy that is applied across
-a much wider domain of problems is checkpointing. In this strategy, a subset of the data from the
-forward pass is stored (and the rest discarded). The discarded data is recomputed when required
-by rerunning the forward pass. Starting from Revolve \cite{griewank2000algorithm}, which is a provably optimal
-scheduling algorithm to implement this strategy, various studies have used and extended this
-technique to various scenarios \cite{wang2009minimal} \cite{aupy2016optimal} \cite{schanen2016asynchronous}.
+Domain decomposition is often used not only to distribute the computational
+workload across more processors, but also to utilise the large amount of memory
+available in distributed systems. While this strategy is very powerful, the
+number of compute nodes and therefore the amount of memory that can be used
+efficiently is limited, for example by communication overheads that start to
+dominate as the domain is split into increasingly small
+pieces~\cite{virieux2009seismic}.
+Another common strategy in FWI is to only store values at the boundaries of the
+domain at each timestep, and reconstruct the rest of the wavefield when
+required~\cite{clapp2009reverse,yang2014rtm}.
+
+Checkpointing is yet another strategy to reduce the memory overhead. Only a
+subset of the program states during the forward pass is stored (and the rest
+discarded).  The discarded data is recomputed when needed by restarting the
+forward pass from the last available stored state. The question which states
+should be stored and which states should be recomputed to minimize the total
+amount of recomputation work has been answerd under certain assumptions with the
+Revolve algorithm~\cite{griewank2000algorithm}.  Other authors have subsequently
+developed extensions to Revolve that are optimal under less restrictive
+assumptions~\cite{wang2009minimal,aupy2016optimal,schanen2016asynchronous}.
 
@@
 
-Another strategy to reduce the memory footprint of scientific applications is
-data compression. Well-known general purpose data compression algorithms like
-Zlib (which is a part of gzip) \cite{deutsch1996zlib}, and compression for video and image
-data such as JPEG-2000 \cite{skodras2001jpeg}have been used for scientific data. More
-recently, special purpose compression algorithms for floating-point scientific data have been
-developed, such as ZFP or SZ \cite{lindstrom2014fixed} \cite{di2018efficient} 
-\cite{Kaklamanis:2012aa}.
+Data compression is also increasingly used to reduce the memory footprint of
+scientific applications. General purpose data compression algorithms like Zlib
+(which is a part of gzip)~\cite{deutsch1996zlib}, and compression algorithms for
+video and image data such as JPEG-2000~\cite{skodras2001jpeg} have been
+presented in previous work. More recently, special purpose compression
+algorithms for floating-point scientific data have been developed, such as ZFP
+or SZ~\cite{Kaklamanis:2012aa,lindstrom2014fixed,di2018efficient}.
 
-Lossless algorithms guarantee that the exact original data can be
-recovered during decompression, whereas lossy algorithms introduce an error, but
-often guarantee that the error does not exceed certain absolute or relative
-error metrics. Typically, lossy compression is more effective in reducing the
-data size, or in other words, offer a greater compression ratio. Most popular
-compression packages offer various settings that allow a tradeoff between
-compression ratio, accuracy, and compression and decompression time. In this
-work we use the ZFP package to perform lossy compresssion.
+Lossless algorithms guarantee that the exact original data can be recovered
+during decompression, whereas lossy algorithms introduce an error, but often
+guarantee that the error does not exceed certain absolute or relative error
+metrics. Typically, lossy compression is more effective in reducing the data
+size. Most popular compression packages offer various settings that allow a
+tradeoff between compression ratio, accuracy, and compression and decompression
+time. In this work we use the ZFP package to perform lossy compresssion.
 
-It is worth noting that another popular data reduction strategy is to typecast
-values into a lower precision format, for example, from double precision to
-single precision TODO cite. This can be seen as a computationally cheap lossy
-compression algorithm with a compression ratio of $2$.
+It is worth noting that another data reduction strategy is to typecast values
+into a lower precision format, for example, from double precision to single
+precision. This can be seen as a computationally cheap lossy compression
+algorithm with a compression ratio of $2$.
 
@@
 reduction in data that needs to be communicated across MPI nodes or between a
-GPU and a host computer. There has been some work on streaming compression
-of floating-point data for this specific intent. \cite{gpu-compression}
+GPU and a host computer~\cite{gpu-compression}.
 
-One possible way of using compression to speed up full-waveform inversion or any adjoint-based
-optimization is by simply compressing the field data before storing it and decompressing it when it
-is required again. If the memory available is sufficient to store the complete forward data after
-compression, the use of compression effectively replaces the use of revolve-based checkpointing.
-This scenario has been discussed in different studies before - they look at Revolve as an alternative
-to compression. This approach was seen for a study of adjoint-based methods on the Navier-Stokes
-equation \cite{cyr2015towards}. This was also studied in the context of seismic imaging and FWI
-to develop tailored compression methods for the wavefield data that is typically seen in this context - 
-including a compression method based on DCT \cite{dalmau2014lossy} and one based on adaptive
-coarsening \cite{boehm2016wavefield}. This paper focusses on a broader region, of which this scenario
-is just one part. In particular, we discuss the scenarios where the forward field can not fit in memory even
-after compression, where compression and Revolve would be used in conjunction.
+Another way of using compression to speed up FWI or other adjoint-based methods
+is to use it \emph{instead} of checkpointing. If the compression ratio is
+sufficient to fit the entire data in memory, checkpointing is no longer
+necessary. Previous work has discussed this situation in the context of
+computational fluid dynamics~\cite{cyr2015towards} and FWI using compression
+method specifically designed for
+wavefields~\cite{dalmau2014lossy,boehm2016wavefield}.
 
-Although most of the compression we discuss here introduces some error which may or may not affect
-the final result of the computation, we try to avoid discussing the error in this paper and focus on the 
-performance impact of compression instead.
+In this paper, we extend the previous studies by \emph{combining} checkpointing
+and compression. An obvious use case would be one where the data does not fit
+in the available memory even after compression. This could happen either for
+very large adjoint problems, or for problems where the required accuracy limits
+the achievable compression ratios.
+
+Compared to using only checkpointing without compression, our combined method
+can often achieve better performance. This is achieved by reducing the size of
+stored program states, thereby allowing more program states to be stored during
+the forward computation. This in turn reduces the amount of recomputations that
+needs to be performed during the adjoint computation. On the other hand, the
+compression and decompression itself takes time.  We thereforen provide a
+comprehensive performance model that predicts whether or not the combined method
+is beneficial. We present various scenarios that use a TODO describe Richter and
+TODO describe JLSE/SKX node system, error tolerances for the lossy ZFP
+compression algorithm ranging from TODO to TODO, and different types of forward
+and adjoint computations that vary in their ratio between compute cost and state
+size.
+
+Our aim is to make this discussion independent of the hardware, application and
+compression algorithm. We therefore do not discuss whether or not the accuracy
+of the decompressed data is sufficient for our application, or whether
+other algorithms might achieve a better accuracy/compression tradeoff.
 

### 2018-10-03 paper/paper.tex
revised Introduction

@@
 In this paper, we extend the previous studies by \emph{combining} checkpointing
-and compression. An obvious use case would be one where the data does not fit
-in the available memory even after compression. This could happen either for
-very large adjoint problems, or for problems where the required accuracy limits
-the achievable compression ratios.
+and compression. This is obviously useful when the data does not fit in the
+available memory even after compression, for example for very large adjoint
+problems, or for problems where the required accuracy limits the achievable
+compression ratios.
 

### 2018-10-03 paper/paper.tex
revised Introduction

@@
 Compared to using only checkpointing without compression, our combined method
-can often achieve better performance. This is achieved by reducing the size of
-stored program states, thereby allowing more program states to be stored during
-the forward computation. This in turn reduces the amount of recomputations that
-needs to be performed during the adjoint computation. On the other hand, the
-compression and decompression itself takes time.  We thereforen provide a
-comprehensive performance model that predicts whether or not the combined method
-is beneficial. We present various scenarios that use a TODO describe Richter and
-TODO describe JLSE/SKX node system, error tolerances for the lossy ZFP
+often improves performance. This is a consequence of the reduced size of stored
+program states, allowing more program states to be stored during the forward
+computation. This in turn reduces the amount of recomputation that needs to be
+performed. On the other hand, the compression and decompression itself takes
+time. We thereforen provide a comprehensive performance model that predicts
+whether the combined method is beneficial, that is, whether the time spent
+compressing and decompressing is less than the time saved by reduced
+recomputations. We present various scenarios that use a TODO describe Richter
+and TODO describe JLSE/SKX node system, error tolerances for the lossy ZFP
 compression algorithm ranging from TODO to TODO, and different types of forward

### 2018-10-03 paper/paper.tex
Allow choosing kernel from command-line

--- 2018-10-03 paper/paper.tex
@@
 \section{Introduction}
-Adjoint-based optimization problems typically consist of a simulation that is
-run forward in simulation time, producing data that is used in reverse order by
-a subsequent adjoint computation that is run backwards in simulation time.
-Figure~\ref{fig:dataflow} shows the resulting data flow. Many important
-applications in science and engineering have this structure, including
-full-waveform inversion (FWI).
-
-In FWI, the propagation of seismic waves through the earth's subsurface is
-simulated and compared with data from real-world measurements. The model of the
-subsurface is iteratively improved by minimizing the misfit between simulation and
-measurement in an adjoint optimization problem.
-Figure~\ref{fig:offshore_survey} shows the setup of the real-world experiment that
-produces the measurement data~\cite{plessix2006review}. Storing all intermediate
-results as required for the data flow shown in Figure~\ref{fig:dataflow} is not
-feasible, since this would require tens of terabytes of memory, which is not
-commonly available on the machines used to solve these problems. To solve this
-problem, a number of strategies are used that are discussed in the following
-paragraphs. While this paper uses FWI as a test case, most of these ideas,
-including the one presented in this work, are also applicable to other adjoint
-optimization problems.
+The class of adjoint-based optimization problems has a very typical data access pattern where a
+simulation is run ``forward" in simulation time, which produces data that is to be used by a
+subsequent adjoint pass which is run backward in simulation time. Figure \ref{fig:dataflow}
+shows the typical data access pattern. This class consists of many different problems of interest in
+science and engineering. One problem in this class is full-waveform inversion (FWI) where physical
+experiments are carried out by propagating seismic waves through the earth's subsurface and the
+solution to the optimization problem gives the most likely structure of the earth's subsurface. Figure
+\ref{fig:offshore_survey} shows the physical setup of the experiment that produces the data that this problem operates
+on. \cite{plessix2006review} The data access pattern from Figure \ref{fig:dataflow} applies here as well, since every step during the
+adjoint propagation requires data from the corresponding step from the forward propagation. In
+particular for FWI, this data can reach tens of terabytes in size, which is beyond the memory
+available on most hardware. While this paper is written to be focussed on FWI as the target problem, 
+the analysis can be applied to any problem with a similar data-access pattern.
 
@@
 
-Domain decomposition is often used not only to distribute the computational
-workload across more processors, but also to utilise the large amount of memory
-available in distributed systems. While this strategy is very powerful, the
-number of compute nodes and therefore the amount of memory that can be used
-efficiently is limited, for example by communication overheads that start to
-dominate as the domain is split into increasingly small
-pieces~\cite{virieux2009seismic}.
-Another common strategy in FWI is to only store values at the boundaries of the
-domain at each timestep, and reconstruct the rest of the wavefield when
-required~\cite{clapp2009reverse,yang2014rtm}.
-
-Checkpointing is yet another strategy to reduce the memory overhead. Only a
-subset of the program states during the forward pass is stored (and the rest
-discarded).  The discarded data is recomputed when needed by restarting the
-forward pass from the last available stored state. The question which states
-should be stored and which states should be recomputed to minimize the total
-amount of recomputation work has been answerd under certain assumptions with the
-Revolve algorithm~\cite{griewank2000algorithm}.  Other authors have subsequently
-developed extensions to Revolve that are optimal under less restrictive
-assumptions~\cite{wang2009minimal,aupy2016optimal,schanen2016asynchronous}.
+Various strategies are routinely employed to reduce this memory requirement, for example 
+Domain decomposition where the domain of the simulation is distributed over
+multiple nodes so as to fit within the memory available on each node. This introduces the overhead
+of network communication and might not efficiently use the computational resources on each node.\cite{virieux2009seismic}
+Another common strategy, which is specific to the domain of FWI, is where only the boundaries of
+the domain are stored at each timestep and the rest of the wavefield is reconstructed from the
+boundaries when required \cite{clapp2009reverse}\cite{yang2014rtm}. Another commonly used strategy that is applied across
+a much wider domain of problems is checkpointing. In this strategy, a subset of the data from the
+forward pass is stored (and the rest discarded). The discarded data is recomputed when required
+by rerunning the forward pass. Starting from Revolve \cite{griewank2000algorithm}, which is a provably optimal
+scheduling algorithm to implement this strategy, various studies have used and extended this
+technique to various scenarios \cite{wang2009minimal} \cite{aupy2016optimal} \cite{schanen2016asynchronous}.
 
@@
 
-Data compression is also increasingly used to reduce the memory footprint of
-scientific applications. General purpose data compression algorithms like Zlib
-(which is a part of gzip)~\cite{deutsch1996zlib}, and compression algorithms for
-video and image data such as JPEG-2000~\cite{skodras2001jpeg} have been
-presented in previous work. More recently, special purpose compression
-algorithms for floating-point scientific data have been developed, such as ZFP
-or SZ~\cite{Kaklamanis:2012aa,lindstrom2014fixed,di2018efficient}.
+Another strategy to reduce the memory footprint of scientific applications is
+data compression. Well-known general purpose data compression algorithms like
+Zlib (which is a part of gzip) \cite{deutsch1996zlib}, and compression for video and image
+data such as JPEG-2000 \cite{skodras2001jpeg}have been used for scientific data. More
+recently, special purpose compression algorithms for floating-point scientific data have been
+developed, such as ZFP or SZ \cite{lindstrom2014fixed} \cite{di2018efficient} 
+\cite{Kaklamanis:2012aa} \cite{ratanaworabhan2006fast}.
 
-Lossless algorithms guarantee that the exact original data can be recovered
-during decompression, whereas lossy algorithms introduce an error, but often
-guarantee that the error does not exceed certain absolute or relative error
-metrics. Typically, lossy compression is more effective in reducing the data
-size. Most popular compression packages offer various settings that allow a
-tradeoff between compression ratio, accuracy, and compression and decompression
-time. In this work we use the ZFP package to perform lossy compresssion.
+Lossless algorithms guarantee that the exact original data can be
+recovered during decompression, whereas lossy algorithms introduce an error, but
+often guarantee that the error does not exceed certain absolute or relative
+error metrics. Typically, lossy compression is more effective in reducing the
+data size, or in other words, offer a greater compression ratio. Most popular
+compression packages offer various settings that allow a tradeoff between
+compression ratio, accuracy, and compression and decompression time. In this
+work we use the ZFP package to perform lossy compresssion.
 
-It is worth noting that another data reduction strategy is to typecast values
-into a lower precision format, for example, from double precision to single
-precision. This can be seen as a computationally cheap lossy compression
-algorithm with a compression ratio of $2$.
+It is worth noting that another popular data reduction strategy is to typecast
+values into a lower precision format, for example, from double precision to
+single precision TODO cite. This can be seen as a computationally cheap lossy
+compression algorithm with a compression ratio of $2$.
 
@@
 reduction in data that needs to be communicated across MPI nodes or between a
-GPU and a host computer~\cite{gpu-compression}.
+GPU and a host computer. There has been some work on streaming compression
+of floating-point data for this specific intent. \cite{gpu-compression}
 
-Another way of using compression to speed up FWI or other adjoint-based methods
-is to use it \emph{instead} of checkpointing. If the compression ratio is
-sufficient to fit the entire data in memory, checkpointing is no longer
-necessary. Previous work has discussed this situation in the context of
-computational fluid dynamics~\cite{cyr2015towards} and FWI using compression
-method specifically designed for
-wavefields~\cite{dalmau2014lossy,boehm2016wavefield}.
+One possible way of using compression to speed up full-waveform inversion or any adjoint-based
+optimization is by simply compressing the field data before storing it and decompressing it when it
+is required again. If the memory available is sufficient to store the complete forward data after
+compression, the use of compression effectively replaces the use of revolve-based checkpointing.
+This scenario has been discussed in different studies before - they look at Revolve as an alternative
+to compression. This approach was seen for a study of adjoint-based methods on the Navier-Stokes
+equation \cite{cyr2015towards}. This was also studied in the context of seismic imaging and FWI
+to develop tailored compression methods for the wavefield data that is typically seen in this context - 
+including a compression method based on DCT \cite{dalmau2014lossy} and one based on adaptive
+coarsening \cite{boehm2016wavefield}. This paper focusses on a broader region, of which this scenario
+is just one part. In particular, we discuss the scenarios where the forward field can not fit in memory even
+after compression, where compression and Revolve would be used in conjunction.
 
-In this paper, we extend the previous studies by \emph{combining} checkpointing
-and compression. This is obviously useful when the data does not fit in the
-available memory even after compression, for example for very large adjoint
-problems, or for problems where the required accuracy limits the achievable
-compression ratios.
-
-Compared to using only checkpointing without compression, our combined method
-can often achieve better performance. This is achieved by reducing the size of
-stored program states, thereby allowing more program states to be stored during
-the forward computation. This in turn reduces the amount of recomputations that
-needs to be performed during the adjoint computation. On the other hand, the
-compression and decompression itself takes time.  We thereforen provide a
-comprehensive performance model that predicts whether or not the combined method
-is beneficial. We present various scenarios that use a TODO describe Richter and
-TODO describe JLSE/SKX node system, error tolerances for the lossy ZFP
-compression algorithm ranging from TODO to TODO, and different types of forward
-and adjoint computations that vary in their ratio between compute cost and state
-size.
-
-Our aim is to make this discussion independent of the hardware, application and
-compression algorithm. We therefore do not discuss whether or not the accuracy
-of the decompressed data is sufficient for our application, or whether
-other algorithms might achieve a better accuracy/compression tradeoff.
+Although most of the compression we discuss here introduces some error which may or may not affect
+the final result of the computation, we try to avoid discussing the error in this paper and focus on the 
+performance impact of compression instead.
 
@@
 of $287 \times 881 \times 881$ points, including an absorbing layer of 40 points on each side. The grid spacing
-was 25m in space and TODO in time. This was run for 2500 timesteps, and the final field stored as
+was 25m. This was run for 2500 timesteps, and the final field stored as
 representative of a typical snapshot. Figure \ref{fig:uncompressed} shows this stored field that was used for the compression tests. 
+The uncompressed size of this field was just under 900MB per timestep.
 
@@
 \subsection{Lossless}
-For our initial experiments, we used the python package \emph{blosc} (TODO CITE). It includes implementations for
-6 different lossless compression algorithms, namely ZLIB, ZSTD, BLOSC, LZ4, LZ4HC and TODO (TODO cite all of them).
-We did a parameter sweep over all available parameters to find the best-performing combinations. 
+For our initial experiments, we used the python package \emph{blosc} \cite{blosc}. It includes implementations for
+6 different lossless compression algorithms, namely ZLIB, ZSTD, BLOSCLZ,
+LZ4, LZ4HC and Snappy. We did a parameter sweep over all available parameters to find the best-performing combinations. 
+The best compression ratio we saw in this study was 1.18x which wasn't very promising.
 \subsection{Lossy}
-Next we tried lossy compression using ZFP \cite{ratanaworabhan2006fast} (TODO verify citation). To use ZFP from python, 
+Next we tried lossy compression using ZFP \cite{lindstrom2014fixed}. To use ZFP from python, 
 we first wrote a python wrapper over the reference implementation of ZFP, which is in C. We plan to release this wrapper
@@
 
-We tried all three modes of ZFP, namely fixed-tolerance, fixed-precision and fixed-rate modes. Figure TODO shows the compression ratios
-we saw for a variety of compression settings for the three modes. As expected, the highest compression efficiency was
+We tried all three modes of ZFP, namely fixed-tolerance, fixed-precision and fixed-rate modes. Figure \ref{fig:tolerance_cf_plot}
+shows the compression ratios we saw for a variety of compression settings for the three modes. As expected, the highest compression efficiency was
 observed in the fixed-tolerance mode albeit with highly unpredictable compression ratios. Figure \ref{fig:decompressed_error} shows the
@@
 platforms.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.8\linewidth]{images/tolerance-cf-richter.pdf}
+\end{center}
+\caption{Compression ratios achieved on compressing the wavefield}
+\label{fig:tolerance_cf_plot}
+\end{figure}
 
@@
 \end{equation}
-where $p(N, M)$ is the minimum number of recomputed steps from TODO cite revolve, reproduced
+where $p(N, M)$ is the minimum number of recomputed steps from \cite{griewank2000algorithm}, reproduced
 here in equation \ref{eqn:recompute}.

### 2018-10-03 plotter.py

@@ -26,6 +26,7 @@ def plot(x, y, filename, title, xlabel, ylabel):
     plt.title(title)
     plt.xticks(rotation=90)
     plt.xscale('log')
+    plt.yscale('log')
     plt.savefig(filename, bbox_inches='tight')
 
 if __name__ == "__main__":

### 2018-10-03 paper/paper.tex
revised section 2

--- 2018-10-03 paper/paper.tex
@@
 compressing and decompressing is less than the time saved by reduced
-recomputations. We present various scenarios that use a TODO describe Richter
-and TODO describe JLSE/SKX node system, error tolerances for the lossy ZFP
-compression algorithm ranging from TODO to TODO, and different types of forward
-and adjoint computations that vary in their ratio between compute cost and state
-size.
+recomputations.
+
+We present test cases that use a TODO describe Richter
+and TODO describe JLSE/SKX node system, henceforth called Richter and Skylake,
+respectively. We discuss performance results for varying error tolerances for
+the lossy ZFP compression algorithm ranging from TODO to TODO, and different
+types of forward and adjoint computations that vary in their ratio between
+compute cost and state size.
 
@@
 \section{Test case using Devito and pyRevolve}
-We use Devito \cite{devito-compiler} \cite{devito-api} to setup wave equation solvers for forward and
-adjoint modes. Devito is a domain-specific language to enable the rapid development of finite-difference
-solvers from a high-level description of the partial differential equation. The simplest version of the wave
-equation is the acoustic:
+We use Devito~\cite{devito-api,devito-compiler} to solve forward and adjoint
+wave equation problems. Devito is a domain-specific language to enable the
+rapid development of finite-difference solvers from a high-level description
+of a partial differential equation. The simplest version of the wave
+equation is the acoustic wave equation given by
 \begin{equation}
-m(x)\frac{\partial^2 u(t, x)}{\partial t^2} - \Laplace u(t, x) = q(t, x)
+m(x)\frac{\partial^2 u(t, x)}{\partial t^2} - \Laplace u(t, x) = q(t, x),
 \label{eqn:wave}
 \end{equation}
-where $m(x) = \frac{1}{c^2(x)}$ is the squared slowness with $c(x)$ the spatially dependent speed of sound, 
-$u(t, x)$ is the wavefield, $\Laplace u(t, x)$ denotes the laplacian of the wavefield and $q(t, x)$ is the source.
-Some of the kernels used in the performance measurements corresponded to a more physically representative
-but complex version of this equation called Tilted Transverse Isotropy (TTI) \cite{zhang2011stable}. We leave the 
-TTI equations out of this paper for brevity.
+where $m(x) = \frac{1}{c^2(x)}$ is the squared slowness, $c(x)$ the spatially
+dependent speed of sound, $u(t, x)$ is the wavefield, $\Laplace u(t, x)$
+denotes the laplacian of the wavefield and $q(t, x)$ is a source term.
+Some of our experiments in Section TODO are performed using a more accurate
+and more complex version of this equation called Tilted Transverse Isotropy
+(TTI)~\cite{zhang2011stable}. We leave the TTI equations out of this paper
+for brevity.
 
-The solution to equation \ref{eqn:wave} form the forward problem. The complete optimization problem involves
-the minimization of the misfit between the simulated signal and the observed signal as follows:
+The solution to equation~\ref{eqn:wave} forms the forward problem. The adjoint
+problem minimizes the misfit between simulated and observed signal given by
+\begin{equation}
+\min_{m} \phi_s(m) = \frac{1}{2} \left\lVert d_{sim} - d_{obs} \right\rVert_2^2.
+\end{equation}
+This optimization problem is usually solved using gradient-descent methods,
+where the gradient is computed using the adjoint-state method, which involves
+the data-flow pattern from figure \ref{fig:dataflow}.
 
-\begin{equation}
-\min_{m} \phi_s(m) = \frac{1}{2} \left\lVert d_{sim} - d_{obs} \right\rVert_2^2
-\end{equation}
-
-This optimization problem is usually solved using gradient-descent methods and the gradient calculation is done
-using the adjoint-state method, which involves the data-flow pattern from figure \ref{fig:dataflow}.
-
-The values of $m(x)$ were derived from the SEAM Overthrust model \cite{aminzadeh1996three} which was over a grid
-of $287 \times 881 \times 881$ points, including an absorbing layer of 40 points on each side. The grid spacing
-was 25m in space and TODO in time. This was run for 2500 timesteps, and the final field stored as
-representative of a typical snapshot. Figure \ref{fig:uncompressed} shows this stored field that was used for the compression tests.
+The values of $m(x)$ are in this work derived from the SEAM Overthrust
+model~\cite{aminzadeh1996three} over a grid of $287 \times 881 \times 881$
+points, including an absorbing layer of 40 points on each side. The grid spacing
+is 25m in space and TODO in time. This is run for 2500 timesteps. The wave field
+at the final time is shown in Figure~\ref{fig:uncompressed}, and only this field
+was used for the compression experiments in this paper.
 
@@
 
-To implement Revolve with Devito, we used pyRevolve \cite{kukreja2018high} which is a python wrapper over the
-original Revolve utility. The performance model that will be developed in section \ref{sec:performance_model} of
-this paper assumes that the implementation is similar to pyRevolve, which stores a checkpoint by copying a portion
-of the operator's working memory to the checkpointing memory and similarly loads a checkpoint by copying from the
-checkpointing memory to the operator's working memory. Although a perfect implementation of checkpointing may
-be able to avoid these copies, the overhead attached to these copies can be ignored for an operator that is
-sufficiently computationally intensive. However, we include the overheads in the model to verify this assumption.
-
-The experiments were run on two different systems, TODO description of Richter, henceforth called Richter and a
-TODO description of skylake, henceforth called Skylake.
+To implement Revolve with Devito, we use pyRevolve~\cite{kukreja2018high} which
+is a python wrapper for the Revolve algorithm. The performance model in
+section~\ref{sec:performance_model} assumes that the implementation is similar
+to pyRevolve, which stores a checkpoint by copying a portion of the operator's
+working memory to the checkpointing memory and similarly loads a checkpoint by
+copying from the checkpointing memory to the operator's working memory. Although
+a perfect implementation of checkpointing may be able to avoid these copies, the
+overhead attached to these copies can be ignored for an operator that is
+sufficiently computationally intensive. However, we include the overheads in the
+model to verify this assumption.
 

### 2018-10-03 paper/paper.tex

@@
 
-The values of $m(x)$ are in this work derived from the SEAM Overthrust
+The values of $m(x)$ used in this work are derived from the SEAM Overthrust
 model~\cite{aminzadeh1996three} over a grid of $287 \times 881 \times 881$
 points, including an absorbing layer of 40 points on each side. The grid spacing
-is 25m in space and TODO in time. This is run for 2500 timesteps. The wave field
+is 25m in space. This is run for 2500 timesteps. The wave field
 at the final time is shown in Figure~\ref{fig:uncompressed}, and only this field
-was used for the compression experiments in this paper.
+was used for the compression experiments in this paper. The uncompressed size of
+this single time step field is just under 900MB.
 

### 2018-10-03 paper/paper.tex
added acknowledgement and DOE box

@@
 TODO (Acknowledgments for Navjot)
-TODO (Acknowledgments for Jan)
-TODO (Acknowledgments for Mathias)
+
+TThis work was supported by the U.S. Department of Energy, Office of Science,
+Office of Advanced Scientific Computing Research, Applied Mathematics and
+Computer Science programs under contract number DE-AC02-06CH11357.
+We would also like to acknowledge the support from the SINBAD II project and
+the member organizations of the SINBAD Consortium.
+
+We gratefully acknowledge the computing resources provided and operated by the
+Joint Laboratory for System Evaluation (JLSE) at Argonne National Laboratory.
 
@@
 
+\vfill
+\begin{flushright}
+\normalsize
+\framebox{\parbox{4.8in}{The submitted manuscript has been created by UChicago
+Argonne, LLC, Operator of Argonne National Laboratory (`Argonne'). Argonne, a
+U.S. Department of Energy Office of Science laboratory, is operated under
+Contract No. DE-AC02-06CH11357. The U.S. Government retains for itself, and
+others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide
+license in said article to reproduce, prepare derivative works, distribute
+copies to the public, and perform publicly and display publicly, by or on behalf
+of the Government.  The Department of Energy will provide public access to these
+results of federally sponsored research in accordance with the DOE Public Access
+Plan. \texttt{http://energy.gov/downloads/doe-public-access-plan}.}}
+\end{flushright}
+
 \end{document}

### 2018-10-03 paper/paper.tex
some language changes

@@
 \subsection{Lossless}
-For our initial experiments, we used the python package \emph{blosc} \cite{blosc}. It includes implementations for
-6 different lossless compression algorithms, namely ZLIB, ZSTD, BLOSCLZ,
-LZ4, LZ4HC and Snappy. We did a parameter sweep over all available parameters to find the best-performing combinations. 
-The best compression ratio we saw in this study was 1.18x which wasn't very promising.
+We use the python package \emph{blosc}~\cite{blosc}, which includes implementations for
+six different lossless compression algorithms, namely ZLIB, ZSTD, BLOSCLZ,
+LZ4, LZ4HC and Snappy. An exhaustive search over all available parameters 
+resulted in a maximum compression ratio of 1.18x. We therefore focus on lossy
+compression in the remainder of this work.
+
 \subsection{Lossy}
-Next we tried lossy compression using ZFP \cite{lindstrom2014fixed}. To use ZFP from python, 
-we first wrote a python wrapper over the reference implementation of ZFP, which is in C. We plan to release this wrapper
-in the open source community shortly.
+We use the lossy compression package ZFP~\cite{lindstrom2014fixed} developed in
+C. To use ZFP from python, we developed a python wrapper for the reference
+implementation of ZFP.
 
@@
 
-We tried all three modes of ZFP, namely fixed-tolerance, fixed-precision and fixed-rate modes. Figure \ref{fig:tolerance_cf_plot}
-shows the compression ratios we saw for a variety of compression settings for the three modes. As expected, the highest compression efficiency was
-observed in the fixed-tolerance mode albeit with highly unpredictable compression ratios. Figure \ref{fig:decompressed_error} shows the
-spatial distribution of the error introduced after compression and decompression using fixed-tolerance mode. Since, in our experiments, the time
-taken to compress a checkpoint was in the same order of magnitude as the time taken to compute a timestep, we build a
-performance model in the next section to evaluate the utility of compression in different scenarios of problem setups and test
-platforms.
+ZFP supports three compression modes, namely fixed-tolerance, fixed-precision
+and fixed-rate. The fixed-tolerance mode limits the absolute error, while the
+fixed-precision mode limits the relative error to some user-specified value.
+The fixed-rate mode achieves a guaranteed compression ratio requested by the
+user, but does not provide any bounds on accuracy loss.
+Figure~\ref{fig:tolerance_cf_plot} shows compression ratios for a variety of
+settings for the three modes.
+
+The highest compression efficiency TODO define this is achieved in the
+fixed-tolerance mode albeit with highly unpredictable compression ratios. Figure
+\ref{fig:decompressed_error} shows the spatial distribution of the error after
+compression and decompression, compared to the original field, using
+fixed-tolerance mode.
+
+In our experiments using Devito, the time required to compress a checkpoint was
+in the same orderof magnitude as the time taken to compute a timestep.
 
@@
 \label{sec:performance_model}
-We work with the assumption that the computation of a single forward time step takes the same wall time
-as the computation of a single reverse time step - calling this $\mathbf{C}$. If the size of a single 
-timestep in memory is $\mathbf{S}$ and the simulation involves $\mathbf{N}$ timesteps, the minimum
-wall time required to run a single forward-adjoint evaluation is given by:
+We assume that the computation of a single forward time step takes the same wall
+time as the computation of a single reverse time step. We denote this time as
+$\mathbf{C}$. For a simulation with $\mathbf{N}$ timesteps, the minimum wall time required
+for the full forward-adjoint evaluation is given by
 \begin{equation}
-T_N = 2 \cdot \mathbf{C} \cdot \mathbf{N}
+T_N = 2 \cdot \mathbf{C} \cdot \mathbf{N}.
 \end{equation}
-However, this would require $\mathbf{S} \cdot \mathbf{N}$ memory. On a hardware platform where enough memory
-is available, this would be the optimal strategy to compute the adjoint solution.
+If the size of a single timestep in memory is given by $\mathbf{S}$, this
+requires a memory of at least size $\mathbf{S} \cdot \mathbf{N}$. If sufficient memory
+is available, no checkpointing or compression is needed.
 
-When the memory available on a system is less than $\mathbf{S} \cdot \mathbf{N}$, Revolve provides
-an optimal strategy to solve for the adjoint field by storing a subset of the $\mathbf{N}$ total checkpoints
+If the memory is smaller than $\mathbf{S} \cdot \mathbf{N}$, Revolve provides
+a strategy to solve for the adjoint field by storing a subset of the $\mathbf{N}$ total checkpoints
 and recomputing the remaining ones. The overhead introduced by this method can be broken down into
 the recomputation overhead $\mathbf{O}_R$ and the storage overhead $\mathbf{O}_S$. The recomputation
-overhead is simply the amount of time spent in recomputation, given by:
+overhead is the amount of time spent in recomputation, given by
 \begin{equation}
-\mathbf{O}_R(N, M) = p(N, M) \cdot \mathbf{C}
+\mathbf{O}_R(N, M) = p(N, M) \cdot \mathbf{C},
 \end{equation}
@@
 \end{figure*}
-In equation \ref{eqn:recompute}, M is the number of checkpoints that can be stored in memory. Note that for $M >=N$, $\mathbf{O}_R$
-would be zero. For $M < N$, $\mathbf{O}_R$ goes up rather quickly as M is reduced in relation to N.
+In equation \ref{eqn:recompute}, M is the number of checkpoints that can be
+stored in memory. Note that for $M >=N$, $\mathbf{O}_R$ would be zero. For $M <
+N$, $\mathbf{O}_R$ grows rapidly as M is reduced relative to N.
 
-In a perfect implementation, the storage overhead $\mathbf{O}_S$ might be zero, since the computation could
+In an ideal implementation, the storage overhead $\mathbf{O}_S$ might be zero, since the computation could
 be done "in-place", but in practice, checkpoints are generally stored in a separate section of memory and they
@@
 implementations, and might pose a non-trivial overhead in scenarios where the computation is heavily bandwidth-bound. 
-This storage overhead is given by:
+This storage overhead is given by
 \begin{equation}
-\mathbf{O}_{SR} = \mathbf{N}_W \cdot \frac{\mathbf{S}}{\mathbf{B}} + \mathbf{N}_R \cdot \frac{\mathbf{S}}{\mathbf{B}}
+\mathbf{O}_{SR} = \mathbf{N}_W \cdot \frac{\mathbf{S}}{\mathbf{B}} +
+\mathbf{N}_R \cdot \frac{\mathbf{S}}{\mathbf{B}}.
 \label{eqn:storage}
@@
 
-The total time to solution in this scenario becomes:
+The total time to solution in this scenario becomes
 \begin{equation}
-T_R = 2 \cdot \mathbf{C} \cdot \mathbf{N} + \mathbf{O}_R(N, M) + \mathbf{O}_S
+T_R = 2 \cdot \mathbf{C} \cdot \mathbf{N} + \mathbf{O}_R(N, M) + \mathbf{O}_S.
 \end{equation}
 
-Adding compression to the mix will increase the number of checkpoints available ($M$ in equation \ref{eqn:recompute}),
-hence reducing $\mathbf{O}_R$ while adding overheads related to compression and decompression in $\mathbf{O}_S$,
-as well as some error introduced by the compression. Assuming that the error remains within the allowed tolerance,
-we hypothesize that, at least for some circumstances, the reduction in $\mathbf{O}_R$ should offset the increase in 
-$\mathbf{O}_S$, i.e. lead to an overall decrease in the total time to solution ($T$). To build a performance model that
-enables us to find the regions where compression pays off, we assume that the compression algorithm behaves uniformly
+By using compression, the size of each checkpoint is reduced and therefore the
+number of checkpoints available is increased ($M$ in equation
+\ref{eqn:recompute}). This reduces the recomputation overhead $\mathbf{O}_R$,
+while at the same time adding overheads related to compression and decompression
+in $\mathbf{O}_S$.
+To be beneficial, the reduction in $\mathbf{O}_R$ must offset the increase in 
+$\mathbf{O}_S$, leading to an overall decrease in the time to solution $T$.
+
+Our performance model assumes
+ that the compression algorithm behaves uniformly
 across the different time steps of the simulation, i.e. that we get the same compression ratio, compression time and 
 decompression time, no matter which of the $N$ possible checkpoints we try to compress/decompress. The storage overhead
-now becomes:
+now becomes
 \begin{equation}
-\mathbf{O}_{SR} = \mathbf{N}_W \cdot (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} + t_c) + \mathbf{N}_R \cdot (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} + t_d)
+\mathbf{O}_{SR} = \mathbf{N}_W \cdot (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} +
+t_c) + \mathbf{N}_R \cdot (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} + t_d),
 \end{equation}
 where $\mathbf{F}$ is the compression ratio (i.e. the ratio between the uncompressed and compressed checkpoint), and $t_c$
-and $t_d$ are compression and decompression times, respectively. At the same time, the recomputation overhead goes down
+and $t_d$ are compression and decompression times, respectively. At the same
+time, the recomputation overhead decreases
 because $\mathbf{F}$ times more checkpoints are now available.
@@
 
-TThis work was supported by the U.S. Department of Energy, Office of Science,
+This work was supported by the U.S. Department of Energy, Office of Science,
 Office of Advanced Scientific Computing Research, Applied Mathematics and

### 2018-10-03 paper/paper.tex
Add plots

--- 2018-10-03 paper/paper.tex
@@
 \IEEEauthorblockN{Mathias Louboutin}
-\IEEEauthorblockA{\textit{TODO Mathias} \\
+\IEEEauthorblockA{\textit{School of Computational Science and Engineering} \\
 \textit{Georgia Institute of Technology}\\
@@
 \and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
+\IEEEauthorblockN{Fabio Luporini}
+\IEEEauthorblockA{\textit{Department of Earth Science and Engineering} \\
+\textit{Imperial College London}\\
+London, UK \\
+}
+\and
+\IEEEauthorblockN{Paul Hovland}
+\IEEEauthorblockA{\textit{Mathematics and Computer Science} \\
+\textit{Argonne National Laboratory}\\
+Lemont, IL, USA \\
+}
 \and
@@
 \end{center}
-\caption{Compression ratios achieved on compressing the wavefield}
+\caption{Compression ratios achieved on compressing the wavefield. The
+highlighted point corresponds to the setting used for the other results here
+unless otherwise specified.}
 \label{fig:tolerance_cf_plot}
@@
 \begin{equation}
-\mathbf{O}_{SR} = \mathbf{N}_W \cdot (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} + t_c) + \mathbf{N}_R \cdot (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} + t_d)
+\mathbf{O}_{SR} = \mathbf{N}_W \cdot (\frac{\mathbf{S}}{\mathbf{F}
+  \cdot \mathbf{B}} + t_c) + \mathbf{N}_R \cdot
+(\frac{\mathbf{S}}{\mathbf{F} \cdot \mathbf{B}} + t_d)
 \end{equation}
@@
 
-TODO add figure here that shows memory on x-axis and speedup on y axis. Split into 3 regions. Region 1, 
-where Revolve+Compression makes sense, Region 2, where compression would probably not bring in enough benefit, 
-and region 3 where compression would enable the computation to be done without revolve. 
 \section{Results}
+Figure \ref{fig:varying_memory} shows the speedups predicted by the
+performance model for varying amount of available memory. A big
+shortcoming that was realised in ZFP was the lack of parallelization
+in the decompression functions. The ZFP manual talks about the
+difficulty in implementing a parallel mode decompressor. To estimate
+the benefit from parallel decompression, we scaled the serial
+decompression time by the same factor that we saw between the serial
+and parallel compressors on a given platform. We called this scaled
+decompression time, ``theoretical decompression time'' and show this
+on the plot with a dotted line. The
+right-most region of this plot has been the focus of many studies
+mentioned earlier in this paper. The extension to the regions with
+lower memory is new here and we see that the maximum benefit is gained
+in the most memory-constrained scenarios (region 1), while there is a
+region where Revolve-based recomputation alone would be more
+beneficial (region 2). As expected, compression alone (with no
+Revolve) dominates region 3. 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\linewidth]{images/varying-memory.pdf}
+\end{center}
+\caption{The speedups predicted by the performance model for varying
+  memory. The baseline
+(1.0) is the performance of a Revolve-only implementation under the
+same conditions. The dotted curve represents the hypothetical speedup
+if ZFP supported parallel decompression. The two vertical lines at 20GB and 52GB clearly
+divide the plot into 3 regions. The left-most region, with less than
+20GB memory sees the maximum benefit from compression. The middle
+region, with memory between 20GB and 52GB, is better off using only
+Revolve and no compression. The right-most region, with memory greater
+than 52GB represents the scenario where the entire history of the
+wavefield can fit in memory post-compression. In this case compression
+eliminates the use of Revolve - hence the speedup. This is the region
+that has been well-studied in past work.}
+\label{fig:varying_memory}
+\end{figure}
 
+Figure \ref{fig:varying_compute} shows the speedups predicted by the
+performance model for kernels of different computational cost. Perhaps
+unsurprisingly, the benefit gained by compression drops rapidly if the
+time to compute a single timestep is less than the time to compress
+the computed data (since recomputing would be cheaper in that case). For
+increasing computational cost, the speedup from compression saturates
+at a bound. 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\linewidth]{images/varying-compute.pdf}
+\end{center}
+\caption{The speedups predicted by the performance model for varying
+  compute cost. The baseline
+(1.0) is the performance of a Revolve-only implementation under the
+same conditions. The benefits of compression drop rapidly if the
+computational cost of the kernel that generated the data is much lower
+than the cost of compressing the data. For increasing computational
+costs, the benefits are bounded.}
+\label{fig:varying_compute}
+\end{figure}
+
+Figure \ref{fig:varying_nt} shows the speedups predicted for differing
+number of timesteps. It is interesting to note that there is a
+consistent increase in the speedup as the number of timesteps goes
+up. 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\linewidth]{images/varying-nt.pdf}
+\end{center}
+\caption{The speedups predicted by the performance model for varying
+  number of timesteps to be reversed. The baseline
+(1.0) is the performance of a Revolve-only implementation under the
+same conditions. It can be seen that compression becomes more
+beneficial as the number of timesteps is increased.}
+\label{fig:varying_nt}
+\end{figure}
 \section{Conclusions and Future work}
@@
 \section*{Acknowledgment}
-
-TODO (Acknowledgments for Navjot)
-
-TThis work was supported by the U.S. Department of Energy, Office of Science,
+This work was funded by the Intel Parallel Computing Centre at
+Imperial College London and EPSRC EP/R029423/1 . 
+This work was supported by the U.S. Department of Energy, Office of Science,
 Office of Advanced Scientific Computing Research, Applied Mathematics and

### 2018-10-03 performance.py

@@ -127,7 +127,7 @@ def platform_name():
     return platform
 
 def plot(x, y, filename, title, xlabel, ylabel, hline=None, more_y=None, more_y_labels=None,
-         fixed=None, xscale=None, yscale=None):
+         fixed=None, xscale=None, yscale=None, vlines=None, textloc=None):
     plt.gcf().clear()
     if more_y_labels is not None:
         # We need labels for every series in more_y and one for the main series
@@ -156,13 +156,19 @@ def plot(x, y, filename, title, xlabel, ylabel, hline=None, more_y=None, more_y_
     if fixed is not None:
         fixed.update({"Platform" : platform_name()})
         fixed_params = ", \n".join(["%s: %s" % (key, value) for key, value in fixed.items()])
-        title += "\n" + fixed_params
+        if textloc is None:
+            textloc = [0.5, 0.5]
+        plt.figtext(textloc[0], textloc[1], fixed_params, size=8)
     plt.title(title)
     if xscale is not None:
         plt.xscale(xscale)
 
     if yscale is not None:
         plt.yscale(yscale)
+    if vlines is not None:
+        for v in vlines:
+            plt.axvline(v, linestyle='-.', color='green')
+            plt.text(v + 100,2.25,str(v),rotation=90)
     plt.savefig(filename, bbox_inches='tight')
 
 
@@ -180,9 +186,9 @@ bandwidth = 8139.2 # Skylake
 platform = "Skylake" # Richter
 
 # Compressions Params
-c_factor = 29.72
-c_time = 0.94
-d_time = 1.5
+c_factor = 41
+c_time = 0.36
+d_time = 1.6
 t_d_time = 0.396
 
 # Plot params
@@ -206,8 +212,8 @@ def varying_peak_memory(nt, size_ts, compute_ts, bw, c_factor, c_time, d_time,
              'Compression Factor': c_factor, 'Compression Time (s)': c_time,
              'Decompression time (s)': d_time,
              'Theoretical decompression time (s)': theoretical_d_time}
-    plot(mems, speedups, "varying-memory.png", "Speedup for varying peak memory", "Memory (MB)",
-             "Speedup (x)", hline=1, more_y=[theoretical_speedup], fixed=fixed, xscale='log')
+    plot(mems, speedups, "varying-memory.pdf", "Speedup for varying peak memory", "Memory (MB)",
[…]

### 2018-10-03 plotter.py

@@ -12,7 +12,7 @@ def read_csv_file(filename):
         data = []
         for i, row in enumerate(csv_reader):
             if i == 0:
-                headings = row
+                headings = [a.replace('"', '') for a in row]
             else:
                 data.append([float(x) for x in row])
     data = list(zip(*data))
@@ -27,6 +27,10 @@ def plot(x, y, filename, title, xlabel, ylabel, highlight=None):
     plt.xticks(rotation=90)
     plt.xscale('log')
     plt.yscale('log')
+    plt.axhline(1, linestyle='dashed', color='red')
+    if highlight is not None:
+        for i in highlight:
+            plt.plot(x[i], y[i], marker='^')
     plt.savefig(filename, bbox_inches='tight')
 
 if __name__ == "__main__":
@@ -45,4 +49,4 @@ if __name__ == "__main__":
     title = args.title
     output_file = args.output_file
     data, headings = read_csv_file(filename)
-    plot(data[xindex], data[yindex], output_file, title, headings[xindex], headings[yindex])
+    plot(data[xindex], data[yindex], output_file, title, headings[xindex], headings[yindex], highlight=[3])

### 2018-10-03 paper/paper.tex

--- 2018-10-03 paper/paper.tex
@@
 \begin{equation}
-\mathbf{O}_{SR} = \mathbf{N}_W \cdot (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} +
-t_c) + \mathbf{N}_R \cdot (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} + t_d),
+\mathbf{O}_{SR} = \mathbf{N}_W \cdot (\frac{\mathbf{S}}{\mathbf{F}
+  \cdot \mathbf{B}} + t_c) + \mathbf{N}_R \cdot
+(\frac{\mathbf{S}}{\mathbf{F} \cdot \mathbf{B}} + t_d)
 \end{equation}
@@
 \section*{Acknowledgment}
-
-TODO (Acknowledgments for Navjot)
-
+This work was funded by the Intel Parallel Computing Centre at
+Imperial College London and EPSRC EP/R029423/1 . 
 This work was supported by the U.S. Department of Energy, Office of Science,

### 2018-10-03 paper/paper.tex
minor typos

@@
 \usepackage{pgfplots}
+\usepackage{url}
 
@@
 In an ideal implementation, the storage overhead $\mathbf{O}_S$ might be zero, since the computation could
-be done "in-place", but in practice, checkpoints are generally stored in a separate section of memory and they
-need to be transferred to a "computational" section of the memory where the compute is performed, and then
+be done ``in-place'', but in practice, checkpoints are generally stored in a separate section of memory and they
+need to be transferred to a ``computational'' section of the memory where the compute is performed, and then
 the results copied back to the checkpointing memory. This copying is a common feature of checkpointing
@@
 \begin{equation}
-\mathbf{O}_{SR} = \mathbf{N}_W \cdot (\frac{\mathbf{S}}{\mathbf{F}
-  \cdot \mathbf{B}} + t_c) + \mathbf{N}_R \cdot
-(\frac{\mathbf{S}}{\mathbf{F} \cdot \mathbf{B}} + t_d)
+\mathbf{O}_{SR} = \mathbf{N}_W \cdot \left(\frac{\mathbf{S}}{\mathbf{F}
+  \cdot \mathbf{B}} + t_c\right) + \mathbf{N}_R \cdot
+\left(\frac{\mathbf{S}}{\mathbf{F} \cdot \mathbf{B}} + t_d\right)
 \end{equation}
@@
 \section{Results}
-Figure \ref{fig:varying_memory} shows the speedups predicted by the
+Figure~\ref{fig:varying_memory} shows the speedups predicted by the
 performance model for varying amount of available memory. A big
@@
 
-Figure \ref{fig:varying_compute} shows the speedups predicted by the
+Figure~\ref{fig:varying_compute} shows the speedups predicted by the
 performance model for kernels of different computational cost. Perhaps
@@
 
-Figure \ref{fig:varying_nt} shows the speedups predicted for differing
+Figure~\ref{fig:varying_nt} shows the speedups predicted for differing
 number of timesteps. It is interesting to note that there is a
@@
 \end{figure}
+
 \section{Conclusions and Future work}
@@
 \normalsize
-\framebox{\parbox{4.8in}{The submitted manuscript has been created by UChicago
+\framebox{\parbox{3in}{The submitted manuscript has been created by UChicago
 Argonne, LLC, Operator of Argonne National Laboratory (`Argonne'). Argonne, a
@@
 results of federally sponsored research in accordance with the DOE Public Access
-Plan. \texttt{http://energy.gov/downloads/doe-public-access-plan}.}}
+Plan.\newline \url{http://energy.gov/downloads/doe-public-access-plan}.}}
 \end{flushright}

### 2018-10-03 paper/paper.tex
Add text in conclusions

--- 2018-10-03 paper/paper.tex
@@
 \end{center}
-\caption{Cross-section of the wavefield used as a reference sample for compression and decompression}
+\caption{Cross-section of the wavefield used as a reference sample for
+  compression and decompression. This field was formed after a Ricker
+  wavelet source was placed at TODO and the wave propagated for 2500
+  timesteps. This is a vertical cross-section of a 3D field, taken at
+  the source location}
 \label{fig:uncompressed}
@@
 \end{center}
-\caption{Cross-section of the field that shows errors introduced during compression and decompression using the fixed-tolerance mode}
+\caption{Cross-section of the field that shows errors introduced
+  during compression and decompression using the fixed-tolerance
+  mode. It is interesting to note that the errors are more or less
+  evenly distributed across the domain with only slight variations
+  corresponding to the wave amplitude (from the field plot in figure
+  \ref{uncompressed}. A small block-like structure characteristic of
+  ZFP can be seen.}
 \label{fig:decompressed_error}
@@
 \section{Results}
+We notice that the combination of compression and checkpointing
+outperforms pure checkpointing in some scenarios, and that wherever
+possible, compression alone outperforms checkpointing. This is made
+evident in figure \ref{fig:varying_memory} which shows a clear split
+in three different regions. The right-most region of this plot has
+been the focus of many studies mentioned earlier in this paper. The
+extension to the regions with lower memory is new here and we see that
+the maximum benefit is gained in the most memory-constrained scenarios
+(region 1), while there is a region where Revolve-based recomputation
+alone would be more beneficial (region 2). As expected, compression
+alone (with no Revolve) dominates region 3.
 Figure \ref{fig:varying_memory} shows the speedups predicted by the
@@
 decompression time, ``theoretical decompression time'' and show this
-on the plot with a dotted line. The
-right-most region of this plot has been the focus of many studies
-mentioned earlier in this paper. The extension to the regions with
-lower memory is new here and we see that the maximum benefit is gained
-in the most memory-constrained scenarios (region 1), while there is a
-region where Revolve-based recomputation alone would be more
-beneficial (region 2). As expected, compression alone (with no
-Revolve) dominates region 3. 
+on the plot with a dotted line. 
 \begin{figure}
@@
 \section{Conclusions and Future work}
+What are the problems?
+The performance we developed might be useful in judging whether compression
+would be helpful in a particular scenario without actually building a
+complete implementation first. We have tried to keep it application
+and platform agnostic in order to maximize its applicability. However
+there are a few limitations in this model. Firstly, we assumed that
+the compression performance we saw for a single checkpoint is
+representative of any checkpoint throughout the simulation. This
+assumption might have been completely wrong if made about a checkpoint
+in the first few time steps since we consistently saw upto 800x
+compression in those cases - since they are mostly zeros. In order for
+this assumption to be realistic, we use a checkpoint close to the end
+of the simulation so that most of the field is perturbed and hence is
+close to the worst case in terms of compression results.
 
-What are the problems?
+In comparing pure compression with pure checkpointing, we assume that every
+checkpoint would be compressed and decompressed. However, in scenarios
+where the available memory is only slightly less than the required
+memory, an implementation that compresses only a subset of the
+checkpoints might outperform the expectations of our model.
 
 Future work:
+
+This work is focussed on checkpointing in memory alone. In a future
+work we would like to extend this performance model to include
+checkpointing to disk.
+
+The ZFP algorithm that the majority of this paper was based on, only
+supports fields with up to 3 dimensions. Since our field is evolving
+in time, we would expect some correlations in the time dimensions that
+are currently not being exploited in the compression.
+
+As can be seen in figure \ref{fig:varying_memory}, the lack of
+parallel decompression has a significant impact on the overall time to
+solution. This strategy would become much more beneficial once we
+start seeing faster decompression speeds.
+
+In this work, we have ignored the error that is induced by the lossy
+compression - assuming that a tolerance of $10^{-4}$ is acceptable for
+the application. This assumption needs to be verified in a formal
+analysis of how the error in the field propagates through the
+simulation, and how it finally affects the cross-correlated gradient. 
+The development of a priori and a posteriori error metrics would
+greatly help the further use of lossy compression in similar
+problems. 
 \begin{itemize}
@@
 
-More than 3D fields (use time correlation)
-
-Parallel decompression
-
-Disk checkpointing
-
 \section*{Acknowledgment}
 This work was funded by the Intel Parallel Computing Centre at
-Imperial College London and EPSRC EP/R029423/1 . 
+Imperial College London and EPSRC EP/R029423/1. 
 This work was supported by the U.S. Department of Energy, Office of Science,

### 2018-10-03 performance.py

@@ -243,7 +243,7 @@ def varying_compute(nt, size_ts, peak_mem, bw, c_factor, c_time, d_time, more_pe
              'Compression Factor': c_factor, 'Compression Time (s)': c_time,
              'Decompression time (s)': d_time,}
     plot(computes, speedups, "varying-compute.pdf",
-         "Speedup for varying compute time/timestep", "Compute time (s)", "Speedup (x)",
+         "Speedup for varying compute/timestep", "Compute time (s)", "Speedup (x)",
          hline=1, more_y=more_speedups,
          more_y_labels=["Peak memory: %d" % x for x in [peak_mem] + more_peak_mem],
          fixed=fixed, xscale='log', textloc=[0.55, 0.55])

### 2018-10-04 performance.py
Update plots

@@ -5,6 +5,8 @@ import math
 import numpy as np
 import matplotlib.pyplot as plt
 import csv
+from matplotlib import ticker
+
 
 
 class CounterOperator(Operator):
@@ -85,6 +87,8 @@ class Problem(object):
 
     def revolve(self, peak_memory):
         ncp = math.floor(peak_memory/self.size_ts)
+        if ncp >= self.nt:
+            ncp = self.nt
         advances, takeshots, useshots = revolve_factors(self.nt, ncp)
         assert(advances >= self.nt)
         time_fw = self.compute_ts * advances
@@ -97,7 +101,7 @@ class Problem(object):
         ncp = math.floor(peak_memory/compressed_size)
 
         if ncp >= self.nt:
-            return self.compression_only(c_time, d_time)
+            ncp = self.nt
         
         # If we're here, we can't fit the whole history in memory even after compression
         advances, takeshots, useshots = revolve_factors(self.nt, ncp)
@@ -121,13 +125,16 @@ class Problem(object):
 def linspace(lower, upper, length):
     return [lower + x*(upper-lower)/length for x in range(length)]
 
+def logspace(lower, upper, length, base=10):
+    return np.logspace(math.log(lower, base), math.log(upper, base), base=base, num=length)
+
 
 def platform_name():
     global platform
     return platform
 
 def plot(x, y, filename, title, xlabel, ylabel, hline=None, more_y=None, more_y_labels=None,
-         fixed=None, xscale=None, yscale=None, vlines=None, textloc=None):
+         fixed=None, xscale=None, yscale=None, vlines=None, textloc=None, basex=None):
     plt.gcf().clear()
     if more_y_labels is not None:
         # We need labels for every series in more_y and one for the main series
@@ -151,7 +158,7 @@ def plot(x, y, filename, title, xlabel, ylabel, hline=None, more_y=None, more_y_
                 label = None
             plt.plot(x, series, linestyle='dotted', label=label)
         if more_y_labels is not None:
-            plt.legend()
+            plt.legend(loc=9)
     
     if fixed is not None:
         fixed.update({"Platform" : platform_name()})
@@ -161,14 +168,18 @@ def plot(x, y, filename, title, xlabel, ylabel, hline=None, more_y=None, more_y_
         plt.figtext(textloc[0], textloc[1], fixed_params, size=8)
     plt.title(title)
     if xscale is not None:
-        plt.xscale(xscale)
[…]

### 2018-10-04 paper/paper.tex
revised result section

@@
 \IEEEauthorblockN{Kaiyuan Hou}
-\IEEEauthorblockA{\textit{Department of EECS} \\
+\IEEEauthorblockA{\textit{Department of Electrical Engineering and Computer Science} \\
 \textit{Northwestern University}\\
@@
 \IEEEauthorblockN{Paul Hovland}
-\IEEEauthorblockA{\textit{Mathematics and Computer Science} \\
+\IEEEauthorblockA{\textit{Mathematics and Computer Science Division} \\
 \textit{Argonne National Laboratory}\\
@@
 \section{Results}
-We notice that the combination of compression and checkpointing
-outperforms pure checkpointing in some scenarios, and that wherever
-possible, compression alone outperforms checkpointing. This is made
-evident in figure \ref{fig:varying_memory} which shows a clear split
-in three different regions. The right-most region of this plot has
-been the focus of many studies mentioned earlier in this paper. The
-extension to the regions with lower memory is new here and we see that
-the maximum benefit is gained in the most memory-constrained scenarios
-(region 1), while there is a region where Revolve-based recomputation
-alone would be more beneficial (region 2). As expected, compression
-alone (with no Revolve) dominates region 3.
-Figure \ref{fig:varying_memory} shows the speedups predicted by the
-performance model for varying amount of available memory. A big
-shortcoming that was realised in ZFP was the lack of parallelization
-in the decompression functions. The ZFP manual talks about the
-difficulty in implementing a parallel mode decompressor. To estimate
-the benefit from parallel decompression, we scaled the serial
-decompression time by the same factor that we saw between the serial
-and parallel compressors on a given platform. We called this scaled
-decompression time, ``theoretical decompression time'' and show this
-on the plot with a dotted line. 
+We can distinguish three different scenarios, depending on the amount of
+available memory.
+\begin{enumerate}
+\item If the memory is insufficient even with compression to store the entire
+trajectory, one can either use checkpointing only, or combine checkpointing with
+compression.
+\item If the available memory is not sufficient to store the uncompressed
+trajectory, but large enough to store the entire compressed trajectory, we study
+the two possible strategies: Either use compression only, or use checkpointing
+only.
+\item If the available system memory is large enough to hold the entire
+uncompressed trajectory, neither compression nor checkpointing is necessary.
+\end{enumerate}
+All three scenarios can be seen in figure~\ref{fig:varying_memory}.
+The second scenario was studied in previous work~\cite{cyr2015towards}, while
+the combined method is also applicable to the first scenario, for which previous
+work has only used checkpointing without compression.
+
+We can identify a number of factors that make compression more likely to be
+beneficial compared to pure checkpointing: A very small system memory size and a
+large number of time steps lead to a rapidly increasing recompute factor, and
+compression can substantially reduce this recompute factor. This can be seen in
+figures~\ref{fig:varying_memory} and~\ref{fig:varying_nt}.
+
+The extent to which the recompute factor affects the overall runtime also
+depends on the cost to compute each individual time step. If the compute cost
+per time step is large compared to the compression and decompression cost, then
+compression is also likely to be beneficial, as shown in
+figure~\ref{fig:varying_compute}. As the time per time step increases and the
+compression cost becomes negligible, we observe that the ratio between the runtime
+of the combined method and that of pure checkpointing is only determined by the
+difference in recompute factors.
+
 \begin{figure}
@@
 
-Figure~\ref{fig:varying_compute} shows the speedups predicted by the
-performance model for kernels of different computational cost. Perhaps
-unsurprisingly, the benefit gained by compression drops rapidly if the
-time to compute a single timestep is less than the time to compress
-the computed data (since recomputing would be cheaper in that case). For
-increasing computational cost, the speedup from compression saturates
-at a bound. 
 \begin{figure}
@@
 
-Figure~\ref{fig:varying_nt} shows the speedups predicted for differing
-number of timesteps. It is interesting to note that there is a
-consistent increase in the speedup as the number of timesteps goes
-up. 
 \begin{figure}
@@
 checkpoints might outperform the expectations of our model.
+
+Our current experimental setup has two  that was realised in ZFP was the lack of parallelization
+in the decompression functions. The ZFP manual talks about the
+difficulty in implementing a parallel mode decompressor.
 

### 2018-10-04 performance.py
Change the dataflow figure to single-column

@@ -232,9 +232,10 @@ def varying_peak_memory(nt, size_ts, compute_ts, bw, c_factor, c_time, d_time,
              'Compression Factor': c_factor, 'Compression Time (s)': c_time,
              'Decompression time (s)': d_time,
              'Theoretical decompression time (s)': theoretical_d_time}
+    legend_labels = ["Measured (TTI)"] + ["%.2f x (c_time + d_time)" % x for x in computes]
     plot(mems, speedups, "varying-memory.pdf", "Speedup for varying peak memory", "Memory (MB)",
-             "Speedup (x)", hline=1, vlines=[overall_peak_mem, overall_peak_mem/c_factor], more_y=more_computes, fixed=fixed, xscale='log', textloc=[0.5, 0.35], more_y_labels=["Orig"]+computes,
-             legend_placement=9)
+             "Speedup (x)", hline=1, vlines=[overall_peak_mem, overall_peak_mem/c_factor], more_y=more_computes, fixed=fixed, xscale='log', textloc=[0.5, 0.40], more_y_labels=legend_labels,
+             legend_placement=(0.41, 0.69))
 
 
 def varying_peak_memory_total(nt, size_ts, compute_ts, bw, c_factor, c_time, d_time,
@@ -258,7 +259,6 @@ def varying_peak_memory_total(nt, size_ts, compute_ts, bw, c_factor, c_time, d_t
              'Compression Factor': c_factor, 'Compression Time (s)': c_time,
              'Decompression time (s)': d_time,
              'Theoretical decompression time (s)': theoretical_d_time}
-
     plot(mems, times_revolve, "varying-memory-total.pdf", "Execution time for varying peak memory", "Memory (MB)",
              "Execution time (s)", more_y=[times_compression], fixed=fixed, xscale='log', textloc=[0.5, 0.5])
 

### 2018-10-04 paper/paper.tex
Change the dataflow figure to single-column
revised conclusion

@@
 \section{Conclusions and Future work}
-What are the problems?
-The performance we developed might be useful in judging whether compression
-would be helpful in a particular scenario without actually building a
-complete implementation first. We have tried to keep it application
-and platform agnostic in order to maximize its applicability. However
-there are a few limitations in this model. Firstly, we assumed that
-the compression performance we saw for a single checkpoint is
-representative of any checkpoint throughout the simulation. This
-assumption might have been completely wrong if made about a checkpoint
-in the first few time steps since we consistently saw upto 800x
-compression in those cases - since they are mostly zeros. In order for
-this assumption to be realistic, we use a checkpoint close to the end
-of the simulation so that most of the field is perturbed and hence is
-close to the worst case in terms of compression results.
+We use lossy compression to reduce the computational overhead of checkpointing
+in an adjoint computation used in full waveform inversion, a common method in
+seismic imaging applications whose memory footprint commonly exceeds the
+available memory size in high performance computing systems. We also developed a
+performance model that computes whether or not the combination of compression
+and checkpointing will outperform pure checkpointing or pure compression in a
+variety of scenarios, depending on the available memory size, computational
+intensity of the application, and compression ratio and throughput of the
+compression algorithm.
 
-In comparing pure compression with pure checkpointing, we assume that every
-checkpoint would be compressed and decompressed. However, in scenarios
-where the available memory is only slightly less than the required
-memory, an implementation that compresses only a subset of the
-checkpoints might outperform the expectations of our model.
-
-Our current experimental setup has two  that was realised in ZFP was the lack of parallelization
-in the decompression functions. The ZFP manual talks about the
-difficulty in implementing a parallel mode decompressor.
-
-Future work:
-
-This work is focussed on checkpointing in memory alone. In a future
-work we would like to extend this performance model to include
-checkpointing to disk.
-
-The ZFP algorithm that the majority of this paper was based on, only
-supports fields with up to 3 dimensions. Since our field is evolving
-in time, we would expect some correlations in the time dimensions that
-are currently not being exploited in the compression.
-
-As can be seen in figure \ref{fig:varying_memory}, the lack of
-parallel decompression has a significant impact on the overall time to
-solution. This strategy would become much more beneficial once we
-start seeing faster decompression speeds.
-
-In this work, we have ignored the error that is induced by the lossy
-compression - assuming that a tolerance of $10^{-4}$ is acceptable for
-the application. This assumption needs to be verified in a formal
-analysis of how the error in the field propagates through the
-simulation, and how it finally affects the cross-correlated gradient. 
-The development of a priori and a posteriori error metrics would
-greatly help the further use of lossy compression in similar
-problems. 
+Our current result has several limitations that we plan to address in future
+work:
 \begin{itemize}
-\item Acceptable error bounds
-\item A posteriori and a priori metrics
-\item Relationship between compression error and field error, propagation of errors in forward simulation
-\item Relationship between compression error and gradient error in cross-correlation
-\item What else?
+\item We do not discuss the accuracy of the results after decompression. This
+depends on the application, compression algorithm, and affects the achievable
+compression ratios. Our performance model only requires knowledge of the
+compression time and ratio, and it is up to the user of this model to determine
+what accuracy is needed and thus what compression ratio is realistic for their
+application.
+\item We only present results using the ZFP compression library. Again, the
+performance model is independent of this, but it would be interesting to try a
+greater variety of compression algorithms.
+\item ZFP currently only supports serial decompression. If ZFP supported
+parallel decompression, our experiments would likely show a geater region in
+which the combined method is faster than pure checkpointing. Furthermore, ZFP
+only supports fields with up to three dimensions, and it is possible that
+exploiting similarities between fields at different time steps would yield a
+better compression ratio.
+\item Our performance model is based on uniform compression ratios and times.
+However, many applications, including FWI, are likely to have initial conditions
+that contain little information and are easily compressed, and the compression
+ratio gradually declines as the field becomes more complex over time. We have
+based our experiments on the final wave field, which is presumably the most
+difficult to compress.
+\item In comparing pure compression with pure checkpointing, we assume that
+every checkpoint would be compressed and decompressed. However, in scenarios
+where the available memory is only slightly less than the required memory, an
+implementation that compresses only a subset of the checkpoints might outperform
+the expectations of our model. 
+\item We do not discuss multi-level checkpointing, where some checkpoints are
+stored on a slower, larger device such as a solid state drive. We expect
+compression to be particularly beneficial in these scenarios due to the reduced
+data transfer sizes.
 \end{itemize}

### 2018-10-04 paper/paper.tex
Version sent to coauthors

@@
 
-\begin{figure*}
+\begin{figure}
 \begin{center}
@@
 \label{fig:dataflow}
-\end{figure*}
+\end{figure}
 
@@
 (1.0) is the performance of a Revolve-only implementation under the
-same conditions. The dotted curve represents the hypothetical speedup
-if ZFP supported parallel decompression. The two vertical lines at 20GB and 52GB clearly
-divide the plot into 3 regions. The left-most region, with less than
-20GB memory sees the maximum benefit from compression. The middle
-region, with memory between 20GB and 52GB, is better off using only
-Revolve and no compression. The right-most region, with memory greater
-than 52GB represents the scenario where the entire history of the
-wavefield can fit in memory post-compression. In this case compression
-eliminates the use of Revolve - hence the speedup. This is the region
-that has been well-studied in past work.}
+same conditions. The different curves represent kernels with differing
+compute times (represented here as a factor of the sum of compression
+and decompression times). The first vertical line at $~$ 53GB marks the
+spot where the compressed wavefield can completely fit in memory and
+Revolve is unnecessary if using compression. The second vertical line
+at $~$ 2.2 TB marks the spot where the entire uncompressed wavefield can
+fit in memory and neither Revolve nor compression is necessary. The
+region to the right is where these optimisations are not necessary or
+relevant. The middle region has been the subject of past studies using
+compression in adjoint problems. The region to the left is the focus
+of this paper.}
 \label{fig:varying_memory}

### 2018-10-04 paper/paper.tex
some rephrasing

--- 2018-10-04 paper/paper.tex
@@
 
-\title{Combining checkpointing and data compression\\ for high frequency seismic imaging
-}
+\title{Combining checkpointing and data compression\\ for large scale seismic inversion}
 
@@
 \begin{abstract}
-Full-waveform inversion is an adjoint-based optimization problem in seismic
-imaging that processes terabytes of data, regularly exceeding the memory
+Seismic inversion and imaging are adjoint-based optimization problems that processes up to terabytes of data, regularly exceeding the memory
 capacity of available computers. Data compression is an effective strategy to
@@
 In this paper we combine compression and checkpointing for the first
-time to compute a realistic full-waveform inversion. The combination of
+time to compute a realistic seismic inversion. The combination of
 checkpointing and compression allows
@@
 applications in science and engineering have this structure, including
-full-waveform inversion (FWI).
+seismic inversion.
 
-In FWI, the propagation of seismic waves through the earth's subsurface is
-simulated and compared with data from real-world measurements. The model of the
-subsurface is iteratively improved by minimizing the misfit between simulation and
-measurement in an adjoint optimization problem.
-Figure~\ref{fig:offshore_survey} shows the setup of the real-world experiment that
-produces the measurement data~\cite{plessix2006review}. Storing all intermediate
+In seismic inversion, the propagation of seismic waves through the earth's subsurface is simulated and compared with data from field measurements.
+The model of the subsurface is iteratively improved by minimizing the misfit between simulated data and field measurement in an adjoint optimization problem.
+Figure~\ref{fig:offshore_survey} shows the setup of the field experiment that produces the measured data~\cite{plessix2006review}. Storing all intermediate
 results as required for the data flow shown in Figure~\ref{fig:dataflow} is not
@@
 problem, a number of strategies are used that are discussed in the following
-paragraphs. While this paper uses FWI as a test case, most of these ideas,
-including the one presented in this work, are also applicable to other adjoint
-optimization problems.
+paragraphs. While this paper uses seismic inversion as a test case, most of these ideas, including the one presented in this work, are also applicable to other adjoint optimization problems.
 
@@
 \end{center}
-\caption{Illustration of an offshore seismic survey that collects the input data for FWI}
+\caption{Illustration of an offshore seismic survey that collects the input data for seismic inversion}
 \label{fig:offshore_survey}
@@
 Domain decomposition is often used not only to distribute the computational
-workload across more processors, but also to utilise the large amount of memory
+workload across more processors, but also to utilize the large amount of memory
 available in distributed systems. While this strategy is very powerful, the
@@
 efficiently is limited, for example by communication overheads that start to
-dominate as the domain is split into increasingly small
-pieces~\cite{virieux2009seismic}.
-Another common strategy in FWI is to only store values at the boundaries of the
-domain at each timestep, and reconstruct the rest of the wavefield when
-required~\cite{clapp2009reverse,yang2014rtm}.
+dominate as the domain is split into increasingly small pieces~\cite{virieux2009seismic}.
+Secondly, this method is broadly used on conventional clusters but can be drastically more complicated to seup and slower due to the communication in the Cloud.
+Another common strategy in seismic inversion is to only store values at the boundaries of the domain at each timestep, and reconstruct the rest of the wavefield when required~\cite{clapp2009reverse,yang2014rtm} with time reversal of the wave equation.
+However, this method is not applicable for wave equations that are not time reversible when for example physical attenuation is included.
 
@@
 should be stored and which states should be recomputed to minimize the total
-amount of recomputation work has been answerd under certain assumptions with the
-Revolve algorithm~\cite{griewank2000algorithm}.  Other authors have subsequently
-developed extensions to Revolve that are optimal under less restrictive
-assumptions~\cite{wang2009minimal,aupy2016optimal,schanen2016asynchronous}.
+amount of recomputation work has been answered under certain assumptions with the Revolve algorithm~\cite{griewank2000algorithm}.
+Other authors have subsequently developed extensions to Revolve that are optimal under less restrictive assumptions~\cite{wang2009minimal,aupy2016optimal,schanen2016asynchronous}.
 
@@
 tradeoff between compression ratio, accuracy, and compression and decompression
-time. In this work we use the ZFP package to perform lossy compresssion.
+time. In this work we use the ZFP package to perform lossy compression.
 
@@
 
-Another way of using compression to speed up FWI or other adjoint-based methods
-is to use it \emph{instead} of checkpointing. If the compression ratio is
-sufficient to fit the entire data in memory, checkpointing is no longer
-necessary. Previous work has discussed this situation in the context of
-computational fluid dynamics~\cite{cyr2015towards} and FWI using compression
-method specifically designed for
+Another way of using compression to speed up seismic inversion or other adjoint-based methods is to use it \emph{instead} of checkpointing.
+If the compression ratio is sufficient to fit the entire data in memory, checkpointing is no longer necessary.
+Previous work has discussed this situation in the context of computational fluid dynamics~\cite{cyr2015towards} and seismic inversion using compression method specifically designed for
 wavefields~\cite{dalmau2014lossy,boehm2016wavefield}.
@@
 
-Compared to using only checkpointing without compression, our combined method
-often improves performance. This is a consequence of the reduced size of stored
-program states, allowing more program states to be stored during the forward
-computation. This in turn reduces the amount of recomputation that needs to be
-performed. On the other hand, the compression and decompression itself takes
-time. We thereforen provide a comprehensive performance model that predicts
-whether the combined method is beneficial, that is, whether the time spent
-compressing and decompressing is less than the time saved by reduced
+Compared to the use of only checkpointing without compression, our combined method often improves performance.
+This is a consequence of the reduced size of stored program states, allowing more program states to be stored during the forward computation.
+This in turn reduces the amount of recomputation that needs to be performed.
+On the other hand, the compression and decompression itself takes time.
+We therefore provide a comprehensive performance model that predicts whether the combined method is beneficial, that is, whether the time spent compressing and decompressing is less than the time saved by reduced
 recomputations.
@@
 We use Devito~\cite{devito-api,devito-compiler} to solve forward and adjoint
-wave equation problems. Devito is a domain-specific language to enable the
+wave equation problems. Devito is a domain-specific language that enables the
 rapid development of finite-difference solvers from a high-level description
-of a partial differential equation. The simplest version of the wave
-equation is the acoustic wave equation given by
+of partial differential equations. The simplest version of the seismic wave
+equation is the acoustic isotropic wave equation defined as:
 \begin{equation}
@@
 where $m(x) = \frac{1}{c^2(x)}$ is the squared slowness, $c(x)$ the spatially
-dependent speed of sound, $u(t, x)$ is the wavefield, $\Laplace u(t, x)$
-denotes the laplacian of the wavefield and $q(t, x)$ is a source term.
+dependent speed of sound, $u(t, x)$ is the pressure wavefield, $\Laplace u(t, x)$ denotes the laplacian of the wavefield and $q(t, x)$ is a source term.
 Some of our experiments in Section TODO are performed using a more accurate
 and more complex version of this equation called Tilted Transverse Isotropy
-(TTI)~\cite{zhang2011stable}. We leave the TTI equations out of this paper
+(TTI)~\cite{zhang2011stable} that takes into account the anisotopic propagation of waves in the earth subsurface (directional dependency of the speed of sound).
+We leave the TTI equations out of this paper
 for brevity.
 
-The solution to equation~\ref{eqn:wave} forms the forward problem. The adjoint
-problem minimizes the misfit between simulated and observed signal given by
+The solution to equation~\ref{eqn:wave} forms the forward problem. The seismic inversion problem minimizes the misfit between simulated and observed signal given by:
 \begin{equation}
@@
 \end{equation}
-This optimization problem is usually solved using gradient-descent methods,
-where the gradient is computed using the adjoint-state method, which involves
+
+This optimization problem is usually solved using gradient based methods such as steepest descent,
+where the gradient is computed using the adjoint-state method that involves
 the data-flow pattern from figure \ref{fig:dataflow}.
 
-The values of $m(x)$ used in this work are derived from the SEAM Overthrust
+The values of $m(x)$ used in this work are derived from the Overthrust
 model~\cite{aminzadeh1996three} over a grid of $287 \times 881 \times 881$
 points, including an absorbing layer of 40 points on each side. The grid spacing
-is 25m in space. This is run for 2500 timesteps. The wave field
+is $25m$ in space. The propagation time is $4sec$ that corresponds to 2500 timesteps. The wave field
 at the final time is shown in Figure~\ref{fig:uncompressed}, and only this field
@@
   compression and decompression. This field was formed after a Ricker
-  wavelet source was placed at TODO and the wave propagated for 2500
-  timesteps. This is a vertical cross-section of a 3D field, taken at
-  the source location}
+  wavelet source was placed at the surface of the model and the wave propagated for 2500
+  timesteps. This is a vertical (x-z) cross-section of a 3D field, taken at
+  the $y$ source location}
 \label{fig:uncompressed}
@@
 would be helpful in a particular scenario without actually building a
-complete implementation first. We have tried to keep it application
+complete implementation first. We have tried to keep the algorithm application
 and platform agnostic in order to maximize its applicability. However
@@
 assumption might have been completely wrong if made about a checkpoint
-in the first few time steps since we consistently saw upto 800x
+in the first few time steps since we consistently saw up to 800x
 compression in those cases - since they are mostly zeros. In order for

### 2018-10-05 paper/paper.tex
Undoing changes to conclusion

@@
 would be helpful in a particular scenario without actually building a
-complete implementation first. We have tried to keep the algorithm application
+complete implementation first. We have tried to keep it application
 and platform agnostic in order to maximize its applicability. However
@@
 assumption might have been completely wrong if made about a checkpoint
-in the first few time steps since we consistently saw up to 800x
+in the first few time steps since we consistently saw upto 800x
 compression in those cases - since they are mostly zeros. In order for

### 2018-10-05 paper/paper.tex
Update plots
removed dept names

@@
 \author{\IEEEauthorblockN{Navjot Kukreja}
-\IEEEauthorblockA{\textit{Department of Earth Science and Engineering} \\
+\IEEEauthorblockA{%\textit{Department of Earth Science and Engineering} \\
 \textit{Imperial College London}\\
@@
 \IEEEauthorblockN{Jan H\"uckelheim}
-\IEEEauthorblockA{\textit{Department of Earth Science and Engineering}\\
+\IEEEauthorblockA{%\textit{Department of Earth Science and Engineering}\\
 \textit{Imperial College London}\\
@@
 \IEEEauthorblockN{Mathias Louboutin}
-\IEEEauthorblockA{\textit{School of Computational Science and Engineering} \\
+\IEEEauthorblockA{%\textit{School of Computational Science and Engineering} \\
 \textit{Georgia Institute of Technology}\\
@@
 \IEEEauthorblockN{Kaiyuan Hou}
-\IEEEauthorblockA{\textit{Department of Electrical Engineering and Computer Science} \\
+\IEEEauthorblockA{%\textit{Department of Electrical Engineering and Computer Science} \\
 \textit{Northwestern University}\\
@@
 \IEEEauthorblockN{Fabio Luporini}
-\IEEEauthorblockA{\textit{Department of Earth Science and Engineering} \\
+\IEEEauthorblockA{%\textit{Department of Earth Science and Engineering} \\
 \textit{Imperial College London}\\
@@
 \IEEEauthorblockN{Paul Hovland}
-\IEEEauthorblockA{\textit{Mathematics and Computer Science Division} \\
+\IEEEauthorblockA{%\textit{Mathematics and Computer Science Division} \\
 \textit{Argonne National Laboratory}\\
@@
 \IEEEauthorblockN{Gerard Gorman}
-\IEEEauthorblockA{\textit{Department of Earth Science and Engineering}\\
+\IEEEauthorblockA{%\textit{Department of Earth Science and Engineering}\\
 \textit{Imperial College London}\\

### 2018-10-05 paper/paper.tex
fixed reference
Finishing up

@@
 
-Another way of using compression to speed up seismic inversion or other adjoint-based methods is to use it \emph{instead} of checkpointing.
-If the compression ratio is sufficient to fit the entire data in memory, checkpointing is no longer necessary.
-Previous work has discussed this situation in the context of computational fluid dynamics~\cite{cyr2015towards} and seismic inversion using compression method specifically designed for
+Another way of using compression to speed up adjoint-based methods is to use it \emph{instead} of
+checkpointing. If the compression ratio is sufficient to fit the
+entire data in memory, checkpointing is no longer necessary. Previous
+work has discussed this in the context of computational
+fluid dynamics~\cite{cyr2015towards} and seismic inversion using
+compression algorithms specifically designed for
 wavefields~\cite{dalmau2014lossy,boehm2016wavefield}.
@@
 
-Compared to the use of only checkpointing without compression, our combined method often improves performance.
-This is a consequence of the reduced size of stored program states, allowing more program states to be stored during the forward computation.
-This in turn reduces the amount of recomputation that needs to be performed.
-On the other hand, the compression and decompression itself takes time.
-We therefore provide a comprehensive performance model that predicts whether the combined method is beneficial, that is, whether the time spent compressing and decompressing is less than the time saved by reduced
-recomputations.
+Compared to the use of only checkpointing without compression, our
+combined method often improves performance. This is a consequence of
+the reduced size of stored program states, allowing more program
+states to be stored during the forward computation. This in turn
+reduces the amount of recomputation that needs to be performed. On the
+other hand, the compression and decompression itself takes time. We
+therefore provide a comprehensive performance model that predicts
+whether the combined method is beneficial, that is, whether the time
+spent compressing and decompressing is less than the time saved by
+reduced recomputations.
 
-We present test cases that use a TODO describe Richter
-and TODO describe JLSE/SKX node system, henceforth called Richter and Skylake,
-respectively. We discuss performance results for varying error tolerances for
-the lossy ZFP compression algorithm ranging from TODO to TODO, and different
+For benchmarking we used a dual-socket
+Intel(R) Xeon(R) Platinum 8180M @ 2.50 Ghz (28 cores each), henceforth
+called Skylake. We discuss performance results for varying error tolerances for
+the lossy ZFP compression algorithm ranging from $10^0$ to $10^{-15}$, and different
 types of forward and adjoint computations that vary in their ratio between
@@
 dependent speed of sound, $u(t, x)$ is the pressure wavefield, $\Laplace u(t, x)$ denotes the laplacian of the wavefield and $q(t, x)$ is a source term.
-Some of our experiments in Section TODO are performed using a more accurate
+Some of our experiments in Section \ref{sec:results} are performed using a more accurate
 and more complex version of this equation called Tilted Transverse Isotropy
@@
 
-The highest compression efficiency TODO define this is achieved in the
-fixed-tolerance mode albeit with highly unpredictable compression ratios. Figure
-\ref{fig:decompressed_error} shows the spatial distribution of the error after
-compression and decompression, compared to the original field, using
-fixed-tolerance mode.
+The compression ratio achieved per unit time spent during compressing
+and decompressing was seen to be highest in the fixed-tolerance mode
+albeit with highly unpredictable compression ratios. Figure
+\ref{fig:decompressed_error} shows the spatial distribution of the
+error after compression and decompression, compared to the original
+field, using fixed-tolerance mode.
 
 In our experiments using Devito, the time required to compress a checkpoint was
-in the same orderof magnitude as the time taken to compute a timestep.
+in the same order of magnitude as the time taken to compute a timestep.
 
@@
 \begin{center}
-\includegraphics[width=0.8\linewidth]{images/tolerance-cf-richter.pdf}
+\includegraphics[width=0.9\linewidth]{images/tolerance-cf-richter.pdf}
 \end{center}
-\caption{Compression ratios achieved on compressing the wavefield. The
-highlighted point corresponds to the setting used for the other results here
-unless otherwise specified.}
+\caption{Compression ratios achieved on compressing the wavefield. We
+  define compression ratio as the ratio between the size of the
+  uncompressed data and the compressed data. The dashed line
+  represents no compression. The highlighted point corresponds to the
+  setting used for the other results here unless otherwise specified.}
 \label{fig:tolerance_cf_plot}
@@
 \end{equation}
-where $\mathbf{N}_W$ is the total number of times Revolve writes checkpoints for a single run, $ \mathbf{N}_R$ 
-is the number of times checkpoints are read, and $\mathbf{B}$ is the memory bandwidth of the target system.
-
-The total time to solution in this scenario becomes
+where $\mathbf{N}_W$ is the total number of times Revolve writes
+checkpoints for a single run, $ \mathbf{N}_R$ is the number of times
+checkpoints are read, and $\mathbf{B}$ is the memory bandwidth of the
+target system. The total time to solution becomes
 \begin{equation}
@@
 \end{equation} 
-
 By using compression, the size of each checkpoint is reduced and therefore the
@@
 \section{Results}
+\label{sec:results}
 We can distinguish three different scenarios, depending on the amount of
@@
 \begin{center}
-\includegraphics[width=\linewidth]{images/varying-memory.pdf}
+\includegraphics[width=0.9\linewidth]{images/varying-memory.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\linewidth]{images/varying-compute.pdf}
+\includegraphics[width=0.9\linewidth]{images/varying-compute.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\linewidth]{images/varying-nt.pdf}
+\includegraphics[width=0.9\linewidth]{images/varying-nt.pdf}
 \end{center}
@@
 \section{Conclusions and Future work}
-We use lossy compression to reduce the computational overhead of checkpointing
-in an adjoint computation used in full waveform inversion, a common method in
-seismic imaging applications whose memory footprint commonly exceeds the
-available memory size in high performance computing systems. We also developed a
-performance model that computes whether or not the combination of compression
-and checkpointing will outperform pure checkpointing or pure compression in a
-variety of scenarios, depending on the available memory size, computational
-intensity of the application, and compression ratio and throughput of the
-compression algorithm.
-
-Our current result has several limitations that we plan to address in future
-work:
+We use lossy compression to reduce the computational overhead of
+checkpointing in an adjoint computation used in full waveform
+inversion, a common method in seismic imaging applications whose
+memory footprint commonly exceeds the available memory size in high
+performance computing systems. We also developed a performance model
+that computes whether or not the combination of compression and
+checkpointing will outperform pure checkpointing or pure compression
+in a variety of scenarios, depending on the available memory size,
+computational intensity of the application, and compression ratio and
+throughput of the compression algorithm. Our current result has
+several limitations that we plan to address in future work:
 \begin{itemize}
@@
 application.
-\item We only present results using the ZFP compression library. Again, the
-performance model is independent of this, but it would be interesting to try a
-greater variety of compression algorithms.
-\item ZFP currently only supports serial decompression. If ZFP supported
+\item We only present results using the ZFP compression library. It
+would be interesting to try a greater variety of compression
+algorithms.
+\item ZFP only supports serial decompression. If ZFP supported
 parallel decompression, our experiments would likely show a geater region in
 which the combined method is faster than pure checkpointing. Furthermore, ZFP
-only supports fields with up to three dimensions, and it is possible that
-exploiting similarities between fields at different time steps would yield a
+only supports fields with up to three dimensions, while
+exploiting similarities between fields at different time steps may yield a
 better compression ratio.
-\item Our performance model is based on uniform compression ratios and times.
-However, many applications, including FWI, are likely to have initial conditions
-that contain little information and are easily compressed, and the compression
-ratio gradually declines as the field becomes more complex over time. We have
-based our experiments on the final wave field, which is presumably the most
-difficult to compress.
-\item In comparing pure compression with pure checkpointing, we assume that
-every checkpoint would be compressed and decompressed. However, in scenarios
-where the available memory is only slightly less than the required memory, an
-implementation that compresses only a subset of the checkpoints might outperform
-the expectations of our model. 
-\item We do not discuss multi-level checkpointing, where some checkpoints are
-stored on a slower, larger device such as a solid state drive. We expect
-compression to be particularly beneficial in these scenarios due to the reduced
-data transfer sizes.
+\item Our performance model is based on uniform compression ratios and
+times.  However, many applications, including FWI, are likely to have
+initial conditions that contain little information and are easily
+compressed, and the compression ratio gradually declines as the field
+becomes more complex. We based our experiments on the final wave
+field, which is presumably difficult to compress.
+\item In comparing pure compression with pure checkpointing, we assume
+that every checkpoint is compressed and decompressed. However, if the
+available memory is only slightly less than the required memory, an
+implementation that compresses only a subset of the checkpoints might
+outperform the expectations of our model.
+\item We do not discuss multi-level checkpointing, where some
+checkpoints are stored on a slower, larger device. We expect
+compression to be beneficial in these scenarios due to reduced data
+transfer sizes.
 \end{itemize}

### 2018-10-05 paper/paper.tex
Update figure and add attribution

@@
 \begin{center}
-\includegraphics[width=0.8\linewidth]{images/survey-ship-diagram.png}
+\includegraphics[width=0.75\linewidth]{images/survey-ship-diagram.jpg}
 \end{center}
-\caption{Illustration of an offshore seismic survey that collects the input data for seismic inversion}
+\caption{Illustration of an offshore seismic survey that collects the
+  input data for seismic inversion. Source: The Open University \cite{surveyship}}
 \label{fig:offshore_survey}
@@
   corresponding to the wave amplitude (from the field plot in figure
-  \ref{uncompressed}. A small block-like structure characteristic of
+  \ref{fig:uncompressed}. A small block-like structure characteristic of
   ZFP can be seen.}

### 2018-10-11 paper/paper.tex
Modifications corresponding to Paul's suggestions

@@
 efficiently is limited, for example by communication overheads that start to
-dominate as the domain is split into increasingly small pieces~\cite{virieux2009seismic}.
-Secondly, this method is broadly used on conventional clusters but can be drastically more complicated to seup and slower due to the communication in the Cloud.
-Another common strategy in seismic inversion is to only store values at the boundaries of the domain at each timestep, and reconstruct the rest of the wavefield when required~\cite{clapp2009reverse,yang2014rtm} with time reversal of the wave equation.
-However, this method is not applicable for wave equations that are not time reversible when for example physical attenuation is included.
+dominate as the domain is split into increasingly small
+pieces~\cite{virieux2009seismic}. Secondly, this method is broadly
+used on conventional clusters but can be drastically more complicated
+to seup and slower due to the communication in a cloud-based
+setup. 
+Another common strategy in seismic inversion is to only store values
+at the boundaries of the domain at each timestep, and reconstruct the
+rest of the wavefield when
+required~\cite{clapp2009reverse,yang2014rtm} with time reversal of the
+wave equation. However, this method is not applicable for wave
+equations that are not time reversible when for example physical
+attenuation is included.
 
@@
 discarded). The discarded data is recomputed when needed by restarting the
-forward pass from the last available stored state. The question which states
-should be stored and which states should be recomputed to minimize the total
-amount of recomputation work has been answered under certain assumptions with the Revolve algorithm~\cite{griewank2000algorithm}.
-Other authors have subsequently developed extensions to Revolve that are optimal under less restrictive assumptions~\cite{wang2009minimal,aupy2016optimal,schanen2016asynchronous}.
+forward pass from the last available stored state. The Revolve
+algorithm~\cite{griewank2000algorithm} provides an answer to the
+question of which states should be stored and which states should be recomputed to minimize the total
+amount of recomputation work. Other authors have subsequently
+developed extensions to Revolve that are optimal under different
+assumptions~\cite{stumm2009multistage, wang2009minimal,aupy2016optimal,schanen2016asynchronous}.
 
@@
 tradeoff between compression ratio, accuracy, and compression and decompression
-time. In this work we use the ZFP package to perform lossy compression.
+time.
 
@@
 where the gradient is computed using the adjoint-state method that involves
-the data-flow pattern from figure \ref{fig:dataflow}.
+the data-flow pattern from Figure \ref{fig:dataflow}.
 
@@
   evenly distributed across the domain with only slight variations
-  corresponding to the wave amplitude (from the field plot in figure
+  corresponding to the wave amplitude (from the field plot in Figure
   \ref{fig:uncompressed}. A small block-like structure characteristic of
@@
       N(N-1) /2, & \text{if}\ M=1 \\
-      \min\limits_{1<=\widetilde{N}<=N} \{\widetilde{N} + \mathbf{O}_R(\widetilde{N}, M) + \mathbf{O}_R(N-\widetilde{N}, M-1)\}, & \text{if}\ M>1
+      \min\limits_{1<=\widetilde{N}<=N} \{\widetilde{N} + p(\widetilde{N}, M) + p(N-\widetilde{N}, M-1)\}, & \text{if}\ M>1
     \end{cases}
@@
 the results copied back to the checkpointing memory. This copying is a common feature of checkpointing
-implementations, and might pose a non-trivial overhead in scenarios where the computation is heavily bandwidth-bound. 
+implementations, and might pose a non-trivial overhead in some scenarios. 
 This storage overhead is given by
 \begin{equation}
-\mathbf{O}_{SR} = \mathbf{N}_W \cdot \frac{\mathbf{S}}{\mathbf{B}} +
-\mathbf{N}_R \cdot \frac{\mathbf{S}}{\mathbf{B}}.
+\mathbf{O}_{SR}(N, M) = \mathbf{W}(N, M) \cdot \frac{\mathbf{S}}{\mathbf{B}} +
+\mathbf{R}(N, M) \cdot \frac{\mathbf{S}}{\mathbf{B}}
 \label{eqn:storage}
 \end{equation}
-where $\mathbf{N}_W$ is the total number of times Revolve writes
-checkpoints for a single run, $ \mathbf{N}_R$ is the number of times
+where $\mathbf{W}$ is the total number of times Revolve writes
+checkpoints for a single run, $ \mathbf{R}$ is the number of times
 checkpoints are read, and $\mathbf{B}$ is the memory bandwidth of the
@@
 \begin{equation}
-T_R = 2 \cdot \mathbf{C} \cdot \mathbf{N} + \mathbf{O}_R(N, M) + \mathbf{O}_S.
+T_R = 2 \cdot \mathbf{C} \cdot \mathbf{N} + \mathbf{O}_R(N, M) +
+\mathbf{O}_{SR}(N, M)
 \end{equation} 
@@
 To be beneficial, the reduction in $\mathbf{O}_R$ must offset the increase in 
-$\mathbf{O}_S$, leading to an overall decrease in the time to solution $T$.
+$\mathbf{O}_{SR}$, leading to an overall decrease in the time to solution $T$.
 
@@
 \begin{equation}
-\mathbf{O}_{SR} = \mathbf{N}_W \cdot \left(\frac{\mathbf{S}}{\mathbf{F}
-  \cdot \mathbf{B}} + t_c\right) + \mathbf{N}_R \cdot
+\begin{split}
+\mathbf{O}_{SR}(N, M) = &\mathbf{W}(N, M \cdot F) \cdot \left(\frac{\mathbf{S}}{\mathbf{F}
+  \cdot \mathbf{B}} + t_c\right) +\\&\mathbf{R}(N, M \cdot F) \cdot
 \left(\frac{\mathbf{S}}{\mathbf{F} \cdot \mathbf{B}} + t_d\right)
+\end{split}
 \end{equation}
@@
 \end{enumerate}
-All three scenarios can be seen in figure~\ref{fig:varying_memory}.
+All three scenarios can be seen in Figure~\ref{fig:varying_memory}.
 The second scenario was studied in previous work~\cite{cyr2015towards}, while
@@
 compression can substantially reduce this recompute factor. This can be seen in
-figures~\ref{fig:varying_memory} and~\ref{fig:varying_nt}.
+Figures~\ref{fig:varying_memory} and~\ref{fig:varying_nt}.
 
@@
 compression is also likely to be beneficial, as shown in
-figure~\ref{fig:varying_compute}. As the time per time step increases and the
+Figure~\ref{fig:varying_compute}. As the time per time step increases and the
 compression cost becomes negligible, we observe that the ratio between the runtime

### 2019-01-11 lossless.py
Use all compressors

@@ -1,0 +1,48 @@
+import blosc
+import sys
+import h5py
+from bloscpack import pack_nd_array_to_bytes, unpack_ndarray_from_bytes
+from itertools import product
+
+
+chunk_size_options = [int(2**i) for i in numpy.arange(19, 23.5, 0.5)]
+shuffle_options = [blosc.SHUFFLE, blosc.BITSHUFFLE, blosc.NOSHUFFLE]
+cname_options = blosc.cnames
+clevel_options = list(range(10))
+
+
+def compress(arr, chunk_size, shuffle, cname, clevel):
+    if arr.dtype == np.float32:
+        typesize = 4
+    elif arr.dtype == np.float64:
+        typesize == 8
+    else:
+        assert(False)
+
+    return pack_ndarray_to_bytes(arr, blosc_args=BloscArgs(shuffle=shuffle,
+                                                           typesize=typesize,
+                                                           cname=cname,
+                                                           clevel=clevel), chunk_size=chunk_size)
+
+
+def decompress(compressed):
+    return unpack_ndarray_from_bytes(compressed)
+
+
+filename = sys.argv[1]
+f = h5py.File(filename, 'r')
+uncompressed = f['data'][()].astype(np.dtype('float32'))
+
+combinations = product(chunk_size_options, shuffle_options, cname_options, clevel_options)
+
+for op_set in combinations:
+    with Timer(factor=1000) as compress_time:
+        compressed = compress(input_array, *op_set)
+
+    compression_factor = (input_array.size * input_array.item_size) / len(compressed)
+
+    with Timer(factor=1000) as decompress_time:
+        decompressed = decompress(compressed)
+
+    assert(np.allclose(input_array, decompressed))
+    print("%f, %d, %s, %d, %f, %f, %f" % (op_set[0], op_set[1], op_set[2], op_set[3], compress_time, decompress_time, compression_factor))

### 2019-01-11 tolerance.py

@@ -3,15 +3,15 @@ import sys
 from util import to_hdf5
 import numpy as np
 from contexttimer import Timer
-from zfp import compress, decompress
 from argparse import ArgumentParser
 
 
 
 
-description = ("Test for fixed accuracy mode of zfp")
+description = ("Test for fixed accuracy mode of different compressors")
 parser = ArgumentParser(description=description)
 parser.add_argument("filename", type=str, help="Filename")
+parser.add_argument("compressor", choices=["zfp", "sz"], default="zfp")
 p_parser = parser.add_mutually_exclusive_group(required=False)
 p_parser.add_argument('--parallel', dest='parallel', action='store_true')
 p_parser.add_argument('--no-parallel', dest='parallel', action='store_false')
@@ -20,6 +20,12 @@ args = parser.parse_args()
 
 filename = args.filename
 parallel = args.parallel
+compressor = args.compressor
+
+if compressor == "zfp":
+    from pyzfp import compress, decompress
+else:
+    from pysz import compress, decompress
 
 f = h5py.File(filename, 'r')
 uncompressed = f['data'][()].astype(np.dtype('float32'))
@@ -32,6 +38,8 @@ for p_i in range(0, 16):
     with Timer(factor=1000) as t2:
         decompressed = decompress(compressed, uncompressed.shape, uncompressed.dtype, tolerance=tolerance, parallel=parallel)
 
-    to_hdf5(decompressed, "decompressed-t-%d.h5"%p_i)
+    #to_hdf5(decompressed, "decompressed-t-%d.h5"%p_i)
     error_matrix = decompressed-uncompressed
+    if p_i in (0, 8, 16):
+        to_hdf5(error_matrix, "error_field-%s-%d.h5"%(compressor, p_i))
     print("%f, %f, %f, %f, %.16f, %f, %f" % (len(compressed), len(uncompressed.tostring())/float(len(compressed)), t.elapsed, t2.elapsed, tolerance, np.linalg.norm(error_matrix), np.max(error_matrix)))

### 2019-01-12 paper/conclusions-futurework.tex
Update before overleaf

--- 2018-10-11 paper/paper.tex
@@
-\documentclass[conference]{IEEEtran}
-%\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{cuted}
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{url}
-
-\usetikzlibrary{calc,trees,positioning,arrows,chains,shapes.geometric,%
-    decorations.pathreplacing,decorations.pathmorphing,shapes,%
-    matrix,shapes.symbols}
-
-\tikzstyle{line} = [draw, -, thick]
-\tikzstyle{nodraw} = [draw, fill, circle, minimum width=0pt, inner sep=0pt]
-\tikzstyle{box} = [line, rectangle, rounded corners, text centered]
-
-\tikzset{
->=stealth',
-  punktchain/.style={
-    rectangle, 
-    rounded corners, 
-    draw=black, very thick,
-    text width=2em, 
-    minimum height=3em, 
-    text centered, 
-    on chain},
-  line/.style={draw, thick, <-},
-  element/.style={
-    tape,
-    top color=white,
-    bottom color=blue!50!black!60!,
-    minimum width=1em,
-    draw=blue!40!black!90, very thick,
-    text width=2em, 
-    minimum height=3.5em, 
-    text centered, 
-    on chain},
-  every join/.style={->, thick,shorten >=1pt},
-  decoration={brace},
-  tuborg/.style={decorate},
-  tubnode/.style={midway, right=2pt},
-}
-
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\newcommand*\Laplace{\mathop{}\!\mathbin\bigtriangleup}
-\begin{document}
-
-\title{Combining checkpointing and data compression\\ for large scale seismic inversion}
-
-\author{\IEEEauthorblockN{Navjot Kukreja}
-\IEEEauthorblockA{%\textit{Department of Earth Science and Engineering} \\
-\textit{Imperial College London}\\
-London, UK \\
-n.kukreja@imperial.ac.uk}
-\and
-\IEEEauthorblockN{Jan H\"uckelheim}
-\IEEEauthorblockA{%\textit{Department of Earth Science and Engineering}\\
-\textit{Imperial College London}\\
-London, UK }
-\and
-\IEEEauthorblockN{Mathias Louboutin}
-\IEEEauthorblockA{%\textit{School of Computational Science and Engineering} \\
-\textit{Georgia Institute of Technology}\\
-Atlanta, GA, USA\\}
-\and
-\IEEEauthorblockN{Kaiyuan Hou}
-\IEEEauthorblockA{%\textit{Department of Electrical Engineering and Computer Science} \\
-\textit{Northwestern University}\\
-Evanston, IL, USA \\
-}
-\and
-\IEEEauthorblockN{Fabio Luporini}
-\IEEEauthorblockA{%\textit{Department of Earth Science and Engineering} \\
-\textit{Imperial College London}\\
-London, UK \\
-}
-\and
-\IEEEauthorblockN{Paul Hovland}
-\IEEEauthorblockA{%\textit{Mathematics and Computer Science Division} \\
-\textit{Argonne National Laboratory}\\
-Lemont, IL, USA \\
-}
-\and
-\IEEEauthorblockN{Gerard Gorman}
-\IEEEauthorblockA{%\textit{Department of Earth Science and Engineering}\\
-\textit{Imperial College London}\\
-London, UK }
-}
-
-\maketitle
-
-\begin{abstract}
-Seismic inversion and imaging are adjoint-based optimization problems that processes up to terabytes of data, regularly exceeding the memory
-capacity of available computers. Data compression is an effective strategy to
-reduce this memory requirement by a certain factor, particularly if some loss in
-accuracy is acceptable. A popular alternative is checkpointing, where data is
-stored at selected points in time, and values at other times are recomputed as
-needed from the last stored state.  This allows arbitrarily large adjoint
-computations with limited memory, at the cost of additional recomputations.
-
-In this paper we combine compression and checkpointing for the first
-time to compute a realistic seismic inversion. The combination of
-checkpointing and compression allows
-larger adjoint computations compared to using only compression, and
-reduces the recomputation overhead significantly compared to using only checkpointing.
-\end{abstract}
-
-\begin{IEEEkeywords}
-Checkpointing, compression, adjoints, inversion
-\end{IEEEkeywords}
-
-\section{Introduction}
-Adjoint-based optimization problems typically consist of a simulation that is
-run forward in simulation time, producing data that is used in reverse order by
-a subsequent adjoint computation that is run backwards in simulation time.
-Figure~\ref{fig:dataflow} shows the resulting data flow. Many important
-applications in science and engineering have this structure, including
-seismic inversion.
-
-In seismic inversion, the propagation of seismic waves through the earth's subsurface is simulated and compared with data from field measurements.
-The model of the subsurface is iteratively improved by minimizing the misfit between simulated data and field measurement in an adjoint optimization problem.
-Figure~\ref{fig:offshore_survey} shows the setup of the field experiment that produces the measured data~\cite{plessix2006review}. Storing all intermediate
-results as required for the data flow shown in Figure~\ref{fig:dataflow} is not
-feasible, since this would require tens of terabytes of memory, which is not
-commonly available on the machines used to solve these problems. To solve this
-problem, a number of strategies are used that are discussed in the following
-paragraphs. While this paper uses seismic inversion as a test case, most of these ideas, including the one presented in this work, are also applicable to other adjoint optimization problems.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.75\linewidth]{images/survey-ship-diagram.jpg}
-\end{center}
-\caption{Illustration of an offshore seismic survey that collects the
-  input data for seismic inversion. Source: The Open University \cite{surveyship}}
-\label{fig:offshore_survey}
-\end{figure}
-
-Domain decomposition is often used not only to distribute the computational
-workload across more processors, but also to utilize the large amount of memory
-available in distributed systems. While this strategy is very powerful, the
-number of compute nodes and therefore the amount of memory that can be used
-efficiently is limited, for example by communication overheads that start to
-dominate as the domain is split into increasingly small
-pieces~\cite{virieux2009seismic}. Secondly, this method is broadly
-used on conventional clusters but can be drastically more complicated
-to seup and slower due to the communication in a cloud-based
-setup. 
-Another common strategy in seismic inversion is to only store values
-at the boundaries of the domain at each timestep, and reconstruct the
-rest of the wavefield when
-required~\cite{clapp2009reverse,yang2014rtm} with time reversal of the
-wave equation. However, this method is not applicable for wave
-equations that are not time reversible when for example physical
-attenuation is included.
-
-Checkpointing is yet another strategy to reduce the memory overhead. Only a
-subset of the program states during the forward pass is stored (and the rest
-discarded). The discarded data is recomputed when needed by restarting the
-forward pass from the last available stored state. The Revolve
-algorithm~\cite{griewank2000algorithm} provides an answer to the
-question of which states should be stored and which states should be recomputed to minimize the total
-amount of recomputation work. Other authors have subsequently
-developed extensions to Revolve that are optimal under different
-assumptions~\cite{stumm2009multistage, wang2009minimal,aupy2016optimal,schanen2016asynchronous}.
-
-\begin{figure}
-\begin{center}
-\input{dataflow}
-\end{center}
-\caption{The dataflow pattern that is typical of adjoint-based optimization problems}
-\label{fig:dataflow}
-\end{figure}
-
-Data compression is also increasingly used to reduce the memory footprint of
-scientific applications. General purpose data compression algorithms like Zlib
-(which is a part of gzip)~\cite{deutsch1996zlib}, and compression algorithms for
-video and image data such as JPEG-2000~\cite{skodras2001jpeg} have been
-presented in previous work. More recently, special purpose compression
-algorithms for floating-point scientific data have been developed, such as ZFP
-or SZ~\cite{Kaklamanis:2012aa,lindstrom2014fixed,di2018efficient}.
-
-Lossless algorithms guarantee that the exact original data can be recovered
-during decompression, whereas lossy algorithms introduce an error, but often
-guarantee that the error does not exceed certain absolute or relative error
-metrics. Typically, lossy compression is more effective in reducing the data
-size. Most popular compression packages offer various settings that allow a
-tradeoff between compression ratio, accuracy, and compression and decompression
-time.
-
-It is worth noting that another data reduction strategy is to typecast values
-into a lower precision format, for example, from double precision to single
-precision. This can be seen as a computationally cheap lossy compression
-algorithm with a compression ratio of $2$.
-
-Perhaps counterintuitively, compression can not only reduce the memory
-footprint, but also speed up an application. Previous work has observed that the
-compression and decompression time can be less than the time saved from the
-reduction in data that needs to be communicated across MPI nodes or between a
-GPU and a host computer~\cite{gpu-compression}.
-
-Another way of using compression to speed up adjoint-based methods is to use it \emph{instead} of
-checkpointing. If the compression ratio is sufficient to fit the
-entire data in memory, checkpointing is no longer necessary. Previous
-work has discussed this in the context of computational
-fluid dynamics~\cite{cyr2015towards} and seismic inversion using
-compression algorithms specifically designed for
-wavefields~\cite{dalmau2014lossy,boehm2016wavefield}.
-
-In this paper, we extend the previous studies by \emph{combining} checkpointing
-and compression. This is obviously useful when the data does not fit in the
-available memory even after compression, for example for very large adjoint
-problems, or for problems where the required accuracy limits the achievable
-compression ratios.
-
-Compared to the use of only checkpointing without compression, our
-combined method often improves performance. This is a consequence of
-the reduced size of stored program states, allowing more program
-states to be stored during the forward computation. This in turn
-reduces the amount of recomputation that needs to be performed. On the
-other hand, the compression and decompression itself takes time. We
-therefore provide a comprehensive performance model that predicts
-whether the combined method is beneficial, that is, whether the time
-spent compressing and decompressing is less than the time saved by
-reduced recomputations.
-
-For benchmarking we used a dual-socket
-Intel(R) Xeon(R) Platinum 8180M @ 2.50 Ghz (28 cores each), henceforth
-called Skylake. We discuss performance results for varying error tolerances for
-the lossy ZFP compression algorithm ranging from $10^0$ to $10^{-15}$, and different
-types of forward and adjoint computations that vary in their ratio between
-compute cost and state size.
-
-Our aim is to make this discussion independent of the hardware, application and
-compression algorithm. We therefore do not discuss whether or not the accuracy
-of the decompressed data is sufficient for our application, or whether
-other algorithms might achieve a better accuracy/compression tradeoff.
-
-\section{Test case using Devito and pyRevolve}
-We use Devito~\cite{devito-api,devito-compiler} to solve forward and adjoint
-wave equation problems. Devito is a domain-specific language that enables the
-rapid development of finite-difference solvers from a high-level description
-of partial differential equations. The simplest version of the seismic wave
-equation is the acoustic isotropic wave equation defined as:
-\begin{equation}
-m(x)\frac{\partial^2 u(t, x)}{\partial t^2} - \Laplace u(t, x) = q(t, x),
-\label{eqn:wave}
-\end{equation}
-where $m(x) = \frac{1}{c^2(x)}$ is the squared slowness, $c(x)$ the spatially
-dependent speed of sound, $u(t, x)$ is the pressure wavefield, $\Laplace u(t, x)$ denotes the laplacian of the wavefield and $q(t, x)$ is a source term.
-Some of our experiments in Section \ref{sec:results} are performed using a more accurate
-and more complex version of this equation called Tilted Transverse Isotropy
-(TTI)~\cite{zhang2011stable} that takes into account the anisotopic propagation of waves in the earth subsurface (directional dependency of the speed of sound).
-We leave the TTI equations out of this paper
-for brevity.
-
-The solution to equation~\ref{eqn:wave} forms the forward problem. The seismic inversion problem minimizes the misfit between simulated and observed signal given by:
-\begin{equation}
-\min_{m} \phi_s(m) = \frac{1}{2} \left\lVert d_{sim} - d_{obs} \right\rVert_2^2.
-\end{equation}
-
-This optimization problem is usually solved using gradient based methods such as steepest descent,
-where the gradient is computed using the adjoint-state method that involves
-the data-flow pattern from Figure \ref{fig:dataflow}.
-
-The values of $m(x)$ used in this work are derived from the Overthrust
-model~\cite{aminzadeh1996three} over a grid of $287 \times 881 \times 881$
-points, including an absorbing layer of 40 points on each side. The grid spacing
-is $25m$ in space. The propagation time is $4sec$ that corresponds to 2500 timesteps. The wave field
-at the final time is shown in Figure~\ref{fig:uncompressed}, and only this field
-was used for the compression experiments in this paper. The uncompressed size of
-this single time step field is just under 900MB.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.8\linewidth]{images/uncompressed.pdf}
-\end{center}
-\caption{Cross-section of the wavefield used as a reference sample for
-  compression and decompression. This field was formed after a Ricker
-  wavelet source was placed at the surface of the model and the wave propagated for 2500
-  timesteps. This is a vertical (x-z) cross-section of a 3D field, taken at
-  the $y$ source location}
-\label{fig:uncompressed}
-\end{figure}
-
-To implement Revolve with Devito, we use pyRevolve~\cite{kukreja2018high} which
-is a python wrapper for the Revolve algorithm. The performance model in
-section~\ref{sec:performance_model} assumes that the implementation is similar
-to pyRevolve, which stores a checkpoint by copying a portion of the operator's
-working memory to the checkpointing memory and similarly loads a checkpoint by
-copying from the checkpointing memory to the operator's working memory. Although
-a perfect implementation of checkpointing may be able to avoid these copies, the
-overhead attached to these copies can be ignored for an operator that is
-sufficiently computationally intensive. However, we include the overheads in the
-model to verify this assumption.
-
-\section{Compression algorithms}
-\subsection{Lossless}
-We use the python package \emph{blosc}~\cite{blosc}, which includes implementations for
-six different lossless compression algorithms, namely ZLIB, ZSTD, BLOSCLZ,
-LZ4, LZ4HC and Snappy. An exhaustive search over all available parameters 
-resulted in a maximum compression ratio of 1.18x. We therefore focus on lossy
-compression in the remainder of this work.
-
-\subsection{Lossy}
-We use the lossy compression package ZFP~\cite{lindstrom2014fixed} developed in
-C. To use ZFP from python, we developed a python wrapper for the reference
-implementation of ZFP.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.8\linewidth]{images/errors.pdf}
-\end{center}
-\caption{Cross-section of the field that shows errors introduced
-  during compression and decompression using the fixed-tolerance
-  mode. It is interesting to note that the errors are more or less
-  evenly distributed across the domain with only slight variations
-  corresponding to the wave amplitude (from the field plot in Figure
-  \ref{fig:uncompressed}. A small block-like structure characteristic of
-  ZFP can be seen.}
-\label{fig:decompressed_error}
-\end{figure}
-
-ZFP supports three compression modes, namely fixed-tolerance, fixed-precision
-and fixed-rate. The fixed-tolerance mode limits the absolute error, while the
-fixed-precision mode limits the relative error to some user-specified value.
-The fixed-rate mode achieves a guaranteed compression ratio requested by the
-user, but does not provide any bounds on accuracy loss.
-Figure~\ref{fig:tolerance_cf_plot} shows compression ratios for a variety of
-settings for the three modes.
-
-The compression ratio achieved per unit time spent during compressing
-and decompressing was seen to be highest in the fixed-tolerance mode
-albeit with highly unpredictable compression ratios. Figure
-\ref{fig:decompressed_error} shows the spatial distribution of the
-error after compression and decompression, compared to the original
-field, using fixed-tolerance mode.
-
-In our experiments using Devito, the time required to compress a checkpoint was
-in the same order of magnitude as the time taken to compute a timestep.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.9\linewidth]{images/tolerance-cf-richter.pdf}
-\end{center}
-\caption{Compression ratios achieved on compressing the wavefield. We
-  define compression ratio as the ratio between the size of the
-  uncompressed data and the compressed data. The dashed line
-  represents no compression. The highlighted point corresponds to the
-  setting used for the other results here unless otherwise specified.}
-\label{fig:tolerance_cf_plot}
-\end{figure}
-
-\section{Performance model for a combination of Revolve and compression}
-\label{sec:performance_model}
-We assume that the computation of a single forward time step takes the same wall
-time as the computation of a single reverse time step. We denote this time as
-$\mathbf{C}$. For a simulation with $\mathbf{N}$ timesteps, the minimum wall time required
-for the full forward-adjoint evaluation is given by
-\begin{equation}
-T_N = 2 \cdot \mathbf{C} \cdot \mathbf{N}.
-\end{equation}
-If the size of a single timestep in memory is given by $\mathbf{S}$, this
-requires a memory of at least size $\mathbf{S} \cdot \mathbf{N}$. If sufficient memory
-is available, no checkpointing or compression is needed.
-
-If the memory is smaller than $\mathbf{S} \cdot \mathbf{N}$, Revolve provides
-a strategy to solve for the adjoint field by storing a subset of the $\mathbf{N}$ total checkpoints
-and recomputing the remaining ones. The overhead introduced by this method can be broken down into
-the recomputation overhead $\mathbf{O}_R$ and the storage overhead $\mathbf{O}_S$. The recomputation
-overhead is the amount of time spent in recomputation, given by
-\begin{equation}
-\mathbf{O}_R(N, M) = p(N, M) \cdot \mathbf{C},
-\end{equation}
-where $p(N, M)$ is the minimum number of recomputed steps from \cite{griewank2000algorithm}, reproduced
-here in equation \ref{eqn:recompute}.
-\begin{figure*}
-\begin{equation}
-p(N, M) = \begin{cases}
-      N(N-1) /2, & \text{if}\ M=1 \\
-      \min\limits_{1<=\widetilde{N}<=N} \{\widetilde{N} + p(\widetilde{N}, M) + p(N-\widetilde{N}, M-1)\}, & \text{if}\ M>1
-    \end{cases}
-    \label{eqn:recompute}
-\end{equation}
-\end{figure*}
-In equation \ref{eqn:recompute}, M is the number of checkpoints that can be
-stored in memory. Note that for $M >=N$, $\mathbf{O}_R$ would be zero. For $M <
-N$, $\mathbf{O}_R$ grows rapidly as M is reduced relative to N.
-
-In an ideal implementation, the storage overhead $\mathbf{O}_S$ might be zero, since the computation could
-be done ``in-place'', but in practice, checkpoints are generally stored in a separate section of memory and they
-need to be transferred to a ``computational'' section of the memory where the compute is performed, and then
-the results copied back to the checkpointing memory. This copying is a common feature of checkpointing
-implementations, and might pose a non-trivial overhead in some scenarios. 
-This storage overhead is given by
-\begin{equation}
-\mathbf{O}_{SR}(N, M) = \mathbf{W}(N, M) \cdot \frac{\mathbf{S}}{\mathbf{B}} +
-\mathbf{R}(N, M) \cdot \frac{\mathbf{S}}{\mathbf{B}}
-\label{eqn:storage}
-\end{equation}
-where $\mathbf{W}$ is the total number of times Revolve writes
-checkpoints for a single run, $ \mathbf{R}$ is the number of times
-checkpoints are read, and $\mathbf{B}$ is the memory bandwidth of the
-target system. The total time to solution becomes
-\begin{equation}
-T_R = 2 \cdot \mathbf{C} \cdot \mathbf{N} + \mathbf{O}_R(N, M) +
-\mathbf{O}_{SR}(N, M)
-\end{equation} 
-By using compression, the size of each checkpoint is reduced and therefore the
-number of checkpoints available is increased ($M$ in equation
-\ref{eqn:recompute}). This reduces the recomputation overhead $\mathbf{O}_R$,
-while at the same time adding overheads related to compression and decompression
-in $\mathbf{O}_S$.
-To be beneficial, the reduction in $\mathbf{O}_R$ must offset the increase in 
-$\mathbf{O}_{SR}$, leading to an overall decrease in the time to solution $T$.
-
-Our performance model assumes
- that the compression algorithm behaves uniformly
-across the different time steps of the simulation, i.e. that we get the same compression ratio, compression time and 
-decompression time, no matter which of the $N$ possible checkpoints we try to compress/decompress. The storage overhead
-now becomes
-\begin{equation}
-\begin{split}
-\mathbf{O}_{SR}(N, M) = &\mathbf{W}(N, M \cdot F) \cdot \left(\frac{\mathbf{S}}{\mathbf{F}
-  \cdot \mathbf{B}} + t_c\right) +\\&\mathbf{R}(N, M \cdot F) \cdot
-\left(\frac{\mathbf{S}}{\mathbf{F} \cdot \mathbf{B}} + t_d\right)
-\end{split}
-\end{equation}
-where $\mathbf{F}$ is the compression ratio (i.e. the ratio between the uncompressed and compressed checkpoint), and $t_c$
-and $t_d$ are compression and decompression times, respectively. At the same
-time, the recomputation overhead decreases
-because $\mathbf{F}$ times more checkpoints are now available.
-
-\section{Results}
-\label{sec:results}
-We can distinguish three different scenarios, depending on the amount of
-available memory.
-\begin{enumerate}
-\item If the memory is insufficient even with compression to store the entire
-trajectory, one can either use checkpointing only, or combine checkpointing with
-compression.
-\item If the available memory is not sufficient to store the uncompressed
-trajectory, but large enough to store the entire compressed trajectory, we study
-the two possible strategies: Either use compression only, or use checkpointing
-only.
-\item If the available system memory is large enough to hold the entire
-uncompressed trajectory, neither compression nor checkpointing is necessary.
-\end{enumerate}
-All three scenarios can be seen in Figure~\ref{fig:varying_memory}.
-The second scenario was studied in previous work~\cite{cyr2015towards}, while
-the combined method is also applicable to the first scenario, for which previous
-work has only used checkpointing without compression.
-
-We can identify a number of factors that make compression more likely to be
-beneficial compared to pure checkpointing: A very small system memory size and a
-large number of time steps lead to a rapidly increasing recompute factor, and
-compression can substantially reduce this recompute factor. This can be seen in
-Figures~\ref{fig:varying_memory} and~\ref{fig:varying_nt}.
-
-The extent to which the recompute factor affects the overall runtime also
-depends on the cost to compute each individual time step. If the compute cost
-per time step is large compared to the compression and decompression cost, then
-compression is also likely to be beneficial, as shown in
-Figure~\ref{fig:varying_compute}. As the time per time step increases and the
-compression cost becomes negligible, we observe that the ratio between the runtime
-of the combined method and that of pure checkpointing is only determined by the
-difference in recompute factors.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.9\linewidth]{images/varying-memory.pdf}
-\end{center}
-\caption{The speedups predicted by the performance model for varying
-  memory. The baseline
-(1.0) is the performance of a Revolve-only implementation under the
-same conditions. The different curves represent kernels with differing
-compute times (represented here as a factor of the sum of compression
-and decompression times). The first vertical line at $~$ 53GB marks the
-spot where the compressed wavefield can completely fit in memory and
-Revolve is unnecessary if using compression. The second vertical line
-at $~$ 2.2 TB marks the spot where the entire uncompressed wavefield can
-fit in memory and neither Revolve nor compression is necessary. The
-region to the right is where these optimisations are not necessary or
-relevant. The middle region has been the subject of past studies using
-compression in adjoint problems. The region to the left is the focus
-of this paper.}
-\label{fig:varying_memory}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.9\linewidth]{images/varying-compute.pdf}
-\end{center}
-\caption{The speedups predicted by the performance model for varying
-  compute cost. The baseline
-(1.0) is the performance of a Revolve-only implementation under the
-same conditions. The benefits of compression drop rapidly if the
-computational cost of the kernel that generated the data is much lower
-than the cost of compressing the data. For increasing computational
-costs, the benefits are bounded.}
-\label{fig:varying_compute}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.9\linewidth]{images/varying-nt.pdf}
-\end{center}
-\caption{The speedups predicted by the performance model for varying
-  number of timesteps to be reversed. The baseline
-(1.0) is the performance of a Revolve-only implementation under the
-same conditions. It can be seen that compression becomes more
-beneficial as the number of timesteps is increased.}
-\label{fig:varying_nt}
-\end{figure}
-
 \section{Conclusions and Future work}
 We use lossy compression to reduce the computational overhead of
-checkpointing in an adjoint computation used in full waveform
+checkpointing in an adjoint computation used in seismic
 inversion, a common method in seismic imaging applications whose
@@
 what accuracy is needed and thus what compression ratio is realistic for their
-application.
-\item We only present results using the ZFP compression library. It
-would be interesting to try a greater variety of compression
-algorithms.
+application. TODO this is partiall addressed now. Talk about
+convergence guarantees instead. 
 \item ZFP only supports serial decompression. If ZFP supported
@@
 exploiting similarities between fields at different time steps may yield a
-better compression ratio.
+better compression ratio. TODO update to make more relevant.
 \item Our performance model is based on uniform compression ratios and
-times.  However, many applications, including FWI, are likely to have
+times.  However, many applications, including seismic inversion, are likely to have
 initial conditions that contain little information and are easily
@@
 transfer sizes.
+\item TODO scheduling
 \end{itemize}
-
-\section*{Acknowledgment}
-This work was funded by the Intel Parallel Computing Centre at
-Imperial College London and EPSRC EP/R029423/1. 
-This work was supported by the U.S. Department of Energy, Office of Science,
-Office of Advanced Scientific Computing Research, Applied Mathematics and
-Computer Science programs under contract number DE-AC02-06CH11357.
-We would also like to acknowledge the support from the SINBAD II project and
-the member organizations of the SINBAD Consortium.
-
-We gratefully acknowledge the computing resources provided and operated by the
-Joint Laboratory for System Evaluation (JLSE) at Argonne National Laboratory.
-
-\bibliographystyle{plain}
-\bibliography{compression}
-
-\vfill
-\begin{flushright}
-\normalsize
-\framebox{\parbox{3in}{The submitted manuscript has been created by UChicago
-Argonne, LLC, Operator of Argonne National Laboratory (`Argonne'). Argonne, a
-U.S. Department of Energy Office of Science laboratory, is operated under
-Contract No. DE-AC02-06CH11357. The U.S. Government retains for itself, and
-others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide
-license in said article to reproduce, prepare derivative works, distribute
-copies to the public, and perform publicly and display publicly, by or on behalf
-of the Government.  The Department of Energy will provide public access to these
-results of federally sponsored research in accordance with the DOE Public Access
-Plan.\newline \url{http://energy.gov/downloads/doe-public-access-plan}.}}
-\end{flushright}
-
-\end{document}

### 2019-01-12 paper/devito.tex
Update before overleaf

--- 2018-10-11 paper/paper.tex
@@
-\documentclass[conference]{IEEEtran}
-%\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{cuted}
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{url}
-
-\usetikzlibrary{calc,trees,positioning,arrows,chains,shapes.geometric,%
-    decorations.pathreplacing,decorations.pathmorphing,shapes,%
-    matrix,shapes.symbols}
-
-\tikzstyle{line} = [draw, -, thick]
-\tikzstyle{nodraw} = [draw, fill, circle, minimum width=0pt, inner sep=0pt]
-\tikzstyle{box} = [line, rectangle, rounded corners, text centered]
-
-\tikzset{
->=stealth',
-  punktchain/.style={
-    rectangle, 
-    rounded corners, 
-    draw=black, very thick,
-    text width=2em, 
-    minimum height=3em, 
-    text centered, 
-    on chain},
-  line/.style={draw, thick, <-},
-  element/.style={
-    tape,
-    top color=white,
-    bottom color=blue!50!black!60!,
-    minimum width=1em,
-    draw=blue!40!black!90, very thick,
-    text width=2em, 
-    minimum height=3.5em, 
-    text centered, 
-    on chain},
-  every join/.style={->, thick,shorten >=1pt},
-  decoration={brace},
-  tuborg/.style={decorate},
-  tubnode/.style={midway, right=2pt},
-}
-
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\newcommand*\Laplace{\mathop{}\!\mathbin\bigtriangleup}
-\begin{document}
-
-\title{Combining checkpointing and data compression\\ for large scale seismic inversion}
-
-\author{\IEEEauthorblockN{Navjot Kukreja}
-\IEEEauthorblockA{%\textit{Department of Earth Science and Engineering} \\
-\textit{Imperial College London}\\
-London, UK \\
-n.kukreja@imperial.ac.uk}
-\and
-\IEEEauthorblockN{Jan H\"uckelheim}
-\IEEEauthorblockA{%\textit{Department of Earth Science and Engineering}\\
-\textit{Imperial College London}\\
-London, UK }
-\and
-\IEEEauthorblockN{Mathias Louboutin}
-\IEEEauthorblockA{%\textit{School of Computational Science and Engineering} \\
-\textit{Georgia Institute of Technology}\\
-Atlanta, GA, USA\\}
-\and
-\IEEEauthorblockN{Kaiyuan Hou}
-\IEEEauthorblockA{%\textit{Department of Electrical Engineering and Computer Science} \\
-\textit{Northwestern University}\\
-Evanston, IL, USA \\
-}
-\and
-\IEEEauthorblockN{Fabio Luporini}
-\IEEEauthorblockA{%\textit{Department of Earth Science and Engineering} \\
-\textit{Imperial College London}\\
-London, UK \\
-}
-\and
-\IEEEauthorblockN{Paul Hovland}
-\IEEEauthorblockA{%\textit{Mathematics and Computer Science Division} \\
-\textit{Argonne National Laboratory}\\
-Lemont, IL, USA \\
-}
-\and
-\IEEEauthorblockN{Gerard Gorman}
-\IEEEauthorblockA{%\textit{Department of Earth Science and Engineering}\\
-\textit{Imperial College London}\\
-London, UK }
-}
-
-\maketitle
-
-\begin{abstract}
-Seismic inversion and imaging are adjoint-based optimization problems that processes up to terabytes of data, regularly exceeding the memory
-capacity of available computers. Data compression is an effective strategy to
-reduce this memory requirement by a certain factor, particularly if some loss in
-accuracy is acceptable. A popular alternative is checkpointing, where data is
-stored at selected points in time, and values at other times are recomputed as
-needed from the last stored state.  This allows arbitrarily large adjoint
-computations with limited memory, at the cost of additional recomputations.
-
-In this paper we combine compression and checkpointing for the first
-time to compute a realistic seismic inversion. The combination of
-checkpointing and compression allows
-larger adjoint computations compared to using only compression, and
-reduces the recomputation overhead significantly compared to using only checkpointing.
-\end{abstract}
-
-\begin{IEEEkeywords}
-Checkpointing, compression, adjoints, inversion
-\end{IEEEkeywords}
-
-\section{Introduction}
-Adjoint-based optimization problems typically consist of a simulation that is
-run forward in simulation time, producing data that is used in reverse order by
-a subsequent adjoint computation that is run backwards in simulation time.
-Figure~\ref{fig:dataflow} shows the resulting data flow. Many important
-applications in science and engineering have this structure, including
-seismic inversion.
-
-In seismic inversion, the propagation of seismic waves through the earth's subsurface is simulated and compared with data from field measurements.
-The model of the subsurface is iteratively improved by minimizing the misfit between simulated data and field measurement in an adjoint optimization problem.
-Figure~\ref{fig:offshore_survey} shows the setup of the field experiment that produces the measured data~\cite{plessix2006review}. Storing all intermediate
-results as required for the data flow shown in Figure~\ref{fig:dataflow} is not
-feasible, since this would require tens of terabytes of memory, which is not
-commonly available on the machines used to solve these problems. To solve this
-problem, a number of strategies are used that are discussed in the following
-paragraphs. While this paper uses seismic inversion as a test case, most of these ideas, including the one presented in this work, are also applicable to other adjoint optimization problems.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.75\linewidth]{images/survey-ship-diagram.jpg}
-\end{center}
-\caption{Illustration of an offshore seismic survey that collects the
-  input data for seismic inversion. Source: The Open University \cite{surveyship}}
-\label{fig:offshore_survey}
-\end{figure}
-
-Domain decomposition is often used not only to distribute the computational
-workload across more processors, but also to utilize the large amount of memory
-available in distributed systems. While this strategy is very powerful, the
-number of compute nodes and therefore the amount of memory that can be used
-efficiently is limited, for example by communication overheads that start to
-dominate as the domain is split into increasingly small
-pieces~\cite{virieux2009seismic}. Secondly, this method is broadly
-used on conventional clusters but can be drastically more complicated
-to seup and slower due to the communication in a cloud-based
-setup. 
-Another common strategy in seismic inversion is to only store values
-at the boundaries of the domain at each timestep, and reconstruct the
-rest of the wavefield when
-required~\cite{clapp2009reverse,yang2014rtm} with time reversal of the
-wave equation. However, this method is not applicable for wave
-equations that are not time reversible when for example physical
-attenuation is included.
-
-Checkpointing is yet another strategy to reduce the memory overhead. Only a
-subset of the program states during the forward pass is stored (and the rest
-discarded). The discarded data is recomputed when needed by restarting the
-forward pass from the last available stored state. The Revolve
-algorithm~\cite{griewank2000algorithm} provides an answer to the
-question of which states should be stored and which states should be recomputed to minimize the total
-amount of recomputation work. Other authors have subsequently
-developed extensions to Revolve that are optimal under different
-assumptions~\cite{stumm2009multistage, wang2009minimal,aupy2016optimal,schanen2016asynchronous}.
-
-\begin{figure}
-\begin{center}
-\input{dataflow}
-\end{center}
-\caption{The dataflow pattern that is typical of adjoint-based optimization problems}
-\label{fig:dataflow}
-\end{figure}
-
-Data compression is also increasingly used to reduce the memory footprint of
-scientific applications. General purpose data compression algorithms like Zlib
-(which is a part of gzip)~\cite{deutsch1996zlib}, and compression algorithms for
-video and image data such as JPEG-2000~\cite{skodras2001jpeg} have been
-presented in previous work. More recently, special purpose compression
-algorithms for floating-point scientific data have been developed, such as ZFP
-or SZ~\cite{Kaklamanis:2012aa,lindstrom2014fixed,di2018efficient}.
-
-Lossless algorithms guarantee that the exact original data can be recovered
-during decompression, whereas lossy algorithms introduce an error, but often
-guarantee that the error does not exceed certain absolute or relative error
-metrics. Typically, lossy compression is more effective in reducing the data
-size. Most popular compression packages offer various settings that allow a
-tradeoff between compression ratio, accuracy, and compression and decompression
-time.
-
-It is worth noting that another data reduction strategy is to typecast values
-into a lower precision format, for example, from double precision to single
-precision. This can be seen as a computationally cheap lossy compression
-algorithm with a compression ratio of $2$.
-
-Perhaps counterintuitively, compression can not only reduce the memory
-footprint, but also speed up an application. Previous work has observed that the
-compression and decompression time can be less than the time saved from the
-reduction in data that needs to be communicated across MPI nodes or between a
-GPU and a host computer~\cite{gpu-compression}.
-
-Another way of using compression to speed up adjoint-based methods is to use it \emph{instead} of
-checkpointing. If the compression ratio is sufficient to fit the
-entire data in memory, checkpointing is no longer necessary. Previous
-work has discussed this in the context of computational
-fluid dynamics~\cite{cyr2015towards} and seismic inversion using
-compression algorithms specifically designed for
-wavefields~\cite{dalmau2014lossy,boehm2016wavefield}.
-
-In this paper, we extend the previous studies by \emph{combining} checkpointing
-and compression. This is obviously useful when the data does not fit in the
-available memory even after compression, for example for very large adjoint
-problems, or for problems where the required accuracy limits the achievable
-compression ratios.
-
-Compared to the use of only checkpointing without compression, our
-combined method often improves performance. This is a consequence of
-the reduced size of stored program states, allowing more program
-states to be stored during the forward computation. This in turn
-reduces the amount of recomputation that needs to be performed. On the
-other hand, the compression and decompression itself takes time. We
-therefore provide a comprehensive performance model that predicts
-whether the combined method is beneficial, that is, whether the time
-spent compressing and decompressing is less than the time saved by
-reduced recomputations.
-
-For benchmarking we used a dual-socket
-Intel(R) Xeon(R) Platinum 8180M @ 2.50 Ghz (28 cores each), henceforth
-called Skylake. We discuss performance results for varying error tolerances for
-the lossy ZFP compression algorithm ranging from $10^0$ to $10^{-15}$, and different
-types of forward and adjoint computations that vary in their ratio between
-compute cost and state size.
-
-Our aim is to make this discussion independent of the hardware, application and
-compression algorithm. We therefore do not discuss whether or not the accuracy
-of the decompressed data is sufficient for our application, or whether
-other algorithms might achieve a better accuracy/compression tradeoff.
-
-\section{Test case using Devito and pyRevolve}
+\section{Problem and test case}
 We use Devito~\cite{devito-api,devito-compiler} to solve forward and adjoint
@@
 The values of $m(x)$ used in this work are derived from the Overthrust
-model~\cite{aminzadeh1996three} over a grid of $287 \times 881 \times 881$
-points, including an absorbing layer of 40 points on each side. The grid spacing
-is $25m$ in space. The propagation time is $4sec$ that corresponds to 2500 timesteps. The wave field
-at the final time is shown in Figure~\ref{fig:uncompressed}, and only this field
-was used for the compression experiments in this paper. The uncompressed size of
-this single time step field is just under 900MB.
+model~\cite{aminzadeh1996three} over a grid of $287 \times 881 \times
+881$ points, including an absorbing layer of 40 points on each
+side. The grid spacing is $25m$ in space. The propagation time is
+$4sec$ that corresponds to 2500 timesteps. The wave field at the final
+time is shown in Figure~\ref{fig:uncompressed}. The uncompressed size
+of this single time step field is just under 900MB. If one were to
+store all the timesteps, this would require 2.3TB of memory.
 
@@
 To implement Revolve with Devito, we use pyRevolve~\cite{kukreja2018high} which
-is a python wrapper for the Revolve algorithm. The performance model in
+is a python library to manage the execution of checkpointed adjoint computations. The performance model in
 section~\ref{sec:performance_model} assumes that the implementation is similar
@@
 
-\section{Compression algorithms}
-\subsection{Lossless}
-We use the python package \emph{blosc}~\cite{blosc}, which includes implementations for
-six different lossless compression algorithms, namely ZLIB, ZSTD, BLOSCLZ,
-LZ4, LZ4HC and Snappy. An exhaustive search over all available parameters 
-resulted in a maximum compression ratio of 1.18x. We therefore focus on lossy
-compression in the remainder of this work.
-
-\subsection{Lossy}
-We use the lossy compression package ZFP~\cite{lindstrom2014fixed} developed in
-C. To use ZFP from python, we developed a python wrapper for the reference
-implementation of ZFP.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.8\linewidth]{images/errors.pdf}
-\end{center}
-\caption{Cross-section of the field that shows errors introduced
-  during compression and decompression using the fixed-tolerance
-  mode. It is interesting to note that the errors are more or less
-  evenly distributed across the domain with only slight variations
-  corresponding to the wave amplitude (from the field plot in Figure
-  \ref{fig:uncompressed}. A small block-like structure characteristic of
-  ZFP can be seen.}
-\label{fig:decompressed_error}
-\end{figure}
-
-ZFP supports three compression modes, namely fixed-tolerance, fixed-precision
-and fixed-rate. The fixed-tolerance mode limits the absolute error, while the
-fixed-precision mode limits the relative error to some user-specified value.
-The fixed-rate mode achieves a guaranteed compression ratio requested by the
-user, but does not provide any bounds on accuracy loss.
-Figure~\ref{fig:tolerance_cf_plot} shows compression ratios for a variety of
-settings for the three modes.
-
-The compression ratio achieved per unit time spent during compressing
-and decompressing was seen to be highest in the fixed-tolerance mode
-albeit with highly unpredictable compression ratios. Figure
-\ref{fig:decompressed_error} shows the spatial distribution of the
-error after compression and decompression, compared to the original
-field, using fixed-tolerance mode.
-
-In our experiments using Devito, the time required to compress a checkpoint was
-in the same order of magnitude as the time taken to compute a timestep.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.9\linewidth]{images/tolerance-cf-richter.pdf}
-\end{center}
-\caption{Compression ratios achieved on compressing the wavefield. We
-  define compression ratio as the ratio between the size of the
-  uncompressed data and the compressed data. The dashed line
-  represents no compression. The highlighted point corresponds to the
-  setting used for the other results here unless otherwise specified.}
-\label{fig:tolerance_cf_plot}
-\end{figure}
-
-\section{Performance model for a combination of Revolve and compression}
-\label{sec:performance_model}
-We assume that the computation of a single forward time step takes the same wall
-time as the computation of a single reverse time step. We denote this time as
-$\mathbf{C}$. For a simulation with $\mathbf{N}$ timesteps, the minimum wall time required
-for the full forward-adjoint evaluation is given by
-\begin{equation}
-T_N = 2 \cdot \mathbf{C} \cdot \mathbf{N}.
-\end{equation}
-If the size of a single timestep in memory is given by $\mathbf{S}$, this
-requires a memory of at least size $\mathbf{S} \cdot \mathbf{N}$. If sufficient memory
-is available, no checkpointing or compression is needed.
-
-If the memory is smaller than $\mathbf{S} \cdot \mathbf{N}$, Revolve provides
-a strategy to solve for the adjoint field by storing a subset of the $\mathbf{N}$ total checkpoints
-and recomputing the remaining ones. The overhead introduced by this method can be broken down into
-the recomputation overhead $\mathbf{O}_R$ and the storage overhead $\mathbf{O}_S$. The recomputation
-overhead is the amount of time spent in recomputation, given by
-\begin{equation}
-\mathbf{O}_R(N, M) = p(N, M) \cdot \mathbf{C},
-\end{equation}
-where $p(N, M)$ is the minimum number of recomputed steps from \cite{griewank2000algorithm}, reproduced
-here in equation \ref{eqn:recompute}.
-\begin{figure*}
-\begin{equation}
-p(N, M) = \begin{cases}
-      N(N-1) /2, & \text{if}\ M=1 \\
-      \min\limits_{1<=\widetilde{N}<=N} \{\widetilde{N} + p(\widetilde{N}, M) + p(N-\widetilde{N}, M-1)\}, & \text{if}\ M>1
-    \end{cases}
-    \label{eqn:recompute}
-\end{equation}
-\end{figure*}
-In equation \ref{eqn:recompute}, M is the number of checkpoints that can be
-stored in memory. Note that for $M >=N$, $\mathbf{O}_R$ would be zero. For $M <
-N$, $\mathbf{O}_R$ grows rapidly as M is reduced relative to N.
-
-In an ideal implementation, the storage overhead $\mathbf{O}_S$ might be zero, since the computation could
-be done ``in-place'', but in practice, checkpoints are generally stored in a separate section of memory and they
-need to be transferred to a ``computational'' section of the memory where the compute is performed, and then
-the results copied back to the checkpointing memory. This copying is a common feature of checkpointing
-implementations, and might pose a non-trivial overhead in some scenarios. 
-This storage overhead is given by
-\begin{equation}
-\mathbf{O}_{SR}(N, M) = \mathbf{W}(N, M) \cdot \frac{\mathbf{S}}{\mathbf{B}} +
-\mathbf{R}(N, M) \cdot \frac{\mathbf{S}}{\mathbf{B}}
-\label{eqn:storage}
-\end{equation}
-where $\mathbf{W}$ is the total number of times Revolve writes
-checkpoints for a single run, $ \mathbf{R}$ is the number of times
-checkpoints are read, and $\mathbf{B}$ is the memory bandwidth of the
-target system. The total time to solution becomes
-\begin{equation}
-T_R = 2 \cdot \mathbf{C} \cdot \mathbf{N} + \mathbf{O}_R(N, M) +
-\mathbf{O}_{SR}(N, M)
-\end{equation} 
-By using compression, the size of each checkpoint is reduced and therefore the
-number of checkpoints available is increased ($M$ in equation
-\ref{eqn:recompute}). This reduces the recomputation overhead $\mathbf{O}_R$,
-while at the same time adding overheads related to compression and decompression
-in $\mathbf{O}_S$.
-To be beneficial, the reduction in $\mathbf{O}_R$ must offset the increase in 
-$\mathbf{O}_{SR}$, leading to an overall decrease in the time to solution $T$.
-
-Our performance model assumes
- that the compression algorithm behaves uniformly
-across the different time steps of the simulation, i.e. that we get the same compression ratio, compression time and 
-decompression time, no matter which of the $N$ possible checkpoints we try to compress/decompress. The storage overhead
-now becomes
-\begin{equation}
-\begin{split}
-\mathbf{O}_{SR}(N, M) = &\mathbf{W}(N, M \cdot F) \cdot \left(\frac{\mathbf{S}}{\mathbf{F}
-  \cdot \mathbf{B}} + t_c\right) +\\&\mathbf{R}(N, M \cdot F) \cdot
-\left(\frac{\mathbf{S}}{\mathbf{F} \cdot \mathbf{B}} + t_d\right)
-\end{split}
-\end{equation}
-where $\mathbf{F}$ is the compression ratio (i.e. the ratio between the uncompressed and compressed checkpoint), and $t_c$
-and $t_d$ are compression and decompression times, respectively. At the same
-time, the recomputation overhead decreases
-because $\mathbf{F}$ times more checkpoints are now available.
-
-\section{Results}
-\label{sec:results}
-We can distinguish three different scenarios, depending on the amount of
-available memory.
-\begin{enumerate}
-\item If the memory is insufficient even with compression to store the entire
-trajectory, one can either use checkpointing only, or combine checkpointing with
-compression.
-\item If the available memory is not sufficient to store the uncompressed
-trajectory, but large enough to store the entire compressed trajectory, we study
-the two possible strategies: Either use compression only, or use checkpointing
-only.
-\item If the available system memory is large enough to hold the entire
-uncompressed trajectory, neither compression nor checkpointing is necessary.
-\end{enumerate}
-All three scenarios can be seen in Figure~\ref{fig:varying_memory}.
-The second scenario was studied in previous work~\cite{cyr2015towards}, while
-the combined method is also applicable to the first scenario, for which previous
-work has only used checkpointing without compression.
-
-We can identify a number of factors that make compression more likely to be
-beneficial compared to pure checkpointing: A very small system memory size and a
-large number of time steps lead to a rapidly increasing recompute factor, and
-compression can substantially reduce this recompute factor. This can be seen in
-Figures~\ref{fig:varying_memory} and~\ref{fig:varying_nt}.
-
-The extent to which the recompute factor affects the overall runtime also
-depends on the cost to compute each individual time step. If the compute cost
-per time step is large compared to the compression and decompression cost, then
-compression is also likely to be beneficial, as shown in
-Figure~\ref{fig:varying_compute}. As the time per time step increases and the
-compression cost becomes negligible, we observe that the ratio between the runtime
-of the combined method and that of pure checkpointing is only determined by the
-difference in recompute factors.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.9\linewidth]{images/varying-memory.pdf}
-\end{center}
-\caption{The speedups predicted by the performance model for varying
-  memory. The baseline
-(1.0) is the performance of a Revolve-only implementation under the
-same conditions. The different curves represent kernels with differing
-compute times (represented here as a factor of the sum of compression
-and decompression times). The first vertical line at $~$ 53GB marks the
-spot where the compressed wavefield can completely fit in memory and
-Revolve is unnecessary if using compression. The second vertical line
-at $~$ 2.2 TB marks the spot where the entire uncompressed wavefield can
-fit in memory and neither Revolve nor compression is necessary. The
-region to the right is where these optimisations are not necessary or
-relevant. The middle region has been the subject of past studies using
-compression in adjoint problems. The region to the left is the focus
-of this paper.}
-\label{fig:varying_memory}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.9\linewidth]{images/varying-compute.pdf}
-\end{center}
-\caption{The speedups predicted by the performance model for varying
-  compute cost. The baseline
-(1.0) is the performance of a Revolve-only implementation under the
-same conditions. The benefits of compression drop rapidly if the
-computational cost of the kernel that generated the data is much lower
-than the cost of compressing the data. For increasing computational
-costs, the benefits are bounded.}
-\label{fig:varying_compute}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.9\linewidth]{images/varying-nt.pdf}
-\end{center}
-\caption{The speedups predicted by the performance model for varying
-  number of timesteps to be reversed. The baseline
-(1.0) is the performance of a Revolve-only implementation under the
-same conditions. It can be seen that compression becomes more
-beneficial as the number of timesteps is increased.}
-\label{fig:varying_nt}
-\end{figure}
-
-\section{Conclusions and Future work}
-We use lossy compression to reduce the computational overhead of
-checkpointing in an adjoint computation used in full waveform
-inversion, a common method in seismic imaging applications whose
-memory footprint commonly exceeds the available memory size in high
-performance computing systems. We also developed a performance model
-that computes whether or not the combination of compression and
-checkpointing will outperform pure checkpointing or pure compression
-in a variety of scenarios, depending on the available memory size,
-computational intensity of the application, and compression ratio and
-throughput of the compression algorithm. Our current result has
-several limitations that we plan to address in future work:
-\begin{itemize}
-\item We do not discuss the accuracy of the results after decompression. This
-depends on the application, compression algorithm, and affects the achievable
-compression ratios. Our performance model only requires knowledge of the
-compression time and ratio, and it is up to the user of this model to determine
-what accuracy is needed and thus what compression ratio is realistic for their
-application.
-\item We only present results using the ZFP compression library. It
-would be interesting to try a greater variety of compression
-algorithms.
-\item ZFP only supports serial decompression. If ZFP supported
-parallel decompression, our experiments would likely show a geater region in
-which the combined method is faster than pure checkpointing. Furthermore, ZFP
-only supports fields with up to three dimensions, while
-exploiting similarities between fields at different time steps may yield a
-better compression ratio.
-\item Our performance model is based on uniform compression ratios and
-times.  However, many applications, including FWI, are likely to have
-initial conditions that contain little information and are easily
-compressed, and the compression ratio gradually declines as the field
-becomes more complex. We based our experiments on the final wave
-field, which is presumably difficult to compress.
-\item In comparing pure compression with pure checkpointing, we assume
-that every checkpoint is compressed and decompressed. However, if the
-available memory is only slightly less than the required memory, an
-implementation that compresses only a subset of the checkpoints might
-outperform the expectations of our model.
-\item We do not discuss multi-level checkpointing, where some
-checkpoints are stored on a slower, larger device. We expect
-compression to be beneficial in these scenarios due to reduced data
-transfer sizes.
-\end{itemize}
-
-\section*{Acknowledgment}
-This work was funded by the Intel Parallel Computing Centre at
-Imperial College London and EPSRC EP/R029423/1. 
-This work was supported by the U.S. Department of Energy, Office of Science,
-Office of Advanced Scientific Computing Research, Applied Mathematics and
-Computer Science programs under contract number DE-AC02-06CH11357.
-We would also like to acknowledge the support from the SINBAD II project and
-the member organizations of the SINBAD Consortium.
-
-We gratefully acknowledge the computing resources provided and operated by the
-Joint Laboratory for System Evaluation (JLSE) at Argonne National Laboratory.
-
-\bibliographystyle{plain}
-\bibliography{compression}
-
-\vfill
-\begin{flushright}
-\normalsize
-\framebox{\parbox{3in}{The submitted manuscript has been created by UChicago
-Argonne, LLC, Operator of Argonne National Laboratory (`Argonne'). Argonne, a
-U.S. Department of Energy Office of Science laboratory, is operated under
-Contract No. DE-AC02-06CH11357. The U.S. Government retains for itself, and
-others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide
-license in said article to reproduce, prepare derivative works, distribute
-copies to the public, and perform publicly and display publicly, by or on behalf
-of the Government.  The Department of Energy will provide public access to these
-results of federally sponsored research in accordance with the DOE Public Access
-Plan.\newline \url{http://energy.gov/downloads/doe-public-access-plan}.}}
-\end{flushright}
-
-\end{document}
+For benchmarking we used a dual-socket
+Intel(R) Xeon(R) Platinum 8180M @ 2.50 Ghz (28 cores each), henceforth
+called Skylake.

### 2019-01-12 paper/introduction.tex
Update before overleaf

\section{Introduction}

\subsection{Adjoint-based optimization}
Adjoint-based optimization problems typically consist of a simulation
that is run forward in simulation time, producing data that is used in
reverse order by a subsequent adjoint computation that is run
backwards in simulation time.  Figure~\ref{fig:dataflow} shows the
resulting data flow. Besides seismic inversion, many important
numerical problems in science and engineering use adjoints and follow
the same pattern.

Since the data for each of the computed timesteps in the forward
simulation will be used later in the adjoint computation, it would be
prudent to store it in memory until it is required again, if the required amount
of memory is indeed available. However, the total size of this data
can often run into tens of terabytes and the management of this data
becomes a problem in itself. This data management problem is the
subject of this paper.

\begin{figure}
\begin{center}
\input{dataflow}
\end{center}
\caption{The dataflow pattern that is typical of adjoint-based optimization problems}
\label{fig:dataflow}
\end{figure}

\subsection{Example adjoint problem: Seismic inversion}
Seismic inversion typically involves the simulation of the propagation
of seismic waves through the earth's subsurface, followed by a
comparison with data from field measurements. The model of the
subsurface is iteratively improved by minimizing the misfit between
simulated data and field measurement in an adjoint optimization
problem.  Figure~\ref{fig:offshore_survey} shows the setup of the
field experiment that produces the measured
data~\cite{plessix2006review}.

The data collected in an offshore survey typically consists of a
number of ``shots'' - each of these shots corresponding to different
locations of sources and receivers. As a loose analogy with machine
learning, these correspond to different data points. Since the
gradient computation over a single shot is complex enough that a
single shot can occupy a complete node for $~10^1-10^2$ minutes, the gradient is computed for each of these
shots independently and then collated across all the shots to form a
single update that is used to update the model. It might be evident
here that the processing across shots is easy to parallelise since it
requires a small amount of communication followed by a relatively long
period of independent computation as part of a single iteration of the
optimisation. Since the number of shots is typically of the order of
$10^4$, this offers ample opportunity to fill up a large cluster with
computation, even if an individual shot is only processed on a single
node of the cluster at a time. Hence, it might be worthy to note that
even though this paper focusses on a single gradient evaluation for a
single shot, the overall problem involves carrying out $~10^5$ such
evaluations and is typically run on large clusters for non-trivial
amounts of time.

The first part of seismic inversion, i.e. the simulation of seismic
wave propagation through the earth's subsurface is typically done
using a finite-difference solver and is called the ``forward problem''
in the context of inversion. Looking at this part in isolation, the
data flow here looks like the one shown in
figure~\ref{fig:dataflowfw}. This illustration assumes a first-order
time-stepper, i.e. the computation of each timestep only depends on
the previous timestep. In such a scenario, only two timesteps need to
be kept in memory - the last computed step and the one currently being
computed. In case of an n-th order time-stepper, $(n+1)$ timesteps
need to be kept in memory at any one time. Hence, the memory
requirements of the forward problem can remain constant, regardless of
the number of timesteps the simulation may be run for.

\begin{figure*}
\begin{center}
\input{dataflowfw}
\end{center}
\caption{The dataflow pattern typical of a forward-only
  simulation. Boxes represent data and arrows represent computation.}
\label{fig:dataflowfw}
\end{figure*}

TODO remove contractions

\subsection{Memory requirements}
A number of strategies are regularly employed to deal with this
enormous volume of data - the simplest of these being to store it to a
disk, to be read later by the adjoint pass in reverse order. However,
typically the computation to be done on this data in the adjoint phase
takes much less time than the time taken to read it from the disk,
reading from the disk becomes the bottleneck for most practical cases,
rendering this method, although possibly the simplest, also the
slowest of the possible alternatives. Seeing this from the perspective
where thousands of such computations might be running in parallel on a
single cluster (for different shots), the network bandwidth might
restrict the use of a network storage further, hence only node-local disks
may be suitable for this strategy.

Domain decomposition, where a single shot may be distributed across more
than one node, is often used not only to distribute the computational
workload across more processors, but also to utilize the large amount of memory
available in distributed systems. While this strategy is very powerful, the
number of compute nodes and therefore the amount of memory that can be used
efficiently is limited, for example by communication overheads that start to
dominate as the domain is split into increasingly small
pieces~\cite{virieux2009seismic}. Secondly, this strategy can lead to
a wastage of resources and sometimes a longer time-to-solution for a
given inversion problem using a given number of nodes, especially when
the number of nodes is less than the number of shots. For example, a
problem setup that requires only 10\% more memory than is available on
a single node might not be a good candidate for domain decomposition
over multiple nodes. Lastly, this method is even less applicable on
cloud-based setups since it can be drastically more complicated
to setup and slower due to the communication.

Another common strategy in seismic inversion is to only store values
at the boundaries of the domain at each timestep, and reconstruct the
rest of the wavefield when
required~\cite{clapp2009reverse,yang2014rtm} with time reversal of the
wave equation. However, this method is not applicable for wave
equations that are not time reversible when for example physical
attenuation is included.

Checkpointing is yet another strategy to reduce the memory
overhead. Only a subset of the timesteps during the forward pass is
stored (and the rest discarded). The discarded data is recomputed when
needed by restarting the forward pass from the last available stored
state. We discuss this strategy in section~\ref{revolve}.

Another strategy commonly employed to reduce the memory footprint of
such applications is data compression. This is discussed in
section~\ref{compression}.

In this paper, we extend the previous studies by \emph{combining} checkpointing
and compression. This is obviously useful when the data does not fit in the
available memory even after compression, for example for very large adjoint
problems, or for problems where the required accuracy limits the achievable
compression ratios.

Compared to the use of only checkpointing without compression, this
combined method often improves performance. This is a consequence of
the reduced size of stored timesteps, allowing more timesteps to be
stored during the forward computation. This in turn reduces the amount
of recomputation that needs to be performed. On the other hand, the
compression and decompression itself takes time. The answer to the
question ``does compression pay off?'', depends on a number of factors including - available
memory, the required precision, the time taken to compress and
decompress, and the achieved compression factors, and various problem specific
parameters like computational intensity of the kernel involved in the
forward and adjoint computations, and the number of timesteps.

Hence, the answer to the compression question depends not only on the
problem one is solving (within seismic inversion, there are numerous
variations of the wave equation that may be solved), but also the
hardware specifics of the machine on which it is being solved. In
fact, as we will see in section TODO, the answer might even change
during the solution process of an individual problem. This brings up
the need to be able to predict whether compression would pay off in a
given scenario, without incurring significant overheads in answering
this question. In this paper, we present the use of a performance
model to answer that question.

\subsection{Summary of contributions}
\begin{itemize}
\item Study the use of different compression algorithms to seismic
  data including 6 lossless and the two most popular lossy compression
  algorithms for floating point data. 
\item Study the compressibility behaviour of seismic data in terms of
  compression factors and compression and decompression times under
  these algorithms
\item Study the performance model for Revolve alone, that takes into
  account time taken to read and write checkpoints
\item An online performance model to predict whether compression would speed up
  an optimization problem
\end{itemize}

### 2019-01-12 paper/perf-revolve.tex
Update before overleaf

\section{Revolve: Performance model}
\label{sec:revolve}
Checkpointing is a commonly used strategy to reduce the memory footprint
of adjoint problems. Here, depending on the memory available, some timesteps
computed in the forward pass are stored, while others are discarded. The ones
that were discarded are later recomputed by rerunning the forward pass from
the last stored checkpoint. 
The Revolve algorithm~\cite{griewank2000algorithm} provides an
answer to the question of which timesteps should be stored and which
states should be recomputed to minimize the total amount of
recomputation work. Other authors have subsequently developed
extensions to Revolve that are optimal under different
assumptions~\cite{stumm2009multistage,
wang2009minimal,aupy2016optimal,schanen2016asynchronous, aupy2017periodicity}. Previous
work has applied checkpointing to seismic imaging and inversion
problems~\cite{symes2007reverse, datta2018asynchronous}.

In this section, we build on the ideas introduced in \cite{stumm2009multistage} to build a performance model that can be used to predict the runtime of an adjoint computation that uses the Revolve checkpointing strategy. 
We call the time taken by a single forward computational step $C_F$
and correspondingly, the time taken by a single backward step $C_R$. For a simulation with $\mathbf{N}$ timesteps, the minimum wall time required
for the full forward-adjoint evaluation is given by
\begin{equation}
T_N = \mathbf{C_F} \cdot \mathbf{N} + \mathbf{C_R} \cdot \mathbf{N}
\end{equation}
If the size of a single timestep in memory is given by $\mathbf{S}$, this
requires a memory of at least size $\mathbf{S} \cdot \mathbf{N}$. If sufficient memory
is available, no checkpointing or compression is needed.

If the memory is smaller than $\mathbf{S} \cdot \mathbf{N}$, Revolve provides
a strategy to solve for the adjoint field by storing a subset of the $\mathbf{N}$ total checkpoints
and recompute the remaining ones. The overhead introduced by this method can be broken down into
the recomputation overhead $\mathbf{O}_R$ and the storage overhead $\mathbf{O}_S$. The recomputation
overhead is the amount of time spent in recomputation, given by
\begin{equation}
\mathbf{O}_R(N, M) = p(N, M) \cdot \mathbf{C_F},
\end{equation}
where $p(N, M)$ is the minimum number of recomputed steps from \cite{griewank2000algorithm}, reproduced
here in equation \ref{eqn:recompute}.
\begin{figure*}
\begin{equation}
p(N, M) = \begin{cases}
      N(N-1) /2, & \text{if}\ M=1 \\
      \min\limits_{1<=\widetilde{N}<=N} \{\widetilde{N} + p(\widetilde{N}, M) + p(N-\widetilde{N}, M-1)\}, & \text{if}\ M>1
    \end{cases}
    \label{eqn:recompute}
\end{equation}
\end{figure*}
In equation \ref{eqn:recompute}, M is the number of checkpoints that can be
stored in memory. Note that for $M >=N$, $\mathbf{O}_R$ would be zero. For $M <
N$, $\mathbf{O}_R$ grows rapidly as M is reduced relative to N.

In an ideal implementation, the storage overhead $\mathbf{O}_S$ might be zero, since the computation could
be done ``in-place'', but in practice, checkpoints are generally stored in a separate section of memory and they
need to be transferred to a ``computational'' section of the memory where the computation is performed, and then
the results copied back to the checkpointing memory. This copying is a common feature of checkpointing
implementations, and might pose a non-trivial overhead when the
computation involved in a single timestep is not very large. 
This storage overhead is given by:
\begin{equation}
\mathbf{O}_{SR}(N, M) = \mathbf{W}(N, M) \cdot \frac{\mathbf{S}}{\mathbf{B}} +
\mathbf{N} \cdot \frac{\mathbf{S}}{\mathbf{B}}
\label{eqn:storage}
\end{equation}
where $\mathbf{W}$ is the total number of times Revolve writes
checkpoints for a single run, $ \mathbf{N}$ is the number of times
checkpoints are read, and $\mathbf{B}$ is the bandwidth at which these
copies happen. The total time to solution becomes
\begin{equation}
T_R = \mathbf{C_F} \cdot \mathbf{N} + \mathbf{C_R} \cdot \mathbf{N} + \mathbf{O}_R(N, M) +
\mathbf{O}_{SR}(N, M)
\end{equation}

### 2019-01-12 temp_plot.py

@@ -1,0 +1,16 @@
+import sys
+from ast import literal_eval
+from plotter import plot
+
+infile = sys.argv[1]
+
+data = []
+with open(infile) as datafile:
+    for i, row in enumerate(datafile):
+        data.append(literal_eval(row))
+
+indices, ratios, times = zip(*data)
+
+
+plot(indices, ratios, filename='compression-ratios-through-simulation.pdf', xlabel='Simulation time (n)', ylabel='Compression ratio (x)', title='Variation of achievable compression ratio as the simulation progresses', xscale='linear')
+        

### 2019-01-15 lossless.py
Full compression runs

@@ -1,11 +1,13 @@
 import blosc
 import sys
 import h5py
-from bloscpack import pack_nd_array_to_bytes, unpack_ndarray_from_bytes
+import numpy as np
+from bloscpack import pack_ndarray_to_bytes, unpack_ndarray_from_bytes, BloscArgs
 from itertools import product
+from contexttimer import Timer
 
 
-chunk_size_options = [int(2**i) for i in numpy.arange(19, 23.5, 0.5)]
+chunk_size_options = [int(2**i) for i in np.arange(19, 23.5, 0.5)]
 shuffle_options = [blosc.SHUFFLE, blosc.BITSHUFFLE, blosc.NOSHUFFLE]
 cname_options = blosc.cnames
 clevel_options = list(range(10))
@@ -28,21 +30,24 @@ def compress(arr, chunk_size, shuffle, cname, clevel):
 def decompress(compressed):
     return unpack_ndarray_from_bytes(compressed)
 
-
+print(sys.argv)
 filename = sys.argv[1]
 f = h5py.File(filename, 'r')
-uncompressed = f['data'][()].astype(np.dtype('float32'))
+input_array = f['data'][()].astype(np.dtype('float32'))
 
 combinations = product(chunk_size_options, shuffle_options, cname_options, clevel_options)
 
 for op_set in combinations:
-    with Timer(factor=1000) as compress_time:
-        compressed = compress(input_array, *op_set)
+    try:
+        with Timer(factor=1000) as compress_time:
+            compressed = compress(input_array, *op_set)
 
-    compression_factor = (input_array.size * input_array.item_size) / len(compressed)
+        compression_factor = (input_array.size * input_array.itemsize) / len(compressed)
 
-    with Timer(factor=1000) as decompress_time:
-        decompressed = decompress(compressed)
+        with Timer(factor=1000) as decompress_time:
+            decompressed = decompress(compressed)
 
-    assert(np.allclose(input_array, decompressed))
-    print("%f, %d, %s, %d, %f, %f, %f" % (op_set[0], op_set[1], op_set[2], op_set[3], compress_time, decompress_time, compression_factor))
+        assert(np.allclose(input_array, decompressed))
+        print("%f, %d, %s, %d, %f, %f, %f" % (op_set[0], op_set[1], op_set[2], op_set[3], compress_time.elapsed, decompress_time.elapsed, compression_factor))
+    except:
+        print("Error")

### 2019-01-15 plot_numpy_hdf5.py

@@ -18,10 +18,12 @@ model = from_hdf5(velocity_model, space_order=2, nbpml=20, datakey='m0', dtype=n
 shape = model.vp.shape
 vmax = np.max(data)
 
-slice_loc = 440
-im = plt.imshow(np.transpose(data[slice_loc]), vmax=.01, vmin=-.01, cmap="seismic",
+if len(data.shape)==3:
+    slice_loc = 440
+    im = plt.imshow(np.transpose(data[slice_loc]), vmax=.01, vmin=-.01, cmap="seismic",
            extent = [0, 20, 0.001*(shape[-1]-1)*25, 0])
-
+else:
+    im = plt.imshow(np.transpose(data), vmax=vmax, vmin=-vmax, cmap="seismic")
 plt.xlabel("X (km)")
 plt.ylabel("Depth (km)")
 cb = plt.colorbar(shrink=.3, pad=.01, aspect=10)

### 2019-01-15 timing_experiment.py

@@ -16,7 +16,7 @@ from timeit import default_timer
 from simple import overthrust_setup
 
 from examples.seismic.acoustic.acoustic_example import acoustic_setup
-
+from util import to_hdf5
 
 class Timer(object):
     def __init__(self, tracker):
@@ -34,12 +34,13 @@ class Timer(object):
         self.tracker.append(self.elapsed)
         
 
-def verify(space_order=4, kernel='OT4', nbpml=40, filename='', **kwargs):
+def verify(space_order=4, kernel='OT4', nbpml=40, filename='', compression_params={}, **kwargs):
     solver = acoustic_setup(shape=(10, 10), spacing=(10, 10), nbpml=10, tn=50,
                             space_order=space_order, kernel=kernel, **kwargs)
     #solver = overthrust_setup(filename=filename, tn=50, nbpml=nbpml, space_order=space_order, kernel=kernel, **kwargs)
     
     u = TimeFunction(name='u', grid=solver.model.grid, time_order=2, space_order=solver.space_order)
+
     rec = Receiver(name='rec', grid=solver.model.grid,
                               time_range=solver.receiver.time_range,
                               coordinates=solver.receiver.coordinates.data)
@@ -51,8 +52,10 @@ def verify(space_order=4, kernel='OT4', nbpml=40, filename='', **kwargs):
     grad = Function(name='grad', grid=solver.model.grid)
     wrap_fw = CheckpointOperator(solver.op_fwd(save=False), src=solver.source, u=u, m=m, rec=rec, dt=dt)
     wrap_rev = CheckpointOperator(solver.op_grad(save=False), u=u, v=v, m=m, rec=rec, dt=dt, grad=grad)
-
-    wrp = Revolver(cp, wrap_fw, wrap_rev, n_checkpoints, rec.data.shape[0]-2)
+    nt = rec.data.shape[0] - 2
+    print("Verifying for %d timesteps" % nt)
+    wrp = Revolver(cp, wrap_fw, wrap_rev, n_checkpoints, nt,
+                   compression_params=compression_params)
     wrp.apply_forward()
     summary = wrp.apply_reverse()
     print(wrp.profiler.timings)
@@ -63,14 +66,16 @@ def verify(space_order=4, kernel='OT4', nbpml=40, filename='', **kwargs):
     with Timer([]) as tr:
         grad2, _ = solver.gradient(rec=rec2, u=u2)
 
[…]

### 2019-01-15 tolerance.py

@@ -28,15 +28,23 @@ else:
     from pysz import compress, decompress
 
 f = h5py.File(filename, 'r')
-uncompressed = f['data'][()].astype(np.dtype('float32'))
+uncompressed = f['data'][()].astype(np.dtype('float64'))
 print("\"Size of compressed field\", \"Compression Factor\", \"Compression time\", \"Decompression time\", \"Tolerance\", \"Error norm\", \"Maximum error\"")
 for p_i in range(0, 16):
     tolerance = 0.1**p_i
     with Timer(factor=1000) as t:
-        compressed = compress(uncompressed, tolerance=tolerance, parallel=parallel)
+        if compressor=="zfp":
+            kwargs = {'parallel':parallel, 'tolerance': tolerance}
+        else:
+            kwargs = {'tolerance': tolerance}
+        compressed = compress(uncompressed, **kwargs)
     
     with Timer(factor=1000) as t2:
-        decompressed = decompress(compressed, uncompressed.shape, uncompressed.dtype, tolerance=tolerance, parallel=parallel)
+        if compressor=="zfp":
+            kwargs = {'parallel': parallel, 'tolerance': tolerance}
+        else:
+            kwargs = {}
+        decompressed = decompress(compressed, uncompressed.shape, uncompressed.dtype, **kwargs)
 
     #to_hdf5(decompressed, "decompressed-t-%d.h5"%p_i)
     error_matrix = decompressed-uncompressed

### 2019-06-06 performance_model_validation.py
Performance model validation

@@ -1,19 +1,47 @@
 from performance import Problem
 from plot_timings import consolidated_results_from_file
 import sys
+import matplotlib.pyplot as plt
 
 
 def revolve_theoretical(nt, size_ts, compute_ts, bw):
     p = Problem(nt, size_ts, compute_ts, bw)
     min_time, overall_peak_mem = p.naive_strategy()
-    mems = range(2, 50) * size_ts
+    mems = [x*size_ts for x in list(range(2, 50))]
     times = [sum(p.revolve(x)) for x in mems]
     return times, mems
 
 def revolve_actual(filename):
     results, consolidated_results, field_names, ncp = consolidated_results_from_file(filename)
     total_exec_times = [sum(results[n].values()) for n in ncp]
-    return total_exec_times
+    return total_exec_times, ncp
 
+nt = 631
+size_ts = 2*287*881*881*4/1000000
+#No compression
+compute_ts = 550
+bw = 1.9
 
-revolve_actual(sys.argv[1])
+#Compression OT2
+compute_ts = 550
+bw = 1.2
+
+#Compression OT4
+compute_ts = 1000
+bw = 1.2
+
+filename = "model_validation_compression_ot4.pdf"
+
+measured_exec_times, ncp = revolve_actual(sys.argv[1])
+measured_mems = [x*size_ts for x in ncp]
+predicted_times, mems = revolve_theoretical(nt, size_ts, compute_ts, bw)
+
+plt.plot(mems, predicted_times, label="Predicted")
+plt.plot(measured_mems, measured_exec_times, label="Measured")
+plt.yscale('log')
+plt.xscale('log', basex=2)
+plt.legend()
+plt.xlabel("Peak memory (MB)")
+plt.ylabel("Total time to solution (ms)")
+plt.title("Comparison of predicted and measured runtimes for OT4 kernel with compression")
+plt.savefig(filename, bbox_inches='tight')

### 2019-06-06 paper/compression.tex

\section{Compression algorithms}

Data compression is increasingly used to reduce the memory footprint of
scientific applications. General purpose data compression algorithms like Zlib
(which is a part of gzip)~\cite{deutsch1996zlib}, and compression algorithms for
video and image data such as JPEG-2000~\cite{skodras2001jpeg} have been
presented in previous work. More recently, special purpose compression
algorithms for floating-point scientific data have been developed, such as ZFP
or SZ~\cite{lindstrom2014fixed,di2018efficient}.

Lossless algorithms guarantee that the exact original data can be recovered
during decompression, whereas lossy algorithms introduce an error, but often
guarantee that the error does not exceed certain absolute or relative error
metrics. Typically, lossy compression is more effective in reducing the data
size. Most popular compression packages offer various settings that allow a
tradeoff between compression ratio, accuracy, and compression and decompression
time.

Another comonly-observed difference between lossless and lossy
compression algorithms is that lossless compression algorithms tend to
interpret all data as one-dimensional series only while SZ and ZFP,
being designed for scientific data, tend to take the dimensionality
into account directly. This makes a difference in the case of a
wavefield, for example, where the data to be compressed corresponds to
a smoothly varying function in (two or) three dimensions and
interpreting this three-dimensional data as one-dimensional would
completely miss the smoothness and predictability of the data values.

It is worth noting that another data reduction strategy is to typecast values
into a lower precision format, for example, from double precision to single
precision. This can be seen as a computationally cheap lossy compression
algorithm with a compression ratio of $2$.

Perhaps counterintuitively, compression can not only reduce the memory
footprint, but also speed up an application. Previous work has observed that the
compression and decompression time can be less than the time saved from the
reduction in data that needs to be communicated across MPI nodes or between a
GPU and a host computer~\cite{gpu-compression}.

One way of using compression in adjoint-based methods is to compress
all the timesteps during the forward pass. If the compression ratio is
sufficient to fit the entire data in memory, this enables solving an
adjoint-based optimisation problem without resorting to any of the
other techniques previously discussed here. Specifically, compression
serves as an \emph{alternate strategy} to checkpointing in this
scenario. Previous work has discussed this in the context of
computational fluid dynamics~\cite{cyr2015towards,marin2016large} and seismic
inversion using compression algorithms specifically designed for
the respective applications~\cite{dalmau2014lossy,boehm2016wavefield}.

Since the time spent on compressing and decompressing data is often
non-negligible, this raises the question whether the computational
time is better spent on this compression and decompression, or on the
recomputation involved in the more traditional checkpointing
approach. This question was previously answered to a limited extent
for the above scenario where compression is an alternative to
checkpointing, in a specific application~\cite{cyr2015towards}. We discuss
that in section TODO.

\subsection{Lossless}
We use the python package \emph{blosc}~\cite{blosc}, which includes implementations for
six different lossless compression algorithms, namely ZLIB, ZSTD, BLOSCLZ,
LZ4, LZ4HC and Snappy. All these algorithms look at the data as a one-dimensional stream of bits
and at least the blosc implementations have a limit on the size of the one-dimensional array that
can be compressed in one call. Therefore we use the python package \emph{blosc-pack}, which is
a wrapper over the blosc library, to implement \emph{chunking}, i.e. breaking up the stream into
chunks of a chosen size, which are compressed one at a time.

\subsection{Lossy}
\subsubsection{ZFP}
We use the lossy compression package ZFP~\cite{lindstrom2014fixed} developed in
C. To use ZFP from python, we developed a python wrapper for the reference
implementation of ZFP \footnote{To be released open source on publication}.

ZFP supports three compression modes, namely fixed-tolerance, fixed-precision
and fixed-rate. The fixed-tolerance mode limits the absolute error, while the
fixed-precision mode limits the error as a ratio of the range of values in the array to be compressed.
The fixed-rate mode achieves a guaranteed compression ratio requested by the
user, but does not provide any bounds on accuracy loss.

\subsection{SZ}
SZ~\cite{di2018efficient} is a more recently developed compression library, also focussed on lossy compression
of floating-point scientific data, also developed in C. We also wrote a python wrapper for the reference
implementation of SZ to use it as part of our benchmark suite. \footnote{Also to be released open source
upon publication}

SZ supports four compression modes, namely absolute error mode, which, similar to ZFP's fixed-tolerance
mode, allows the user to control the maximum pointwise error in absolute values. The relative ratio mode 
of SZ allows the user to specify a maximum error as a ratio of the range of values in the array, which is
effectively similar to ZFP's fixed-precision mode but not exactly. SZ has two other modes that are missing 
in ZFP, namely pointwise relative error and pointwise SNR mode. In the pointwise relative error mode, the user
can provide a relative error ratio and SZ will ensure that the error at each point is within that ratio, considering
its absolute value. In the pointwise SNR mode, the user provides a signal-to-noise ratio value that SZ respects
at each point.

\subsection{Combining lossy and lossless compression}
Another approach we attempted is a combination of lossy and lossless compression 
schemes to achieve an overall lossless compression scheme. Here, the array is first compressed using a 
lossy compression scheme, following which the errors incurred by this lossy scheme are passed on to a 
lossless scheme for compression. The idea is that the distribution of errors incurred by a lossy compression
algorithm might make it more favourable for lossless compression than the original array.

### 2019-06-06 paper/errors-convergence.tex

\section{Acceptable errors and convergence}
\label{sec:errors}

Our performance model is designed to be agnostic of the specific adjoint-based 
optimization problem being solved. This is because we envision its use in a generic
checkpointing runtime that manages the checkpointed execution of the optimization
problem that accepts an acceptable error tolerance as an input parameter for each
gradient evaluation and determines whether or not compression can pay off for that
iteration, and if yes, which of the available strategies is to be used. This last question 
has previously been addressed previously in literature but in more specific contexts 
\cite{kunkel2017toward, tao2018optimizing}.

One question that arises in evaluating derivatives on grids compressed (and decompressed)
using lossy compression is the numerical stability of the computed derivatives, since errors
in neighbouring points can accumulate in the derivative rather quickly, rendering the derivatives
unusable. This question was addressed for ZFP \cite{zfp-derivatives} and SZ\cite{tao2017z}
separately.

In the context of seismic inversion, it has been shown before that the precision required
in the gradient evaluation is very low in the beginning of the optimization and
accurate gradients are not needed until the optimization is close to a minimum 
\cite{van20143d, boehm2016wavefield}. This is perhaps quite intuitive since, being far
from a minimum in the beginning, a gradient pointing in the approximate direction of the
relevant minimum is sufficient to make progress. These initial iterations could use a more
aggressive lossy compression strategy to accelerate (through compression) the progress 
towards the minimum. Once the optimization is within the vicinity of the minimum, the
gradient is required at a higher accuracy to make any progress, and this performance
model can then dynamically decide to disable compression for those iterations where
a more accurate, albeit slower, gradient evaluation is preferred.

There is also a body of work that addresses convergence guarantees of trust-region
based optimization methods in the presence of unreliable gradients. This was primarily
done for the scenario where the gradient (and sometimes the functional itself) is known
with a probability $p$.~\cite{blanchet2016convergence,cartis2017global,chen2018stochastic}
It was shown here that the convergence rate is only affected by a factor that is 
a linear function of $p$. This analytical framework could be extended to provide bounds
on the accuracy required in a particular gradient evaluation in order to guarantee a certain
convergence rate.

Both these analyses stop at the required accuracy in the gradient evaluation. This needs
to be extended to derive acceptable error tolerances in individual grid points corresponding to a specific
error bound in the overall gradient evaluation.

### 2019-06-06 paper/paper.tex

@@
 
-\title{Combining checkpointing and data compression\\ for large scale seismic inversion}
+\title{Combining checkpointing and data compression to accelerate
+  adjoint-based optimization problems}
 
@@
 Atlanta, GA, USA\\}
-\and
-\IEEEauthorblockN{Kaiyuan Hou}
-\IEEEauthorblockA{%\textit{Department of Electrical Engineering and Computer Science} \\
-\textit{Northwestern University}\\
-Evanston, IL, USA \\
-}
-\and
-\IEEEauthorblockN{Fabio Luporini}
-\IEEEauthorblockA{%\textit{Department of Earth Science and Engineering} \\
-\textit{Imperial College London}\\
-London, UK \\
-}
+%\and
+%\IEEEauthorblockN{Kaiyuan Hou}
+%\IEEEauthorblockA{%\textit{Department of Electrical Engineering and Computer Science} \\
+%\textit{Northwestern University}\\
+%Evanston, IL, USA \\
+%}
+%\and
+%\IEEEauthorblockN{Fabio Luporini}
+%\IEEEauthorblockA{%\textit{Department of Earth Science and Engineering} \\
+%\textit{Imperial College London}\\
+%London, UK \\
+%}
 \and
@@
 \begin{IEEEkeywords}
-Checkpointing, compression, adjoints, inversion
+Checkpointing, compression, adjoints, inversion, seismic
 \end{IEEEkeywords}
 
-\section{Introduction}
-Adjoint-based optimization problems typically consist of a simulation that is
-run forward in simulation time, producing data that is used in reverse order by
-a subsequent adjoint computation that is run backwards in simulation time.
-Figure~\ref{fig:dataflow} shows the resulting data flow. Many important
-applications in science and engineering have this structure, including
-seismic inversion.
+\input{introduction}
 
-In seismic inversion, the propagation of seismic waves through the earth's subsurface is simulated and compared with data from field measurements.
-The model of the subsurface is iteratively improved by minimizing the misfit between simulated data and field measurement in an adjoint optimization problem.
-Figure~\ref{fig:offshore_survey} shows the setup of the field experiment that produces the measured data~\cite{plessix2006review}. Storing all intermediate
-results as required for the data flow shown in Figure~\ref{fig:dataflow} is not
-feasible, since this would require tens of terabytes of memory, which is not
-commonly available on the machines used to solve these problems. To solve this
-problem, a number of strategies are used that are discussed in the following
-paragraphs. While this paper uses seismic inversion as a test case, most of these ideas, including the one presented in this work, are also applicable to other adjoint optimization problems.
+\input{compression.tex}
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.75\linewidth]{images/survey-ship-diagram.jpg}
-\end{center}
-\caption{Illustration of an offshore seismic survey that collects the
-  input data for seismic inversion. Source: The Open University \cite{surveyship}}
-\label{fig:offshore_survey}
-\end{figure}
+\input{perf-revolve.tex}
 
-Domain decomposition is often used not only to distribute the computational
-workload across more processors, but also to utilize the large amount of memory
-available in distributed systems. While this strategy is very powerful, the
-number of compute nodes and therefore the amount of memory that can be used
-efficiently is limited, for example by communication overheads that start to
-dominate as the domain is split into increasingly small
-pieces~\cite{virieux2009seismic}. Secondly, this method is broadly
-used on conventional clusters but can be drastically more complicated
-to seup and slower due to the communication in a cloud-based
-setup. 
-Another common strategy in seismic inversion is to only store values
-at the boundaries of the domain at each timestep, and reconstruct the
-rest of the wavefield when
-required~\cite{clapp2009reverse,yang2014rtm} with time reversal of the
-wave equation. However, this method is not applicable for wave
-equations that are not time reversible when for example physical
-attenuation is included.
+\input{perf-compression.tex}
 
-Checkpointing is yet another strategy to reduce the memory overhead. Only a
-subset of the program states during the forward pass is stored (and the rest
-discarded). The discarded data is recomputed when needed by restarting the
-forward pass from the last available stored state. The Revolve
-algorithm~\cite{griewank2000algorithm} provides an answer to the
-question of which states should be stored and which states should be recomputed to minimize the total
-amount of recomputation work. Other authors have subsequently
-developed extensions to Revolve that are optimal under different
-assumptions~\cite{stumm2009multistage, wang2009minimal,aupy2016optimal,schanen2016asynchronous}.
+\input{errors-convergence.tex}
 
-\begin{figure}
-\begin{center}
-\input{dataflow}
-\end{center}
-\caption{The dataflow pattern that is typical of adjoint-based optimization problems}
-\label{fig:dataflow}
-\end{figure}
+\input{devito.tex}
 
-Data compression is also increasingly used to reduce the memory footprint of
-scientific applications. General purpose data compression algorithms like Zlib
-(which is a part of gzip)~\cite{deutsch1996zlib}, and compression algorithms for
-video and image data such as JPEG-2000~\cite{skodras2001jpeg} have been
-presented in previous work. More recently, special purpose compression
-algorithms for floating-point scientific data have been developed, such as ZFP
-or SZ~\cite{Kaklamanis:2012aa,lindstrom2014fixed,di2018efficient}.
+\input{results-discussion.tex}
 
-Lossless algorithms guarantee that the exact original data can be recovered
-during decompression, whereas lossy algorithms introduce an error, but often
-guarantee that the error does not exceed certain absolute or relative error
-metrics. Typically, lossy compression is more effective in reducing the data
-size. Most popular compression packages offer various settings that allow a
-tradeoff between compression ratio, accuracy, and compression and decompression
-time.
+\input{conclusions-futurework.tex}
 
-It is worth noting that another data reduction strategy is to typecast values
-into a lower precision format, for example, from double precision to single
-precision. This can be seen as a computationally cheap lossy compression
-algorithm with a compression ratio of $2$.
-
-Perhaps counterintuitively, compression can not only reduce the memory
-footprint, but also speed up an application. Previous work has observed that the
-compression and decompression time can be less than the time saved from the
-reduction in data that needs to be communicated across MPI nodes or between a
-GPU and a host computer~\cite{gpu-compression}.
-
-Another way of using compression to speed up adjoint-based methods is to use it \emph{instead} of
-checkpointing. If the compression ratio is sufficient to fit the
-entire data in memory, checkpointing is no longer necessary. Previous
-work has discussed this in the context of computational
-fluid dynamics~\cite{cyr2015towards} and seismic inversion using
-compression algorithms specifically designed for
-wavefields~\cite{dalmau2014lossy,boehm2016wavefield}.
-
-In this paper, we extend the previous studies by \emph{combining} checkpointing
-and compression. This is obviously useful when the data does not fit in the
-available memory even after compression, for example for very large adjoint
-problems, or for problems where the required accuracy limits the achievable
-compression ratios.
-
-Compared to the use of only checkpointing without compression, our
-combined method often improves performance. This is a consequence of
-the reduced size of stored program states, allowing more program
-states to be stored during the forward computation. This in turn
-reduces the amount of recomputation that needs to be performed. On the
-other hand, the compression and decompression itself takes time. We
-therefore provide a comprehensive performance model that predicts
-whether the combined method is beneficial, that is, whether the time
-spent compressing and decompressing is less than the time saved by
-reduced recomputations.
-
-For benchmarking we used a dual-socket
-Intel(R) Xeon(R) Platinum 8180M @ 2.50 Ghz (28 cores each), henceforth
-called Skylake. We discuss performance results for varying error tolerances for
-the lossy ZFP compression algorithm ranging from $10^0$ to $10^{-15}$, and different
-types of forward and adjoint computations that vary in their ratio between
-compute cost and state size.
-
-Our aim is to make this discussion independent of the hardware, application and
-compression algorithm. We therefore do not discuss whether or not the accuracy
-of the decompressed data is sufficient for our application, or whether
-other algorithms might achieve a better accuracy/compression tradeoff.
-
-\section{Test case using Devito and pyRevolve}
-We use Devito~\cite{devito-api,devito-compiler} to solve forward and adjoint
-wave equation problems. Devito is a domain-specific language that enables the
-rapid development of finite-difference solvers from a high-level description
-of partial differential equations. The simplest version of the seismic wave
-equation is the acoustic isotropic wave equation defined as:
-\begin{equation}
-m(x)\frac{\partial^2 u(t, x)}{\partial t^2} - \Laplace u(t, x) = q(t, x),
-\label{eqn:wave}
-\end{equation}
-where $m(x) = \frac{1}{c^2(x)}$ is the squared slowness, $c(x)$ the spatially
-dependent speed of sound, $u(t, x)$ is the pressure wavefield, $\Laplace u(t, x)$ denotes the laplacian of the wavefield and $q(t, x)$ is a source term.
-Some of our experiments in Section \ref{sec:results} are performed using a more accurate
-and more complex version of this equation called Tilted Transverse Isotropy
-(TTI)~\cite{zhang2011stable} that takes into account the anisotopic propagation of waves in the earth subsurface (directional dependency of the speed of sound).
-We leave the TTI equations out of this paper
-for brevity.
-
-The solution to equation~\ref{eqn:wave} forms the forward problem. The seismic inversion problem minimizes the misfit between simulated and observed signal given by:
-\begin{equation}
-\min_{m} \phi_s(m) = \frac{1}{2} \left\lVert d_{sim} - d_{obs} \right\rVert_2^2.
-\end{equation}
-
-This optimization problem is usually solved using gradient based methods such as steepest descent,
-where the gradient is computed using the adjoint-state method that involves
-the data-flow pattern from Figure \ref{fig:dataflow}.
-
-The values of $m(x)$ used in this work are derived from the Overthrust
-model~\cite{aminzadeh1996three} over a grid of $287 \times 881 \times 881$
-points, including an absorbing layer of 40 points on each side. The grid spacing
-is $25m$ in space. The propagation time is $4sec$ that corresponds to 2500 timesteps. The wave field
-at the final time is shown in Figure~\ref{fig:uncompressed}, and only this field
-was used for the compression experiments in this paper. The uncompressed size of
-this single time step field is just under 900MB.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.8\linewidth]{images/uncompressed.pdf}
-\end{center}
-\caption{Cross-section of the wavefield used as a reference sample for
-  compression and decompression. This field was formed after a Ricker
-  wavelet source was placed at the surface of the model and the wave propagated for 2500
-  timesteps. This is a vertical (x-z) cross-section of a 3D field, taken at
-  the $y$ source location}
-\label{fig:uncompressed}
-\end{figure}
-
-To implement Revolve with Devito, we use pyRevolve~\cite{kukreja2018high} which
-is a python wrapper for the Revolve algorithm. The performance model in
-section~\ref{sec:performance_model} assumes that the implementation is similar
-to pyRevolve, which stores a checkpoint by copying a portion of the operator's
-working memory to the checkpointing memory and similarly loads a checkpoint by
-copying from the checkpointing memory to the operator's working memory. Although
-a perfect implementation of checkpointing may be able to avoid these copies, the
-overhead attached to these copies can be ignored for an operator that is
-sufficiently computationally intensive. However, we include the overheads in the
-model to verify this assumption.
-
-\section{Compression algorithms}
-\subsection{Lossless}
-We use the python package \emph{blosc}~\cite{blosc}, which includes implementations for
-six different lossless compression algorithms, namely ZLIB, ZSTD, BLOSCLZ,
-LZ4, LZ4HC and Snappy. An exhaustive search over all available parameters 
-resulted in a maximum compression ratio of 1.18x. We therefore focus on lossy
-compression in the remainder of this work.
-
-\subsection{Lossy}
-We use the lossy compression package ZFP~\cite{lindstrom2014fixed} developed in
-C. To use ZFP from python, we developed a python wrapper for the reference
-implementation of ZFP.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.8\linewidth]{images/errors.pdf}
-\end{center}
-\caption{Cross-section of the field that shows errors introduced
-  during compression and decompression using the fixed-tolerance
-  mode. It is interesting to note that the errors are more or less
-  evenly distributed across the domain with only slight variations
-  corresponding to the wave amplitude (from the field plot in Figure
-  \ref{fig:uncompressed}. A small block-like structure characteristic of
-  ZFP can be seen.}
-\label{fig:decompressed_error}
-\end{figure}
-
-ZFP supports three compression modes, namely fixed-tolerance, fixed-precision
-and fixed-rate. The fixed-tolerance mode limits the absolute error, while the
-fixed-precision mode limits the relative error to some user-specified value.
-The fixed-rate mode achieves a guaranteed compression ratio requested by the
-user, but does not provide any bounds on accuracy loss.
-Figure~\ref{fig:tolerance_cf_plot} shows compression ratios for a variety of
-settings for the three modes.
-
-The compression ratio achieved per unit time spent during compressing
-and decompressing was seen to be highest in the fixed-tolerance mode
-albeit with highly unpredictable compression ratios. Figure
-\ref{fig:decompressed_error} shows the spatial distribution of the
-error after compression and decompression, compared to the original
-field, using fixed-tolerance mode.
-
-In our experiments using Devito, the time required to compress a checkpoint was
-in the same order of magnitude as the time taken to compute a timestep.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.9\linewidth]{images/tolerance-cf-richter.pdf}
-\end{center}
-\caption{Compression ratios achieved on compressing the wavefield. We
-  define compression ratio as the ratio between the size of the
-  uncompressed data and the compressed data. The dashed line
-  represents no compression. The highlighted point corresponds to the
-  setting used for the other results here unless otherwise specified.}
-\label{fig:tolerance_cf_plot}
-\end{figure}
-
-\section{Performance model for a combination of Revolve and compression}
-\label{sec:performance_model}
-We assume that the computation of a single forward time step takes the same wall
-time as the computation of a single reverse time step. We denote this time as
-$\mathbf{C}$. For a simulation with $\mathbf{N}$ timesteps, the minimum wall time required
-for the full forward-adjoint evaluation is given by
-\begin{equation}
-T_N = 2 \cdot \mathbf{C} \cdot \mathbf{N}.
-\end{equation}
-If the size of a single timestep in memory is given by $\mathbf{S}$, this
-requires a memory of at least size $\mathbf{S} \cdot \mathbf{N}$. If sufficient memory
-is available, no checkpointing or compression is needed.
-
-If the memory is smaller than $\mathbf{S} \cdot \mathbf{N}$, Revolve provides
-a strategy to solve for the adjoint field by storing a subset of the $\mathbf{N}$ total checkpoints
-and recomputing the remaining ones. The overhead introduced by this method can be broken down into
-the recomputation overhead $\mathbf{O}_R$ and the storage overhead $\mathbf{O}_S$. The recomputation
-overhead is the amount of time spent in recomputation, given by
-\begin{equation}
-\mathbf{O}_R(N, M) = p(N, M) \cdot \mathbf{C},
-\end{equation}
-where $p(N, M)$ is the minimum number of recomputed steps from \cite{griewank2000algorithm}, reproduced
-here in equation \ref{eqn:recompute}.
-\begin{figure*}
-\begin{equation}
-p(N, M) = \begin{cases}
-      N(N-1) /2, & \text{if}\ M=1 \\
-      \min\limits_{1<=\widetilde{N}<=N} \{\widetilde{N} + p(\widetilde{N}, M) + p(N-\widetilde{N}, M-1)\}, & \text{if}\ M>1
-    \end{cases}
-    \label{eqn:recompute}
-\end{equation}
-\end{figure*}
-In equation \ref{eqn:recompute}, M is the number of checkpoints that can be
-stored in memory. Note that for $M >=N$, $\mathbf{O}_R$ would be zero. For $M <
-N$, $\mathbf{O}_R$ grows rapidly as M is reduced relative to N.
-
-In an ideal implementation, the storage overhead $\mathbf{O}_S$ might be zero, since the computation could
-be done ``in-place'', but in practice, checkpoints are generally stored in a separate section of memory and they
-need to be transferred to a ``computational'' section of the memory where the compute is performed, and then
-the results copied back to the checkpointing memory. This copying is a common feature of checkpointing
-implementations, and might pose a non-trivial overhead in some scenarios. 
-This storage overhead is given by
-\begin{equation}
-\mathbf{O}_{SR}(N, M) = \mathbf{W}(N, M) \cdot \frac{\mathbf{S}}{\mathbf{B}} +
-\mathbf{R}(N, M) \cdot \frac{\mathbf{S}}{\mathbf{B}}
-\label{eqn:storage}
-\end{equation}
-where $\mathbf{W}$ is the total number of times Revolve writes
-checkpoints for a single run, $ \mathbf{R}$ is the number of times
-checkpoints are read, and $\mathbf{B}$ is the memory bandwidth of the
-target system. The total time to solution becomes
-\begin{equation}
-T_R = 2 \cdot \mathbf{C} \cdot \mathbf{N} + \mathbf{O}_R(N, M) +
-\mathbf{O}_{SR}(N, M)
-\end{equation} 
-By using compression, the size of each checkpoint is reduced and therefore the
-number of checkpoints available is increased ($M$ in equation
-\ref{eqn:recompute}). This reduces the recomputation overhead $\mathbf{O}_R$,
-while at the same time adding overheads related to compression and decompression
-in $\mathbf{O}_S$.
-To be beneficial, the reduction in $\mathbf{O}_R$ must offset the increase in 
-$\mathbf{O}_{SR}$, leading to an overall decrease in the time to solution $T$.
-
-Our performance model assumes
- that the compression algorithm behaves uniformly
-across the different time steps of the simulation, i.e. that we get the same compression ratio, compression time and 
-decompression time, no matter which of the $N$ possible checkpoints we try to compress/decompress. The storage overhead
-now becomes
-\begin{equation}
-\begin{split}
-\mathbf{O}_{SR}(N, M) = &\mathbf{W}(N, M \cdot F) \cdot \left(\frac{\mathbf{S}}{\mathbf{F}
-  \cdot \mathbf{B}} + t_c\right) +\\&\mathbf{R}(N, M \cdot F) \cdot
-\left(\frac{\mathbf{S}}{\mathbf{F} \cdot \mathbf{B}} + t_d\right)
-\end{split}
-\end{equation}
-where $\mathbf{F}$ is the compression ratio (i.e. the ratio between the uncompressed and compressed checkpoint), and $t_c$
-and $t_d$ are compression and decompression times, respectively. At the same
-time, the recomputation overhead decreases
-because $\mathbf{F}$ times more checkpoints are now available.
-
-\section{Results}
-\label{sec:results}
-We can distinguish three different scenarios, depending on the amount of
-available memory.
-\begin{enumerate}
-\item If the memory is insufficient even with compression to store the entire
-trajectory, one can either use checkpointing only, or combine checkpointing with
-compression.
-\item If the available memory is not sufficient to store the uncompressed
-trajectory, but large enough to store the entire compressed trajectory, we study
-the two possible strategies: Either use compression only, or use checkpointing
-only.
-\item If the available system memory is large enough to hold the entire
-uncompressed trajectory, neither compression nor checkpointing is necessary.
-\end{enumerate}
-All three scenarios can be seen in Figure~\ref{fig:varying_memory}.
-The second scenario was studied in previous work~\cite{cyr2015towards}, while
-the combined method is also applicable to the first scenario, for which previous
-work has only used checkpointing without compression.
-
-We can identify a number of factors that make compression more likely to be
-beneficial compared to pure checkpointing: A very small system memory size and a
-large number of time steps lead to a rapidly increasing recompute factor, and
-compression can substantially reduce this recompute factor. This can be seen in
-Figures~\ref{fig:varying_memory} and~\ref{fig:varying_nt}.
-
-The extent to which the recompute factor affects the overall runtime also
-depends on the cost to compute each individual time step. If the compute cost
-per time step is large compared to the compression and decompression cost, then
-compression is also likely to be beneficial, as shown in
-Figure~\ref{fig:varying_compute}. As the time per time step increases and the
-compression cost becomes negligible, we observe that the ratio between the runtime
-of the combined method and that of pure checkpointing is only determined by the
-difference in recompute factors.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.9\linewidth]{images/varying-memory.pdf}
-\end{center}
-\caption{The speedups predicted by the performance model for varying
-  memory. The baseline
-(1.0) is the performance of a Revolve-only implementation under the
-same conditions. The different curves represent kernels with differing
-compute times (represented here as a factor of the sum of compression
-and decompression times). The first vertical line at $~$ 53GB marks the
-spot where the compressed wavefield can completely fit in memory and
-Revolve is unnecessary if using compression. The second vertical line
-at $~$ 2.2 TB marks the spot where the entire uncompressed wavefield can
-fit in memory and neither Revolve nor compression is necessary. The
-region to the right is where these optimisations are not necessary or
-relevant. The middle region has been the subject of past studies using
-compression in adjoint problems. The region to the left is the focus
-of this paper.}
-\label{fig:varying_memory}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.9\linewidth]{images/varying-compute.pdf}
-\end{center}
-\caption{The speedups predicted by the performance model for varying
-  compute cost. The baseline
-(1.0) is the performance of a Revolve-only implementation under the
-same conditions. The benefits of compression drop rapidly if the
-computational cost of the kernel that generated the data is much lower
-than the cost of compressing the data. For increasing computational
-costs, the benefits are bounded.}
-\label{fig:varying_compute}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.9\linewidth]{images/varying-nt.pdf}
-\end{center}
-\caption{The speedups predicted by the performance model for varying
-  number of timesteps to be reversed. The baseline
-(1.0) is the performance of a Revolve-only implementation under the
-same conditions. It can be seen that compression becomes more
-beneficial as the number of timesteps is increased.}
-\label{fig:varying_nt}
-\end{figure}
-
-\section{Conclusions and Future work}
-We use lossy compression to reduce the computational overhead of
-checkpointing in an adjoint computation used in full waveform
-inversion, a common method in seismic imaging applications whose
-memory footprint commonly exceeds the available memory size in high
-performance computing systems. We also developed a performance model
-that computes whether or not the combination of compression and
-checkpointing will outperform pure checkpointing or pure compression
-in a variety of scenarios, depending on the available memory size,
-computational intensity of the application, and compression ratio and
-throughput of the compression algorithm. Our current result has
-several limitations that we plan to address in future work:
-\begin{itemize}
-\item We do not discuss the accuracy of the results after decompression. This
-depends on the application, compression algorithm, and affects the achievable
-compression ratios. Our performance model only requires knowledge of the
-compression time and ratio, and it is up to the user of this model to determine
-what accuracy is needed and thus what compression ratio is realistic for their
-application.
-\item We only present results using the ZFP compression library. It
-would be interesting to try a greater variety of compression
-algorithms.
-\item ZFP only supports serial decompression. If ZFP supported
-parallel decompression, our experiments would likely show a geater region in
-which the combined method is faster than pure checkpointing. Furthermore, ZFP
-only supports fields with up to three dimensions, while
-exploiting similarities between fields at different time steps may yield a
-better compression ratio.
-\item Our performance model is based on uniform compression ratios and
-times.  However, many applications, including FWI, are likely to have
-initial conditions that contain little information and are easily
-compressed, and the compression ratio gradually declines as the field
-becomes more complex. We based our experiments on the final wave
-field, which is presumably difficult to compress.
-\item In comparing pure compression with pure checkpointing, we assume
-that every checkpoint is compressed and decompressed. However, if the
-available memory is only slightly less than the required memory, an
-implementation that compresses only a subset of the checkpoints might
-outperform the expectations of our model.
-\item We do not discuss multi-level checkpointing, where some
-checkpoints are stored on a slower, larger device. We expect
-compression to be beneficial in these scenarios due to reduced data
-transfer sizes.
-\end{itemize}
-
-\section*{Acknowledgment}
+\section*{Acknowledgments}
 This work was funded by the Intel Parallel Computing Centre at
@@
 Joint Laboratory for System Evaluation (JLSE) at Argonne National Laboratory.
+
+TODO Kaiyuan, Fabio, Thomas Matthews, Paul Kelly, Oana Marin
 

### 2019-06-06 paper/perf-compression.tex

\section{Performance model including compression}

By using compression, the size of each checkpoint is reduced and therefore the
number of checkpoints available is increased ($M$ in equation
\ref{eqn:recompute}). This reduces the recomputation overhead $\mathbf{O}_R$,
while at the same time adding overheads related to compression and decompression
in $\mathbf{O}_S$.
To be beneficial, the reduction in $\mathbf{O}_R$ must offset the increase in 
$\mathbf{O}_{SR}$, leading to an overall decrease in the time to solution $T$.

Our performance model assumes
 that the compression algorithm behaves uniformly
across the different time steps of the simulation, i.e. that we get the same compression ratio, compression time and 
decompression time, no matter which of the $N$ possible checkpoints we try to compress/decompress. The storage overhead
now becomes
\begin{equation}
\begin{split}
\mathbf{O}_{SR}(N, M) = &\mathbf{W}(N, M \cdot F) \cdot \left(\frac{\mathbf{S}}{\mathbf{F}
  \cdot \mathbf{B}} + t_c\right) +\\&\mathbf{N} \cdot
\left(\frac{\mathbf{S}}{\mathbf{F} \cdot \mathbf{B}} + t_d\right)
\end{split}
\end{equation}
where $\mathbf{F}$ is the compression ratio (i.e. the ratio between the uncompressed and compressed checkpoint), and $t_c$
and $t_d$ are compression and decompression times, respectively. At the same
time, the recomputation overhead decreases
because $\mathbf{F}$ times more checkpoints are now available.

### 2019-06-06 paper/results-discussion.tex

\section{Results and discussion}
\label{sec:results}

TODO add figure about evolution of compressibility

To understand the compressibility of the data produced in a typical
wave-propagation simulation, we ran a simulation as per the setup
described in section TODO, and tried to compress every single timestep
of that simulation and noted the compression ratios achieved at every
timestep. As figure TODO shows, the initial timesteps are much easier
to compress than the later ones. This is not surprising since most
wave simulations start with the field at rest, i.e. filled with zeros.

As the wave reaches more parts of the domain, the field becomes less
compressible until it achieves a stable state when the wave has
reached most of the domain.

If the simulation had started with the field already oscillating in a
wave, it is likely that the compressibility curve for that simulation
would be flat.

This tells us that the compressibility of the last timestep of the
solution is representative of the worst-case compressibility and hence
we used the last timestep as our reference for comparison of
compression in the following section.

The compression ratio achieved per unit time spent during compressing
and decompressing was seen to be highest in the fixed-tolerance mode
albeit with highly unpredictable compression ratios. Figure~\ref{fig:tolerance_cf_plot} shows compression ratios for a variety of
settings for the fixed-tolerance mode. Figure
\ref{fig:decompressed_error} shows the spatial distribution of the
error after compression and decompression, compared to the original
field, using fixed-tolerance mode.
\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{images/errors.pdf}
\end{center}
\caption{Cross-section of the field that shows errors introduced
  during compression and decompression using the fixed-tolerance
  mode. It is interesting to note that the errors are more or less
  evenly distributed across the domain with only slight variations
  corresponding to the wave amplitude (from the field plot in Figure
  \ref{fig:uncompressed}. A small block-like structure characteristic of
  ZFP can be seen.}
\label{fig:decompressed_error}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.9\linewidth]{images/tolerance-cf-richter.pdf}
\end{center}
\caption{Compression ratios achieved on compressing the wavefield. We
  define compression ratio as the ratio between the size of the
  uncompressed data and the compressed data. The dashed line
  represents no compression. The highlighted point corresponds to the
  setting used for the other results here unless otherwise specified.}
\label{fig:tolerance_cf_plot}
\end{figure}

TODO: -Complete forward backward run to calculate gradient and angular
difference with true gradient

We can distinguish three different scenarios, depending on the amount of
available memory.
\begin{enumerate}
\item If the memory is insufficient even with compression to store the entire
trajectory, one can either use checkpointing only, or combine checkpointing with
compression.
\item If the available memory is not sufficient to store the uncompressed
trajectory, but large enough to store the entire compressed trajectory, we study
the two possible strategies: Either use compression only, or use checkpointing
only.
\item If the available system memory is large enough to hold the entire
uncompressed trajectory, neither compression nor checkpointing is necessary.
\end{enumerate}
All three scenarios can be seen in
Figure~\ref{fig:varying_memory}. TODO elaborate on which part of the
figure shows which scenario. 
The second scenario was studied in previous work~\cite{cyr2015towards}, while
the combined method is also applicable to the first scenario, for which previous
work has only used checkpointing without compression.

We can identify a number of factors that make compression more likely to be
beneficial compared to pure checkpointing: A very small system memory size and a
large number of time steps lead to a rapidly increasing recompute factor, and
compression can substantially reduce this recompute factor. This can be seen in
Figures~\ref{fig:varying_memory} and~\ref{fig:varying_nt}.

The extent to which the recompute factor affects the overall runtime also
depends on the cost to compute each individual time step. If the compute cost
per time step is large compared to the compression and decompression cost, then
compression is also likely to be beneficial, as shown in
Figure~\ref{fig:varying_compute}. As the time per time step increases and the
compression cost becomes negligible, we observe that the ratio between the runtime
of the combined method and that of pure checkpointing is only determined by the
difference in recompute factors.

\subsection{Validation of model}
TODO write this section with a plot where theoretical prediction is
shown overlaid with real measurements

Do this for: 
- Acoustic, where compression doesn't pay off
- TTI where compression performs better than recomputation

\begin{figure}
\begin{center}
\includegraphics[width=0.9\linewidth]{images/varying-memory.pdf}
\end{center}
\caption{The speedups predicted by the performance model for varying
  memory. The baseline
(1.0) is the performance of a Revolve-only implementation under the
same conditions. The different curves represent kernels with differing
compute times (represented here as a factor of the sum of compression
and decompression times). The first vertical line at $~$ 53GB marks the
spot where the compressed wavefield can completely fit in memory and
Revolve is unnecessary if using compression. The second vertical line
at $~$ 2.2 TB marks the spot where the entire uncompressed wavefield can
fit in memory and neither Revolve nor compression is necessary. The
region to the right is where these optimisations are not necessary or
relevant. The middle region has been the subject of past studies using
compression in adjoint problems. The region to the left is the focus
of this paper.}
\label{fig:varying_memory}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.9\linewidth]{images/varying-compute.pdf}
\end{center}
\caption{The speedups predicted by the performance model for varying
  compute cost. The baseline
(1.0) is the performance of a Revolve-only implementation under the
same conditions. The benefits of compression drop rapidly if the
computational cost of the kernel that generated the data is much lower
than the cost of compressing the data. For increasing computational
costs, the benefits are bounded.}
\label{fig:varying_compute}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.9\linewidth]{images/varying-nt.pdf}
\end{center}
\caption{The speedups predicted by the performance model for varying
  number of timesteps to be reversed. The baseline
(1.0) is the performance of a Revolve-only implementation under the
same conditions. It can be seen that compression becomes more
beneficial as the number of timesteps is increased.}
\label{fig:varying_nt}
\end{figure}
